---
title: "Final Vetting for ERS Comparative Database"
author: "Kira FitzGerald"
reviewer: "Jenny Katz"
date: "7/17/25"
output: 
 word_document:
    keep_md: true
    toc: yes
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
options(scipen = 10L)


```

# Part I: Load Data

This part of the script loads all R packages and data files that are used throughout the script.

## 1.1 Load Packages

**What:** this code block loads in all relevant function packages

```{r, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(erstools)
#loading functions used by R
library(readxl) #for loading excel files
library(tidyverse) #for all data analysis
library(data.table) #for all data analysis
library(tm) #for bag of words
library(textstem) #for turning words into their stem form
library(randomForest) #for random forest model
library(ranger) #for random forest model
library(xgboost) #for random forest model

```

## 1.2 Load Coded File

```{r}
#set sharepoint site URL for access to our org-wide files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"

#access sharepoint site URL you may be prompted to sign in here
#orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)

```

```{r}

#1. folder path to data
raw_data_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2025%20Cohort/1.%20Data%20&%20Analysis%20-%20Secure/2.%20Data%20Coding/"

#2. district folder
district_folder <-"Baltimore County"

file_name = paste0(district_folder, "_Homebase.xlsx")

```

```{r}

## For testing purposes ONLY - switch to your district's script as needed
final_coded_file <- ers_read_sharepoint(folder_path = raw_data_folder_path,
                   file_name_with_extension = file_name, skip_rows = 5, sheet_name = "Budgeted Positions")


student_enrollment <- 43479

```

## 1.3 Cleanup Coded File

```{r}

## This standardizes the column names of your file so the coding checks script can be run consistently across ERS coded fields. You should switch the column names as needed. All final coded fields should start with "C" and district fields should start with "D."

final_coded_file <- final_coded_file %>%
  mutate(H_Project = H_project,
         H_District = H_district,
         H_Year = H_project_year,
         C_op_status = C_operating_status_final,
         C_reporting = C_reporting_final,
         C_object_type = NA_character_,
         C_position_type = C_position_type_final,
         C_function = C_function_final,
         C_use = NA_character_,
         C_prek = C_prek_expense_final,
         C_student_type = C_student_type_final,
         C_sharing = C_sharing_final,
         D_FTE = M_fte_adjusted,
         M_FTE_500 = M_fte_adjusted / student_enrollment * 500,
         # Add any columns that might be helpful for your vetting purposes (they will be deleted later)
         #H_source = Source,
         .keep = "none")


```

# Part 2: Vetting

## Step 2.1 Catching typos and missing values

**What:** Make sure that all values are spelled correctly and align to ERS fields. For example, if you are seeing no Nurses, it might be because your district contracts them out - make sure that this is the case before proceeding.

```{r}

# making the final dataframe to append rows 
tier_1 <- tibble(
  Flag_Category = character(),
  Flag_Type = character(),
  ERS_Coded_Field = character(),
  Dollars_FTE_Flag = character(),
  Flag_Description = character(),
  Impact_Scaling = numeric()
) 


```

```{r}

## Operating Status

op_status_values <- c("Non-Expenditure",
                      "Non-Operating",
                      "Operating Not PreK-12",
                      "PreK-12 Operating")


#--------------------------------------------------------------

op_status_incorrect <- final_coded_file %>%
  filter(!C_op_status %in% op_status_values) %>%
  group_by(C_op_status) %>% 
  summarize(total_fte = sum(D_FTE, na.rm = TRUE), .groups = "drop") %>% 
  mutate(
    Flag_Category = "Missing or Misaligned Values",
    Flag_Type = "Red",
    ERS_Coded_Field = "Operating Status",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = paste("Misaligned Value:", C_op_status),
    Impact_Scaling = total_fte) %>% 
  select(Flag_Category, Flag_Type, ERS_Coded_Field, Dollars_FTE_Flag, Flag_Description, Impact_Scaling)

#-----------------------------------------------------------------------
op_status_missing <- tibble(C_op_status = op_status_values) %>%
  left_join(final_coded_file %>%
              group_by(C_op_status) %>%
              summarise(total_fte = sum(D_FTE, na.rm = TRUE),
                        .groups = "drop"),
            by = "C_op_status") %>%
  mutate(total_fte = ifelse(is.na(total_fte), 0, total_fte)) %>% 
  filter(total_fte == 0) %>%
  mutate(
    Flag_Category = "Missing or Misaligned Values",
    Flag_Type = "Yellow",
    ERS_Coded_Field = "Operating Status",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = paste("Missing Value:", C_op_status),
    Impact_Scaling = NA) %>% 
  select(Flag_Category, Flag_Type, ERS_Coded_Field, Dollars_FTE_Flag, Flag_Description, Impact_Scaling)

tier_1 <- bind_rows(tier_1, op_status_incorrect)
tier_1 <- bind_rows(tier_1, op_status_missing)


rm(op_status_incorrect, op_status_missing)
  
```

```{r}
# For all future checks, we filter to just PreK-12 Operating
#TO DO... THERE NEEDS TO BE SOME SORT OF CHECK THAT FIXES PREK12 OPERATING OTHERWISE THIS WILL FILTER OUT ALL OF THE 

final_coded_file <-final_coded_file %>%
  mutate(C_op_status = case_when(C_op_status == "PreK12 Operating"
         ~"PreK-12 Operating", TRUE ~ C_op_status))


final_coded_file <- final_coded_file %>% 
  filter(C_op_status == "PreK-12 Operating")

```

**ADDED FOR BOULDER**

```{r}
# #
# #TRIM Spaces so the rest of the white spaces will work
# 
# cols_to_clean <- c(
#   "C_op_status", "C_reporting", "C_object_type", 
#   "C_position_type", "C_function", "C_use", 
#   "C_prek", "C_student_type", "C_sharing"
# )
# 
# final_coded_file <- final_coded_file %>%
#   mutate(
#     across(
#       all_of(cols_to_clean),
#       ~ str_trim(.x),  # trims 
#       .names = "{.col}"
#     ))
```

```{r}
## Reporting

reporting_values <- c("School",
                      "Non-School")


reporting_incorrect <- final_coded_file %>%
  filter(!C_reporting %in% reporting_values) %>%
  group_by(C_reporting) %>% 
  summarize(total_fte = sum(D_FTE, na.rm = TRUE), .groups = "drop") %>% 
  mutate(
    Flag_Category = "Missing or Misaligned Values",
    Flag_Type = "Red",
    ERS_Coded_Field = "Reporting",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = paste("Misaligned Value:", C_reporting),
    Impact_Scaling = total_fte) %>% 
  select(Flag_Category, Flag_Type, ERS_Coded_Field, Dollars_FTE_Flag, Flag_Description, Impact_Scaling)

#-----------------------------------------------------------------------
reporting_missing <- tibble(C_reporting = reporting_values) %>%
  left_join(final_coded_file %>%
              group_by(C_reporting) %>%
              summarise(total_fte = sum(D_FTE, na.rm = TRUE),
                        .groups = "drop"),
            by = "C_reporting") %>%
  mutate(total_fte = ifelse(is.na(total_fte), 0, total_fte)) %>% 
  filter(total_fte == 0) %>%
  mutate(
    Flag_Category = "Missing or Misaligned Values",
    Flag_Type = "Yellow",
    ERS_Coded_Field = "Reporting",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = paste("Missing Value:", C_reporting),
    Impact_Scaling = NA) %>% 
  select(Flag_Category, Flag_Type, ERS_Coded_Field, Dollars_FTE_Flag, Flag_Description, Impact_Scaling)


tier_1 <- bind_rows(tier_1, reporting_incorrect)
tier_1 <- bind_rows(tier_1, reporting_missing)


rm(reporting_incorrect, reporting_missing)


```

```{r}
# 
# ## Object Type
# 
# object_values <- c("Base Salary/Compensation",
#                    "Substitute Compensation",
#                    "Other Compensation/Stipend",
#                    "Benefit",
#                    "Contracted Services",
#                    "Supplies/Materials",
#                    "Equipment & Equipment Lease",
#                    "Travel & Conferences",
#                    "Rent/Utilities",
#                    "Other Non-Compensation")
# 
# 
# 
# object_type_incorrect <- final_coded_file %>%
#   filter(!C_object_type %in% object_values) %>%
#   group_by(C_object_type) %>% 
#   summarize(total_fte = sum(D_FTE, na.rm = TRUE), .groups = "drop") %>% 
#   mutate(
#     Flag_Category = "Missing or Misaligned Values",
#     Flag_Type = "Red",
#     ERS_Coded_Field = "Object Type",
#     Dollars_FTE_Flag = "FTE",
#     Flag_Description = paste("Misaligned Value:", C_object_type),
#     Impact_Scaling = total_fte) %>% 
#   select(Flag_Category, Flag_Type, ERS_Coded_Field, Dollars_FTE_Flag, Flag_Description, Impact_Scaling)
# 
# #-----------------------------------------------------------------------
# object_type_missing <- tibble(C_object_type = object_values) %>%
#   left_join(final_coded_file %>%
#               group_by(C_object_type) %>%
#               summarise(total_fte = sum(D_FTE, na.rm = TRUE),
#                         .groups = "drop"),
#             by = "C_object_type") %>%
#   mutate(total_fte = ifelse(is.na(total_fte), 0, total_fte)) %>% 
#   filter(total_fte == 0) %>%
#   mutate(
#     Flag_Category = "Missing or Misaligned Values",
#     Flag_Type = "Yellow",
#     ERS_Coded_Field = "Object Type",
#     Dollars_FTE_Flag = "FTE",
#     Flag_Description = paste("Missing Value:", C_object_type),
#     Impact_Scaling = NA) %>% 
#   select(Flag_Category, Flag_Type, ERS_Coded_Field, Dollars_FTE_Flag, Flag_Description, Impact_Scaling)
# 
# 
# tier_1 <- bind_rows(tier_1, object_type_incorrect)
# tier_1 <- bind_rows(tier_1, object_type_missing)
# 
# 
# rm(object_type_incorrect, object_type_missing)

```

```{r}

## Position Type

position_values <- c("Area Officers",
                     "Club Advisor/Coach",
                     "Coordinator/Manager",
                     "Custodian",
                     "(Exec) Director",
                     "Guidance Counselor",
                     "Instructional Coach",
                     "Instructional Coach Supervisors",
                     "Librarian",
                     "Non-Position",
                     "Nurse",
                     "Nurse Aide",
                     "Occupational Therapist",
                     "Other",
                     "Physical Therapist",
                     "Principal",
                     "Psychologist",
                     "Sec/Clerk/Other Admin",
                     "School Monitor/Security",
                     "School Resource Officer/Police",
                     "Social Worker",
                     "Speech Therapist",
                     "Substitute",
                     "TA",
                     "Teacher",
                     "Vice Principal")


position_type_incorrect <- final_coded_file %>%
  filter(!C_position_type %in% position_values) %>%
  group_by(C_position_type) %>% 
  summarize(total_fte = sum(D_FTE, na.rm = TRUE), .groups = "drop") %>% 
  mutate(
    Flag_Category = "Missing or Misaligned Values",
    Flag_Type = "Red",
    ERS_Coded_Field = "Position Type",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = paste("Misaligned Value:", C_position_type),
    Impact_Scaling = total_fte) %>% 
  select(Flag_Category, Flag_Type, ERS_Coded_Field, Dollars_FTE_Flag, Flag_Description, Impact_Scaling)

#-----------------------------------------------------------------------
position_type_missing <- tibble(C_position_type = position_values) %>%
  left_join(final_coded_file %>%
              group_by(C_position_type) %>%
              summarise(total_fte = sum(D_FTE, na.rm = TRUE),
                        .groups = "drop"),
            by = "C_position_type") %>%
  mutate(total_fte = ifelse(is.na(total_fte), 0, total_fte)) %>% 
  filter(total_fte == 0) %>%
  mutate(
    Flag_Category = "Missing or Misaligned Values",
    Flag_Type = "Yellow",
    ERS_Coded_Field = "Position Type",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = paste("Missing Value:", C_position_type),
    Impact_Scaling = NA) %>% 
  select(Flag_Category, Flag_Type, ERS_Coded_Field, Dollars_FTE_Flag, Flag_Description, Impact_Scaling)

tier_1 <- bind_rows(tier_1, position_type_incorrect)
tier_1 <- bind_rows(tier_1, position_type_missing)


rm(position_type_incorrect, position_type_missing)


```

```{r}

# Function

function_values <- c("Teacher Compensation",
                     "Aides Compensation",
                     "Substitute Compensation",
                     "Library & Media",
                     "Instructional Materials & Supplies",
                     "Other Non-Compensation",
                     "Other Compensation",
                     "Extended Time & Tutoring",
                     "Enrichment",
                     "Social & Emotional",
                     "Physical Health & Services",
                     "Career & Academic Counseling",
                     "Parent & Community Relations",
                     "Professional Growth",
                     "Curriculum Development",
                     "Recruitment",
                     "Special Population Program Management & Support",
                     "Governance",
                     "School Supervision",
                     "School Administration",
                     "Research & Accountability",
                     "Communications",
                     "Student Assignment",
                     "Facilities & Maintenance",
                     "Security & Safety",
                     "Food Services",
                     "Student Transportation",
                     "Utilities",
                     "Human Resources",
                     "Finance, Budget, Purchasing & Distribution",
                     "Data Processing & Information Services",
                     "Facilities Planning",
                     "Development & Fundraising",
                     "Legal",
                     "Insurance",
                     "Untracked Budget Set-Aside")


function_incorrect <- final_coded_file %>%
  filter(!C_function %in% function_values) %>%
  group_by(C_function) %>% 
  summarize(total_fte = sum(D_FTE, na.rm = TRUE), .groups = "drop") %>% 
  mutate(
    Flag_Category = "Missing or Misaligned Values",
    Flag_Type = "Red",
    ERS_Coded_Field = "Function",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = paste("Misaligned Value:", C_function),
    Impact_Scaling = total_fte) %>% 
  select(Flag_Category, Flag_Type, ERS_Coded_Field, Dollars_FTE_Flag, Flag_Description, Impact_Scaling)

#-----------------------------------------------------------------------
function_missing <- tibble(C_function = function_values) %>%
  left_join(final_coded_file %>%
              group_by(C_function) %>%
              summarise(total_fte = sum(D_FTE, na.rm = TRUE),
                        .groups = "drop"),
            by = "C_function") %>%
  mutate(total_fte = ifelse(is.na(total_fte), 0, total_fte)) %>% 
  filter(total_fte == 0) %>%
  mutate(
    Flag_Category = "Missing or Misaligned Values",
    Flag_Type = "Yellow",
    ERS_Coded_Field = "Function",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = paste("Missing Value:", C_function),
    Impact_Scaling = NA) %>% 
  select(Flag_Category, Flag_Type, ERS_Coded_Field, Dollars_FTE_Flag, Flag_Description, Impact_Scaling)


tier_1 <- bind_rows(tier_1, function_incorrect)
tier_1 <- bind_rows(tier_1, function_missing)


rm(function_incorrect, function_missing)


```

```{r}

# use_values <- c("Instruction",
#                      "Pupil Services & Enrichment",
#                      "Instructional Support & Professional Growth",
#                      "Leadership",
#                      "Operations & Maintenance",
#                      "Business Services",
#                       "Untracked Budget Set-Aside")
# 
# 
# 
# use_incorrect <- final_coded_file %>%
#   filter(!C_use %in% use_values) %>%
#   group_by(C_use) %>% 
#   summarize(total_fte = sum(D_FTE, na.rm = TRUE), .groups = "drop") %>% 
#   mutate(
#     Flag_Category = "Missing or Misaligned Values",
#     Flag_Type = "Red",
#     ERS_Coded_Field = "Use",
#     Dollars_FTE_Flag = "FTE",
#     Flag_Description = paste("Misaligned Value:", C_use),
#     Impact_Scaling = total_fte) %>% 
#   select(Flag_Category, Flag_Type, ERS_Coded_Field, Dollars_FTE_Flag, Flag_Description, Impact_Scaling)
# 
# #-----------------------------------------------------------------------
# use_missing <- tibble(C_use = use_values) %>%
#   left_join(final_coded_file %>%
#               group_by(C_use) %>%
#               summarise(total_fte = sum(D_FTE, na.rm = TRUE),
#                         .groups = "drop"),
#             by = "C_use") %>%
#   mutate(total_fte = ifelse(is.na(total_fte), 0, total_fte)) %>% 
#   filter(total_fte == 0) %>%
#   mutate(
#     Flag_Category = "Missing or Misaligned Values",
#     Flag_Type = "Yellow",
#     ERS_Coded_Field = "Use",
#     Dollars_FTE_Flag = "FTE",
#     Flag_Description = paste("Missing Value:", C_use),
#     Impact_Scaling = NA) %>% 
#   select(Flag_Category, Flag_Type, ERS_Coded_Field, Dollars_FTE_Flag, Flag_Description, Impact_Scaling)
# 
# 
# tier_1 <- bind_rows(tier_1, use_incorrect)
# tier_1 <- bind_rows(tier_1, use_missing)
# 
# 
# rm(use_incorrect, use_missing)



```

```{r}

student_values <- c("Special Education",
                         "English Language Learner",
                         "Poverty",
                         "Alternative",
                         "Gifted",
                         "Unspecified")




student_type_incorrect <- final_coded_file %>%
  filter(!C_student_type %in% student_values) %>%
  group_by(C_student_type) %>% 
  summarize(total_fte = sum(D_FTE, na.rm = TRUE), .groups = "drop") %>% 
  mutate(
    Flag_Category = "Missing or Misaligned Values",
    Flag_Type = "Red",
    ERS_Coded_Field = "Student Type",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = paste("Misaligned Value:", C_student_type),
    Impact_Scaling = total_fte) %>% 
  select(Flag_Category, Flag_Type, ERS_Coded_Field, Dollars_FTE_Flag, Flag_Description, Impact_Scaling)

#-----------------------------------------------------------------------
student_type_missing <- tibble(C_student_type = student_values) %>%
  left_join(final_coded_file %>%
              group_by(C_student_type) %>%
              summarise(total_fte = sum(D_FTE, na.rm = TRUE),
                        .groups = "drop"),
            by = "C_student_type") %>%
  mutate(total_fte = ifelse(is.na(total_fte), 0, total_fte)) %>% 
  filter(total_fte == 0) %>%
  mutate(
    Flag_Category = "Missing or Misaligned Values",
    Flag_Type = "Yellow",
    ERS_Coded_Field = "Student Type",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = paste("Missing Value:", C_student_type),
    Impact_Scaling = NA) %>% 
  select(Flag_Category, Flag_Type, ERS_Coded_Field, Dollars_FTE_Flag, Flag_Description, Impact_Scaling)

tier_1 <- bind_rows(tier_1, student_type_incorrect)
tier_1 <- bind_rows(tier_1, student_type_missing)


rm(student_type_incorrect, student_type_missing)


```

```{r}

prek_values <- c("PreK",
                 "Non PreK")



prek_incorrect <- final_coded_file %>%
  filter(!C_prek %in% prek_values) %>%
  group_by(C_prek) %>% 
  summarize(total_fte = sum(D_FTE, na.rm = TRUE), .groups = "drop") %>% 
  mutate(
    Flag_Category = "Missing or Misaligned Values",
    Flag_Type = "Red",
    ERS_Coded_Field = "PreK",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = paste("Misaligned Value:", C_prek),
    Impact_Scaling = total_fte) %>% 
  select(Flag_Category, Flag_Type, ERS_Coded_Field, Dollars_FTE_Flag, Flag_Description, Impact_Scaling)

#-----------------------------------------------------------------------
prek_missing <- tibble(C_prek = prek_values) %>%
  left_join(final_coded_file %>%
              group_by(C_prek) %>%
              summarise(total_fte = sum(D_FTE, na.rm = TRUE),
                        .groups = "drop"),
            by = "C_prek") %>%
  mutate(total_fte = ifelse(is.na(total_fte), 0, total_fte)) %>% 
  filter(total_fte == 0) %>%
  mutate(
    Flag_Category = "Missing or Misaligned Values",
    Flag_Type = "Yellow",
    ERS_Coded_Field = "PreK",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = paste("Missing Value:", C_prek),
    Impact_Scaling = NA) %>% 
  select(Flag_Category, Flag_Type, ERS_Coded_Field, Dollars_FTE_Flag, Flag_Description, Impact_Scaling)


tier_1 <- bind_rows(tier_1, prek_incorrect)
tier_1 <- bind_rows(tier_1, prek_missing)


rm(prek_incorrect, prek_missing)



```

```{r}

sharing_values <- c("Leadership & Management",
                    "Shared Services",
                    "School on Central Budgets",
                    "School Reported")


sharing_incorrect <- final_coded_file %>%
  filter(!C_sharing %in% sharing_values) %>%
  group_by(C_sharing) %>% 
  summarize(total_fte = sum(D_FTE, na.rm = TRUE), .groups = "drop") %>% 
  mutate(
    Flag_Category = "Missing or Misaligned Values",
    Flag_Type = "Red",
    ERS_Coded_Field = "Sharing",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = paste("Misaligned Value:", C_sharing),
    Impact_Scaling = total_fte) %>% 
  select(Flag_Category, Flag_Type, ERS_Coded_Field, Dollars_FTE_Flag, Flag_Description, Impact_Scaling)

#-----------------------------------------------------------------------
sharing_missing <- tibble(C_sharing = sharing_values) %>%
  left_join(final_coded_file %>%
              group_by(C_sharing) %>%
              summarise(total_fte = sum(D_FTE, na.rm = TRUE),
                        .groups = "drop"),
            by = "C_sharing") %>%
  mutate(total_fte = ifelse(is.na(total_fte), 0, total_fte)) %>% 
  filter(total_fte == 0) %>%
  mutate(
    Flag_Category = "Missing or Misaligned Values",
    Flag_Type = "Yellow",
    ERS_Coded_Field = "Sharing",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = paste("Missing Value:", C_sharing),
    Impact_Scaling = NA) %>% 
  select(Flag_Category, Flag_Type, ERS_Coded_Field, Dollars_FTE_Flag, Flag_Description, Impact_Scaling)


tier_1 <- bind_rows(tier_1, sharing_incorrect)
tier_1 <- bind_rows(tier_1, sharing_missing)


rm(sharing_incorrect, sharing_missing)



```

```{r}
# 
# # Sample mapping: function -> use
# function_to_use <- c(
#   "Teacher Compensation" = "Instruction",
#   "Aides Compensation" = "Instruction",
#   "Substitute Compensation" = "Instruction",
#   "Library & Media" = "Instruction",
#   "Instructional Materials & Supplies" = "Instruction",
#   "Other Non-Compensation" = "Instruction",
#   "Other Compensation" = "Instruction",
#   "Extended Time & Tutoring" = "Instruction",
#   "Enrichment" = "Pupil Services & Enrichment",
#   "Social & Emotional" = "Pupil Services & Enrichment",
#   "Physical Health & Services" = "Pupil Services & Enrichment",
#   "Career & Academic Counseling" = "Pupil Services & Enrichment",
#   "Parent & Community Relations" = "Pupil Services & Enrichment",
#   "Professional Growth" = "Instructional Support & Professional Growth",
#   "Curriculum Developement" = "Instructional Support & Professional Growth",
#   "Recruitment" = "Instructional Support & Professional Growth",
#   "Special Population Program Management & Support" = "Instructional Support & Professional Growth",
#   "Governance" = "Leadership",
#   "School Supervision" = "Leadership",
#   "School Administration" = "Leadership",
#   "Research & Accountability" = "Leadership",
#   "Communications" = "Leadership",
#   "Student Assignment" = "Leadership",
#   "Facilities & Maintenance" = "Operations & Maintenance",
#   "Security & Safety" = "Operations & Maintenance",
#   "Food Services" = "Operations & Maintenance",
#   "Student Transportation" = "Operations & Maintenance",
#   "Utilities" = "Operations & Maintenance",
#   "Human Resources" = "Business Services",
#   "Finance, Budget, Purchasing & Distribution" = "Business Services",
#   "Data Processing & Information Services" = "Business Services",
#   "Facilities Planning" = "Business Services",
#   "Development & Fundraising" = "Business Services",
#   "Legal" = "Business Services",
#   "Insurance" = "Business Services")
# 
# 
# 
# # Add expected mapping + check
# final_coded_file$expected_use <- function_to_use[final_coded_file$C_function]
# final_coded_file$use_match <- final_coded_file$C_use == final_coded_file$expected_use
# 
# # Filter mismatches
# mismatches <- final_coded_file %>%
#   filter(!use_match)
# 
# # Summarize by function-use combo
# function_mismatches <- mismatches %>%
#   group_by(C_function, C_use, expected_use) %>%
#   summarise(total_fte = sum(D_FTE, na.rm = TRUE), .groups = "drop") %>%
#   mutate(
#     Flag_Category   = "Missing or Misaligned Values",
#     Flag_Type       = "Red",
#     ERS_Coded_Field = "Function",
#     Dollars_FTE_Flag = "FTE",
#     Flag_Description = paste0(C_function, 
#                               " coded as ", C_use, 
#                               " but expected ", expected_use),
#     Impact_Scaling  = total_fte
#   ) %>%
#   select(Flag_Category, Flag_Type, ERS_Coded_Field, 
#          Dollars_FTE_Flag, Flag_Description, Impact_Scaling)
# 
# 
# tier_1 <- rbind(tier_1, function_mismatches)
# rm(mismatches, function_mismatches)

```

```{r}
# Filtering out expected mismatches/devations (i.e no FTE in non-personnel functions)

tier_1 <- tier_1 %>%
  filter(
    !(Dollars_FTE_Flag == "FTE" & 
        str_detect(Flag_Description, 
                   regex("Non-Position|Utilities|Other Non-Compensation|Benefit|Contracted Services|Supplies/Materials|Equipment & Equipment Lease|Travel & Conferences|Instructional Materials & Supplies|Untracked Budget Set-Aside", ignore_case = TRUE)))
  )


```

## Step 2.2 Checking for internal consistency. Do the position and function values make sense in combination?

```{r}

tier_2 <- tibble(
  Flag_Category = character(),
  Flag_Type = character(),
  ERS_Coded_Field = character(),
  Dollars_FTE_Flag = character(),
  Flag_Description = character(),
  Impact_Scaling = numeric()
) 


```

```{r}
## Position to Function

# * Note that sometimes Substitutes can be coded to Professional Growth
position_to_function <- c(
  "Teacher" = "Teacher Compensation",
  "TA" = "Aides Compensation",
  "Custodian" = "Facilities & Maintenance",
  "Principal" = "School Administration",
  "Vice Principal" = "School Administration",
  "Substitute" = "Substitute Compensation",
  "Librarian" = "Library & Media",
  "Nurse" = "Physical Health & Services",
  "Nurse Aide" = "Physical Health & Services",
  "Physical Therapist" = "Physical Health & Services",
  "Occupational Therapist" = "Physical Health & Services",
  "Speech Therapist" = "Physical Health & Services",
  "School Monitor/Security" = "Security & Safety",
  "School Resource Officer/Police" = "Security & Safety"
)

# Check for mismatches
final_coded_file$expected_function <- position_to_function[final_coded_file$C_position_type]
final_coded_file$function_match <- final_coded_file$C_function == final_coded_file$expected_function

# Filter rows where the use does NOT match the expected one
mismatches <- final_coded_file[!final_coded_file$function_match, ]


# Summarize by function-use combo
position_mismatches <- mismatches %>%
  group_by(C_position_type, C_function, expected_function) %>%
  summarise(total_fte = sum(D_FTE, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    Flag_Category   = "Internal Consistency",
    Flag_Type       = "Red",
    ERS_Coded_Field = "Position Type & Function",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = paste0(C_position_type, 
                              " coded as ", C_function, 
                              " but expected ", expected_function),
    Impact_Scaling  = total_fte
  ) %>%
  select(Flag_Category, Flag_Type, ERS_Coded_Field, 
         Dollars_FTE_Flag, Flag_Description, Impact_Scaling)


tier_2 <- rbind(tier_2, position_mismatches)
rm(mismatches, position_mismatches)



```

```{r}

# Function to Sharing

# Define the rules you want to flag
rules <- tibble::tribble(
  ~C_function,              ~sharing_match,              ~Flag_Type,
  "Teacher Compensation",     "Leadership & Management", "Red",
  "Aides Compensation","Leadership & Management",          "Red",
  "Substitute Compensation","Leadership & Management","Red",
  "Library & Media",     "Leadership & Management", "Red",
  "Other Non-Compensation",     "Leadership & Management", "Red",
  "Other Compensation",     "Leadership & Management", "Red",
  "Instructional Materials & Supplies",     "Leadership & Management", "Red",
  "Extended Time & Tutoring",     "Leadership & Management", "Red",
  "Teacher Compensation",     "Shared Services", "Yellow",
  "Aides Compensation","Shared Services",          "Yellow",
  "Substitute Compensation","Shared Services","Yellow",
  "Library & Media",     "Shared Services", "Yellow",
  "Other Non-Compensation",     "Shared Services", "Yellow",
  "Other Compensation",     "Shared Services", "Yellow",
  "Instructional Materials & Supplies",     "Shared Services", "Yellow",
  "Extended Time & Tutoring",     "Shared Services", "Yellow",
  "School Administration",     "Leadership & Management", "Red",
  "School Administration",     "Shared Services", "Yellow",
  "School Supervision",     "School Reported", "Red",
  "School Supervision",     "School on Central Budgets", "Yellow",
)


# Apply rules in one step
flags_out <- final_coded_file %>%
  inner_join(rules, by = "C_function", relationship = "many-to-many") %>%
  filter(str_detect(C_sharing, sharing_match)) %>%
  group_by(C_function, sharing_match, Flag_Type) %>%
  summarise(total_fte = sum(D_FTE, na.rm = TRUE),
            .groups = "drop") %>%
  mutate(
    Flag_Category    = "Internal Consistency",
    Flag_Type = Flag_Type,
    ERS_Coded_Field  = "Function & Sharing",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = paste0(C_function, 
                              " appears with ", sharing_match),
    Impact_Scaling   = total_fte
  ) %>%
  select(Flag_Category, Flag_Type, ERS_Coded_Field,
         Dollars_FTE_Flag, Flag_Description, Impact_Scaling)


tier_2 <- rbind(tier_2, flags_out)
rm(rules, flags_out)


```

```{r}

## Majority of funds are L&M


# Define the target functions you want to check
target_functions <- c("Governance", "Student Assignment", "Curriculum Development", "Special Population Program Management & Support", "Research & Accountability", "Human Resources", "Finance, Budget, Purchasing & Distribution", "Data Processing & Information Services", "Facilities Planning", "Development & Fundraising", "Legal", "Insurance")

# Step 1: Filter data to just those functions
function_sharing_summary <- final_coded_file %>%
  filter(C_function %in% target_functions) %>%
  group_by(C_function, C_sharing) %>%
  summarise(total_fte = sum(D_FTE, na.rm = TRUE), .groups = "drop")

# Step 2: Calculate L&M share of total function spending
lm_share_summary <- function_sharing_summary %>%
  group_by(C_function) %>%
  mutate(total_function_spending = sum(total_fte)) %>%
  filter(C_sharing == "Leadership & Management") %>%
  mutate(lm_share = total_fte / total_function_spending) %>%
  ungroup()

# Step 3: Flag functions where L&M is less than 50%
red_flags <- lm_share_summary %>%
  filter(is.na(lm_share) | lm_share < 0.5) %>%
  mutate(
    Flag_Category   = "Internal Consistency",
    Flag_Type       = "Red",
    ERS_Coded_Field = "Function & Sharing",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = paste0(C_function, 
                              " has less than 50% FTE in L&M"),
    Impact_Scaling  = total_function_spending
  ) %>%
  select(Flag_Category, Flag_Type, ERS_Coded_Field, 
         Dollars_FTE_Flag, Flag_Description, Impact_Scaling)


tier_2 <- rbind(tier_2, red_flags)
rm(red_flags, function_sharing_summary, lm_share_summary)


```

```{r}

# Position Type to Sharing


# Define the rules you want to flag
rules_pos <- tibble::tribble(
  ~C_position_type,              ~sharing_match,              ~Flag_Type,
  "Principal",     "Leadership & Management", "Red",
  "Vice Principal","Leadership & Management",          "Red",
  "Teacher","Leadership & Management","Red",
  "TA",     "Leadership & Management", "Red",
  "Substitute",     "Leadership & Management", "Red",
  "Instructional Coach",     "Leadership & Management", "Red",
  "Guidance Counselor",     "Leadership & Management", "Red",
  "Social Worker",     "Leadership & Management", "Red",
  "Psychologist",     "Leadership & Management", "Red",
  "Physical Therapist",     "Leadership & Management", "Red",
  "Speech Therapist",     "Leadership & Management", "Red",
  "Occupational Therapist",     "Leadership & Management", "Red",
  "Principal",     "Shared Services", "Red",
  "Vice Principal",     "Shared Services", "Red",
  "Teacher",     "Shared Services", "Red",
  "TA",     "Shared Services", "Red",
  "Substitute",     "Shared Services", "Red",
  "(Exec) Director",     "School Reported", "Red",
  "Area Officer",     "School Reported", "Red",
  "(Exec) Director",     "School on Central Budgets", "Yellow",
  "(Exec) Director",     "Shared Services", "Yellow",
  "Principal",     "School on Central Budgets", "Yellow",
  "Vice Principal",     "School on Central Budgets", "Yellow",
  "Guidance Counselor",     "Shared Services", "Yellow",
)


# Apply rules in one step
flags_out_pos <- final_coded_file %>%
  inner_join(rules_pos, by = "C_position_type", relationship = "many-to-many") %>%
  filter(str_detect(C_sharing, sharing_match)) %>%
  group_by(C_position_type, sharing_match, Flag_Type) %>%
  summarise(total_fte = sum(D_FTE, na.rm = TRUE),
            .groups = "drop") %>%
  mutate(
    Flag_Category    = "Internal Consistency",
    Flag_Type = Flag_Type,
    ERS_Coded_Field  = "Position Type & Sharing",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = paste0(C_position_type, 
                              " appears with ", sharing_match),
    Impact_Scaling   = total_fte
  ) %>%
  select(Flag_Category, Flag_Type, ERS_Coded_Field,
         Dollars_FTE_Flag, Flag_Description, Impact_Scaling)


tier_2 <- rbind(tier_2, flags_out_pos)
rm(flags_out_pos, rules_pos)
```

```{r}
## Student Type to Sharing


# Define student types and sharing categories
target_student_types <- c("English Language Learner", "Special Education")
required_sharing <- c("Leadership & Management", "School Reported")

student_type_flags <- tibble()  # start empty

for (stype in target_student_types) {
  for (share in required_sharing) {
    total_fte <- sum(final_coded_file$D_FTE[
      final_coded_file$C_student_type == stype &
      final_coded_file$C_sharing == share
    ], na.rm = TRUE)

    if (total_fte == 0) {
      student_type_flags <- bind_rows(student_type_flags, tibble(
        Flag_Category   = "Internal Consistency",
        Flag_Type       = "Red",
        ERS_Coded_Field = "Student Type & Sharing",
        Dollars_FTE_Flag = "FTE",
        Flag_Description = paste0(stype,
                                  " has no FTE in ", share),
        Impact_Scaling  = NA
      ))
    }
  }
}


tier_2 <- rbind(tier_2, student_type_flags)
rm(student_type_flags)



```

```{r}

# Filtering out expected mismatches/devations (i.e row mismatches that don't bear any FTE - they likely are just dollar impacts)

tier_2 <- tier_2 %>%
  filter(Dollars_FTE_Flag == "FTE" & Impact_Scaling > 0)
        

```

## Step 2.3 Does the breakdown of spend generally align with average breakdowns? Any large fluctutations should be investigated

```{r}
tier_3 <- tibble(
  Flag_Category = character(),
  Flag_Type = character(),
  ERS_Coded_Field = character(),
  Dollars_FTE_Flag = character(),
  Flag_Description = character(),
  Impact_Scaling = numeric()
) 


```

### Step 2.3.1 Breakdown of FTE according to % Distribution

```{r}

# Percent Breakdown by FTE

expected_reporting <- c(
  "School" = 0.80,
  "Non-School" = 0.20)

reporting_summary <- final_coded_file %>%
  group_by(C_reporting) %>%
  summarise(total = sum(D_FTE, na.rm = TRUE), .groups = "drop") %>%
  mutate(actual_percent = total / sum(total),
         expected_percent = expected_reporting[C_reporting],
         diff = actual_percent - expected_percent)
         

reporting_flags <- reporting_summary %>%
  filter(is.na(expected_percent) | abs(diff) >= 0.05) %>%   # flag only bad ones
  mutate(
    Flag_Category   = "Distribution Deviations",
    Flag_Type       = "Yellow",
    ERS_Coded_Field = "Reporting",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = case_when(
    is.na(expected_percent) ~ paste0("Unexpected: ", C_reporting),
    TRUE ~ paste0(C_reporting, ": ", 
                scales::percent(actual_percent), 
                "; expected: ", 
                scales::percent(expected_percent))
    ),
    Impact_Scaling  = total
  ) %>%
  select(Flag_Category, Flag_Type, ERS_Coded_Field,
         Dollars_FTE_Flag, Flag_Description, Impact_Scaling)

# Append into tier_3
tier_3 <- bind_rows(tier_3, reporting_flags)
rm(reporting_summary, reporting_flags)


```

Note: Skip Object Type for FTE Only Checks

```{r}


expected_position_type <- c(
  "Area Officers" = 0.0,
  "Club Advisor/Coach" = 0.0,
  "Coordinator/Manager" = 0.03,
  "Custodian" = 0.05,
  "(Exec) Director" = 0.01,
  "Guidance Counselor" = 0.02,
  "Instructional Coach" = 0.02,
  "Instructional Coach Supervisors" = 0.0,
  "Librarian" = 0.01,
  "Non-Position" = 0.0,
  "Nurse" = 0.01,
  "Nurse Aide" = 0.0,
  "Occupational Therapist" = 0.0,
  "Other" = 0.14,
  "Physical Therapist" = 0.0,
  "Principal" = 0.01,
  "Psychologist" = 0.01,
  "Sec/Clerk/Other Admin" = 0.05,
  "School Monitor/Security" = 0.01,
  "School Resource Officer/Police" = 0.0,
  "Social Worker" = 0.01,
  "Speech Therapist" = 0.0,
  "Substitute" = 0.01,
  "TA" = 0.11,
  "Teacher" = 0.48,
  "Vice Principal" = 0.02)

position_type_summary <- final_coded_file %>%
  group_by(C_position_type) %>%
  summarise(total = sum(D_FTE, na.rm = TRUE), .groups = "drop") %>%
  mutate(actual_percent = total / sum(total),
         expected_percent = expected_position_type[C_position_type],
         diff = actual_percent - expected_percent)
         

position_type_flags <- position_type_summary %>%
  filter(is.na(expected_percent) | abs(diff) >= 0.05) %>%   # flag only bad ones
  mutate(
    Flag_Category   = "Distribution Deviations",
    Flag_Type       = "Yellow",
    ERS_Coded_Field = "Position Type",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = case_when(
    is.na(expected_percent) ~ paste0("Unexpected: ", C_position_type),
    TRUE ~ paste0(C_position_type, ": ", 
                scales::percent(actual_percent), 
                "; expected: ", 
                scales::percent(expected_percent))
    ),
    Impact_Scaling  = total
  ) %>%
  select(Flag_Category, Flag_Type, ERS_Coded_Field,
         Dollars_FTE_Flag, Flag_Description, Impact_Scaling)

# Append into tier_3
tier_3 <- bind_rows(tier_3, position_type_flags)
rm(position_type_summary, position_type_flags)


```

```{r}

expected_function <- c(
  "Teacher Compensation" = 0.48,
  "Aides Compensation" = 0.11,
  "Substitute Compensation" = 0.01,
  "Library & Media" = 0.01,
  "Instructional Materials & Supplies" = 0.0,
  "Other Non-Compensation" = 0.0,
  "Other Compensation" = 0.0,
  "Extended Time & Tutoring" = 0.0,
  "Enrichment" = 0.0,
  "Social & Emotional" = 0.03,
  "Physical Health & Services" = 0.02,
  "Career & Academic Counseling" = 0.02,
  "Parent & Community Relations" = 0.01,
  "Professional Growth" = 0.02,
  "Curriculum Development" = 0.0,
  "Recruitment" = 0.0,
  "Special Population Program Management & Support" = 0.02,
  "Governance" = 0.01,
  "School Supervision" = 0.0,
  "School Administration" = 0.06,
  "Research & Accountability" = 0.0,
  "Communications" = 0.0,
  "Student Assignment" = 0.0,
  "Facilities & Maintenance" = 0.07,
  "Security & Safety" = 0.01,
  "Food Services" = 0.05,
  "Student Transportation" = 0.04,
  "Utilities" = 0.0,
  "Human Resources" = 0.01,
  "Finance, Budget, Purchasing & Distribution" = 0.01,
  "Data Processing & Information Services" = 0.01,
  "Facilities Planning" = 0.0,
  "Development & Fundraising" = 0.0,
  "Legal" = 0.0,
  "Insurance" = 0.0,
  "Untracked Budget Set-Aside" = 0.0)


function_summary <- final_coded_file %>%
  group_by(C_function) %>%
  summarise(total = sum(D_FTE, na.rm = TRUE), .groups = "drop") %>%
  mutate(actual_percent = total / sum(total),
         expected_percent = expected_function[C_function],
         diff = actual_percent - expected_percent)
         

function_flags <- function_summary %>%
  filter(is.na(expected_percent) | abs(diff) >= 0.05) %>%   # flag only bad ones
  mutate(
    Flag_Category   = "Distribution Deviations",
    Flag_Type       = "Yellow",
    ERS_Coded_Field = "Function",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = case_when(
    is.na(expected_percent) ~ paste0("Unexpected: ", C_function),
    TRUE ~ paste0(C_function, ": ", 
                scales::percent(actual_percent), 
                "; expected: ", 
                scales::percent(expected_percent))
    ),
    Impact_Scaling  = total
  ) %>%
  select(Flag_Category, Flag_Type, ERS_Coded_Field,
         Dollars_FTE_Flag, Flag_Description, Impact_Scaling)

# Append into tier_3
tier_3 <- bind_rows(tier_3, function_flags)
rm(function_summary, function_flags)



```

```{r}

expected_student_type <- c(
  "Special Education" = 0.19,
  "English Language Learner" = 0.03,
  "Poverty" = 0.03,
  "Alternative" = 0.01,
  "Gifted" = 0.0,
  "Unspecified" = 0.74)


student_type_summary <- final_coded_file %>%
  group_by(C_student_type) %>%
  summarise(total = sum(D_FTE, na.rm = TRUE), .groups = "drop") %>%
  mutate(actual_percent = total / sum(total),
         expected_percent = expected_student_type[C_student_type],
         diff = actual_percent - expected_percent)
         

student_type_flags <- student_type_summary %>%
  filter(is.na(expected_percent) | abs(diff) >= 0.05) %>%   # flag only bad ones
  mutate(
    Flag_Category   = "Distribution Deviations",
    Flag_Type       = "Yellow",
    ERS_Coded_Field = "Student Type",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = case_when(
    is.na(expected_percent) ~ paste0("Unexpected: ", C_student_type),
    TRUE ~ paste0(C_student_type, ": ", 
                scales::percent(actual_percent), 
                "; expected: ", 
                scales::percent(expected_percent))
    ),
    Impact_Scaling  = total
  ) %>%
  select(Flag_Category, Flag_Type, ERS_Coded_Field,
         Dollars_FTE_Flag, Flag_Description, Impact_Scaling)

# Append into tier_3
tier_3 <- bind_rows(tier_3, student_type_flags)
rm(student_type_summary, student_type_flags)


```

```{r}


expected_prek <- c(
  "Non PreK" = 0.94,
  "PreK" = 0.06)

prek_summary <- final_coded_file %>%
  group_by(C_prek) %>%
  summarise(total = sum(D_FTE, na.rm = TRUE), .groups = "drop") %>%
  mutate(actual_percent = total / sum(total),
         expected_percent = expected_prek[C_prek],
         diff = actual_percent - expected_percent)
         

prek_flags <- prek_summary %>%
  filter(is.na(expected_percent) | abs(diff) >= 0.05) %>%   # flag only bad ones
  mutate(
    Flag_Category   = "Distribution Deviations",
    Flag_Type       = "Yellow",
    ERS_Coded_Field = "PreK",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = case_when(
    is.na(expected_percent) ~ paste0("Unexpected: ", C_prek),
    TRUE ~ paste0(C_prek, ": ", 
                scales::percent(actual_percent), 
                "; expected: ", 
                scales::percent(expected_percent))
    ),
    Impact_Scaling  = total
  ) %>%
  select(Flag_Category, Flag_Type, ERS_Coded_Field,
         Dollars_FTE_Flag, Flag_Description, Impact_Scaling)

# Append into tier_3
tier_3 <- bind_rows(tier_3, prek_flags)
rm(prek_summary, prek_flags)


```

```{r}

expected_sharing <- c(
  "Leadership & Management" = 0.06,
  "School on Central Budgets" = 0.08,
  "Shared Services" = 0.12,
  "School Reported" = 0.74)


sharing_summary <- final_coded_file %>%
  group_by(C_sharing) %>%
  summarise(total = sum(D_FTE, na.rm = TRUE), .groups = "drop") %>%
  mutate(actual_percent = total / sum(total),
         expected_percent = expected_sharing[C_sharing],
         diff = actual_percent - expected_percent)
         

sharing_flags <- sharing_summary %>%
  filter(is.na(expected_percent) | abs(diff) >= 0.05) %>% #flag only bad ones
  mutate(
    Flag_Category   = "Distribution Deviations",
    Flag_Type       = "Yellow",
    ERS_Coded_Field = "Sharing",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = case_when(
    is.na(expected_percent) ~ paste0("Unexpected: ", C_sharing),
    TRUE ~ paste0(C_sharing, ": ", 
                scales::percent(actual_percent), 
                "; expected: ", 
                scales::percent(expected_percent))
    ),
    Impact_Scaling  = total
  ) %>%
  select(Flag_Category, Flag_Type, ERS_Coded_Field,
         Dollars_FTE_Flag, Flag_Description, Impact_Scaling)

# Append into tier_3
tier_3 <- bind_rows(tier_3, sharing_flags)
rm(sharing_summary, sharing_flags)


```

```{r}
tier_4 <- tibble(
  Flag_Category = character(),
  Flag_Type = character(),
  ERS_Coded_Field = character(),
  Dollars_FTE_Flag = character(),
  Flag_Description = character(),
  Impact_Scaling = numeric()
) 


```

### Step 2.3.2 Breakdown of FTE according to expected FTE/500 values

```{r}

# Reporting by FTE/500

expected_reporting <- c(
  "School" = 54.0,
  "Non-School" = 15.0)

reporting_summary <- final_coded_file %>%
  group_by(C_reporting) %>%
  summarise(total = sum(M_FTE_500, na.rm = TRUE), .groups = "drop") %>%
  mutate(percent_diff = (total - expected_reporting[C_reporting]) / expected_reporting[C_reporting])
         

reporting_flags <- reporting_summary %>%
  filter(is.na(percent_diff) | abs(percent_diff) >= 0.5) %>%
  filter((total - expected_reporting[C_reporting]) > 1) %>%# flag only bad ones
  mutate(
    Flag_Category   = "Distribution Deviations",
    Flag_Type       = "Yellow",
    ERS_Coded_Field = "Reporting",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = case_when(
    is.na(percent_diff) ~ paste0("Unexpected: ", C_reporting),
    TRUE ~ paste0(C_reporting, " FTE/500: ", 
                round(total, 1), 
                "; expected: ", 
                expected_reporting[C_reporting])
    ),
    Impact_Scaling  = total
  ) %>%
  select(Flag_Category, Flag_Type, ERS_Coded_Field,
         Dollars_FTE_Flag, Flag_Description, Impact_Scaling)

# Append into tier_3
tier_4 <- bind_rows(tier_4, reporting_flags)
rm(reporting_summary, reporting_flags)


```

Note: Exclude Object Type for FTE only Checks

```{r}


expected_position_type <- c(
  "Area Officers" = 0.07,
  "Club Advisor/Coach" = 0.13,
  "Coordinator/Manager" = 1.87,
  "Custodian" = 3.59,
  "(Exec) Director" = 0.37,
  "Guidance Counselor" = 1.12,
  "Instructional Coach" = 1.23,
  "Instructional Coach Supervisors" = 0.02,
  "Librarian" = 0.70,
  "Non-Position" = 0.0,
  "Nurse" = 0.60,
  "Nurse Aide" = 0.18,
  "Occupational Therapist" = 0.17,
  "Other" = 9.67,
  "Physical Therapist" = 0.07,
  "Principal" = 0.86,
  "Psychologist" = 0.51,
  "Sec/Clerk/Other Admin" = 3.19,
  "School Monitor/Security" = 1.10,
  "School Resource Officer/Police" = 1.06,
  "Social Worker" = 0.57,
  "Speech Therapist" = 0.62,
  "Substitute" = 0.17,
  "TA" = 7.73,
  "Teacher" = 33.18,
  "Vice Principal" = 1.06)


position_type_summary <- final_coded_file %>%
  group_by(C_position_type) %>%
  summarise(total = sum(M_FTE_500, na.rm = TRUE), .groups = "drop") %>%
  mutate(percent_diff = (total - expected_position_type[C_position_type]) / expected_position_type[C_position_type])
         

position_type_flags <- position_type_summary %>%
  filter(is.na(percent_diff) | abs(percent_diff) >= 0.5) %>%
  filter((total - expected_position_type[C_position_type]) > 1) %>% # flag only bad ones
  mutate(
    Flag_Category   = "Distribution Deviations",
    Flag_Type       = "Yellow",
    ERS_Coded_Field = "Position Type",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = case_when(
    is.na(percent_diff) ~ paste0("Unexpected: ", C_position_type),
    TRUE ~ paste0(C_position_type, " FTE/500: ", 
                round(total, 1), 
                "; expected: ", 
                expected_position_type[C_position_type])
    ),
    Impact_Scaling  = total
  ) %>%
  select(Flag_Category, Flag_Type, ERS_Coded_Field,
         Dollars_FTE_Flag, Flag_Description, Impact_Scaling)


# Append into tier_3
tier_4 <- bind_rows(tier_4, position_type_flags)
rm(position_type_summary, position_type_flags)


```

```{r}

expected_function <- c(
  "Teacher Compensation" = 33.00,
  "Aides Compensation" = 7.75,
  "Substitute Compensation" = 0.15,
  "Library & Media" = 0.80,
  "Instructional Materials & Supplies" = 0.0,
  "Other Non-Compensation" = 0.0,
  "Other Compensation" = 0.13,
  "Extended Time & Tutoring" = 0.10,
  "Enrichment" = 0.31,
  "Social & Emotional" = 1.67,
  "Physical Health & Services" = 1.60,
  "Career & Academic Counseling" = 1.12,
  "Parent & Community Relations" = 0.51,
  "Professional Growth" = 1.45,
  "Curriculum Development" = 0.29,
  "Recruitment" = 0.03,
  "Special Population Program Management & Support" = 1.01,
  "Governance" = 0.30,
  "School Supervision" = 0.19,
  "School Administration" = 4.36,
  "Research & Accountability" = 0.03,
  "Communications" = 0.10,
  "Student Assignment" = 0.16,
  "Facilities & Maintenance" = 4.83,
  "Security & Safety" = 1.21,
  "Food Services" = 2.84,
  "Student Transportation" = 2.92,
  "Utilities" = 0.0,
  "Human Resources" = 0.40,
  "Finance, Budget, Purchasing & Distribution" = 0.66,
  "Data Processing & Information Services" = 0.58,
  "Facilities Planning" = 0.03,
  "Development & Fundraising" = 0.05,
  "Legal" = 0.07,
  "Insurance" = 0.0,
  "Untracked Budget Set-Aside" = 0.0)



function_summary <- final_coded_file %>%
  group_by(C_function) %>%
  summarise(total = sum(M_FTE_500, na.rm = TRUE), .groups = "drop") %>%
  mutate(percent_diff = (total - expected_function[C_function]) / expected_function[C_function])
         

function_flags <- function_summary %>%
  filter(is.na(percent_diff) | abs(percent_diff) >= 0.5) %>%
  filter((total - expected_function[C_function]) > 1) %>%# flag only bad ones
  mutate(
    Flag_Category   = "Distribution Deviations",
    Flag_Type       = "Yellow",
    ERS_Coded_Field = "Function",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = case_when(
    is.na(percent_diff) ~ paste0("Unexpected: ", C_function),
    TRUE ~ paste0(C_function, " FTE/500: ", 
                round(total, 1), 
                "; expected: ", 
                expected_function[C_function])
    ),
    Impact_Scaling  = total
  ) %>%
  select(Flag_Category, Flag_Type, ERS_Coded_Field,
         Dollars_FTE_Flag, Flag_Description, Impact_Scaling)


# Append into tier_3
tier_4 <- bind_rows(tier_4, function_flags)
rm(function_summary, function_flags)



```

```{r}

expected_student_type <- c(
  "Special Education" = 12.92,
  "English Language Learner" = 2.06,
  "Poverty" = 2.29,
  "Alternative" = 0.56,
  "Gifted" = 0.38,
  "Unspecified" = 49.75)



student_type_summary <- final_coded_file %>%
  group_by(C_student_type) %>%
  summarise(total = sum(M_FTE_500, na.rm = TRUE), .groups = "drop") %>%
  mutate(percent_diff = (total - expected_student_type[C_student_type]) / expected_student_type[C_student_type])
         

student_type_flags <- student_type_summary %>%
  filter(is.na(percent_diff) | abs(percent_diff) >= 0.5) %>%   # flag only bad ones
  filter((total - expected_student_type[C_student_type]) > 1) %>%
  mutate(
    Flag_Category   = "Distribution Deviations",
    Flag_Type       = "Yellow",
    ERS_Coded_Field = "Student Type",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = case_when(
    is.na(percent_diff) ~ paste0("Unexpected: ", C_student_type),
    TRUE ~ paste0(C_student_type, " FTE/500: ", 
                round(total, 1), 
                "; expected: ", 
                expected_student_type[C_student_type])
    ),
    Impact_Scaling  = total
  ) %>%
  select(Flag_Category, Flag_Type, ERS_Coded_Field,
         Dollars_FTE_Flag, Flag_Description, Impact_Scaling)


# Append into tier_3
tier_4 <- bind_rows(tier_4, student_type_flags)
rm(student_type_summary, student_type_flags)


```

```{r}


expected_prek <- c(
  "Non PreK" = 64.62,
  "PreK" = 2.40)

prek_summary <- final_coded_file %>%
  group_by(C_prek) %>%
  summarise(total = sum(M_FTE_500, na.rm = TRUE), .groups = "drop") %>%
  mutate(percent_diff = (total - expected_prek[C_prek]) / expected_prek[C_prek])
         

prek_flags <- prek_summary %>%
  filter(is.na(percent_diff) | abs(percent_diff) >= 0.5) %>%
  filter((total - expected_prek[C_prek]) > 1) %>% # flag only bad ones
  mutate(
    Flag_Category   = "Distribution Deviations",
    Flag_Type       = "Yellow",
    ERS_Coded_Field = "PreK",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = case_when(
    is.na(percent_diff) ~ paste0("Unexpected: ", C_prek),
    TRUE ~ paste0(C_prek, " FTE/500: ", 
                round(total, 1), 
                "; expected: ", 
                expected_prek[C_prek])
    ),
    Impact_Scaling  = total
  ) %>%
  select(Flag_Category, Flag_Type, ERS_Coded_Field,
         Dollars_FTE_Flag, Flag_Description, Impact_Scaling)


# Append into tier_3
tier_4 <- bind_rows(tier_4, prek_flags)
rm(prek_summary, prek_flags)


```

```{r}

expected_sharing <- c(
  "Leadership & Management" = 3.65,
  "School on Central Budgets" = 5.25,
  "Shared Services" = 7.57,
  "School Reported" = 51.52)


sharing_summary <- final_coded_file %>%
  group_by(C_sharing) %>%
  summarise(total = sum(M_FTE_500, na.rm = TRUE), .groups = "drop") %>%
  mutate(percent_diff = (total - expected_sharing[C_sharing]) / expected_sharing[C_sharing])
         

sharing_flags <- sharing_summary %>%
  filter(is.na(percent_diff) | abs(percent_diff) >= 0.5) %>% 
  filter((total - expected_sharing[C_sharing]) > 1) %>% # flag only bad ones
  mutate(
    Flag_Category   = "Distribution Deviations",
    Flag_Type       = "Yellow",
    ERS_Coded_Field = "Sharing",
    Dollars_FTE_Flag = "FTE",
    Flag_Description = case_when(
    is.na(percent_diff) ~ paste0("Unexpected: ", C_sharing),
    TRUE ~ paste0(C_sharing, " FTE/500: ", 
                round(total, 1), 
                "; expected: ", 
                expected_sharing[C_sharing])
    ),
    Impact_Scaling  = total
  ) %>%
  select(Flag_Category, Flag_Type, ERS_Coded_Field,
         Dollars_FTE_Flag, Flag_Description, Impact_Scaling)


# Append into tier_3
tier_4 <- bind_rows(tier_4, sharing_flags)
rm(sharing_summary, sharing_flags)


```

```{r}

tier_4 <- tier_4 %>%
  filter(
    !(Dollars_FTE_Flag == "FTE" & 
        str_detect(Flag_Description, 
                   regex("Non-Position|Utilities|Other Non-Compensation|Benefit|Contracted Services|Supplies/Materials|Equipment & Equipment Lease|Travel & Conferences|Instructional Materials & Supplies|Untracked Budget Set-Aside|Insurance", ignore_case = TRUE)))
  )



```

```{r}

# Exporting the file

final_checks <- rbind(tier_1, tier_2, tier_3, tier_4)

processed_data_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2025%20Cohort/1.%20Data%20&%20Analysis%20-%20Secure/3.%20Data%20Vetting"

#budgeted positions
ers_write_sharepoint(data = final_checks,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = paste0(district_folder,"_coding_vetting.csv"))


```