---
title: "03_Stack the Files"
author: "Jenny Katz"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3   # Customize the depth of the table of contents for HTML document
    toc_float:
      collapsed: true    # Start with the table of contents expanded
      smooth_scroll: true  # Enable smooth scrolling
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "knitted files", output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 0: Read Libraries and connect to Sharepoint

```{r}
erstools::ers_load_packages()
options(scipen = 99) #prevent any numbers from showing up as scientific notation
```

```{r}
#set sharepoint site URL for access to our org-wide files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"

#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)
```

```{r}
ers_read_sharepoint_local <- function(folder_path,
                                file_name_with_extension,
                                sheet_name = NULL,
                                drive_name = "client_work_drive",
                                skip_rows = 0,
                                na_values = c(""),
                                clean_names = "FALSE")
{
  # Replace '%20' in the file name with spaces to clean the file name
  file_name_with_extension <- gsub("%20", " ", file_name_with_extension)

  # Clean the folder path and extract the drive name from it if not explicitly provided
  cleaned <- erstools::ers_sharepoint_path_clean(folder_path)
  folder_path <- cleaned$path_clean  # Cleaned folder path for SharePoint

  # If drive_name is not explicitly provided, use the drive_name extracted from the folder path
  if (missing(drive_name)) {
    drive_name <- cleaned$drive_name
  }

  # Suppress messages while connecting to SharePoint sites
  orgfiles_site <- get_sharepoint_site(site_url = "https://erstrategies1.sharepoint.com/sites/orgfiles") %>% suppressMessages
  external_site <- get_sharepoint_site(site_url = "https://erstrategies1.sharepoint.com/sites/external") %>% suppressMessages
  data_hub_site <- get_sharepoint_site(site_url = "https://erstrategies1.sharepoint.com/sites/ers-data") %>% suppressMessages

  # Retrieve drives from each SharePoint site
  client_work_drive <- orgfiles_site$get_drive("Client Work")
  internal_drive <- orgfiles_site$get_drive("Internal")
  external_drive <- external_site$get_drive("Documents")
  data_hub_drive <- data_hub_site$get_drive("Documents")

  # Determine which drive to use based on the drive_name parameter
  if (drive_name == "client_work_drive") {
    drive <- client_work_drive
  }
  else if (drive_name == "internal_drive") {
    drive <- internal_drive
  }
  else if (drive_name == "external_drive") {
    drive <- external_drive
  }
  else if (drive_name == "data_hub_drive") {
    drive <- data_hub_drive
  }
  else {
    # If the drive_name is invalid, stop and raise an error
    stop("Invalid drive name. Use 'client_work_drive', 'internal_drive', 'external_drive', or 'data_hub_drive'.")
  }

  # Try to get the folder item from the drive
  folder_item <- try(drive$get_item(folder_path), silent = TRUE)
  if (inherits(folder_item, "try-error")) {
    # Stop execution if the folder path does not exist
    stop("Folder path does not exist: '", folder_path, "'")
  }

  # Construct the full path to the data file within the folder
  data_path <- file.path(folder_path, file_name_with_extension)

  # Try to get the file item from the drive
  file_item <- try(drive$get_item(data_path), silent = TRUE)
  if (inherits(file_item, "try-error")) {
    # Stop execution if the file does not exist
    stop("File '", file_name_with_extension, "' does not exist in folder: '",
         folder_path, "'")
  }

  # Determine the file extension of the target file
  file_extension <- paste0(".", tools::file_ext(file_name_with_extension))

  # Create a temporary file with the appropriate extension for downloading
  temp_file <- tempfile(fileext = file_extension)

  # Download the file from SharePoint to the temporary file
  drive$download_file(data_path, dest = temp_file)

  # Read the file based on its extension
  if (tools::file_ext(file_name_with_extension) == "csv") {
    # For CSV files, use `fread` with specified skip and NA values
    raw_data <- fread(temp_file, skip = skip_rows, na.strings = na_values, colClasses = list(character = "D_location_code"))
  }
  else if (tools::file_ext(file_name_with_extension) %in% c("xls", "xlsx")) {
    # For Excel files, use `read_excel` with the specified sheet name, skip rows, and NA values
    raw_data <- read_excel(temp_file, sheet = sheet_name,
                           skip = skip_rows, na = na_values, guess_max = 100000, col_types = "text")
  }
  else if (tools::file_ext(file_name_with_extension) == "shp") {
    # Create a temporary directory
    temp_dir <- tempdir()
    # Define possible extensions
    shp_components <- c("shp", "shx", "dbf", "prj", "cpg")

    # Download each component
    for (ext in shp_components) {
      file_path <- file.path(temp_dir, paste0(tools::file_path_sans_ext(file_name_with_extension), ".", ext))
      drive$download_file(file.path(folder_path, paste0(tools::file_path_sans_ext(file_name_with_extension), ".", ext)), dest = file_path)
    }

    # Read the shapefile
    raw_data <- st_read(file.path(temp_dir, paste0(tools::file_path_sans_ext(file_name_with_extension), ".shp")))
  }
  else {
    # Stop execution for unsupported file formats
    stop("Unsupported file format. Supported formats are xls, xlsx, csv, or shp: '",
         file_name_with_extension, "'")
  }

  # Check if clean_names is TRUE and file name does NOT contain ".shp"
  if (clean_names == "TRUE" && !grepl("\\.shp$",
                                    file_name_with_extension,
                                    ignore.case = TRUE)) {
    raw_data <- raw_data %>%
      janitor::clean_names()
  }


  # Return the loaded data
  return(raw_data)
}



```

# Part 1: File & Path Variables

## 1.1 Current Cohort Variables

```{r}
#ADJUST THESE VARIABLES AS NEEDED
#current cohort
raw_data_folder_path_current <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2025%20Cohort/1.%20Data%20&%20Analysis%20-%20Secure/2.%20Data%20Coding"


#District List for current cohort (adjust as needed)
district_list <- c("Boulder",
                   "Cleveland",
                   "Columbus",
                   "Denver",
                    "Detroit",
                   "Greenville",
                   "Hamilton",
                   "Horry",
                   "Indianapolis",
                   "Prince William",
                  "Indianapolis_Innovation")

#takes indy innovation out
school_file_list <- c("Boulder",
                   "Cleveland",
                   "Columbus",
                   "Denver",
                    "Detroit",
                   "Greenville",
                   "Hamilton",
                   "Horry",
                   "Indianapolis",
                   "Prince William")
```

## 1.2 Previous Cohort Variables

```{r}
#NOTE UPDATE ONCE COMP DATA IN PERM PLACE

#1. Previous Cohort
#Folder names
raw_data_folder_path_previous <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2024%20Pilot/1.%20Data%20&%20Analysis%20-%20Secure/0.%20Comp%20Data%20Exports/"

#file Names
budgeted_positions_previous_cohort <- "FInal Merged Data for Comp Database.xlsx"
school_file_previous_cohort <- "schools_merged_data_11.27.24.csv"

#DATA DICTIONARY
#File path
data_dictionary_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/1.%20BB%20Resources%20&%20Admin/3.%20BB%20Methodology%20&%20Guidance"

#File Names
bb_data_dictionary_file <- "BB Data Dictionary.xlsx"

```

# Part 2: Clean up Previous Cohort Data

## 2.1 Upload Data

```{r a.bring in previous cohort data}
#budgeted positions
budgeted_positions_previous_cohort <- ers_read_sharepoint_local(folder_path = paste0(raw_data_folder_path_previous),
                   file_name_with_extension = budgeted_positions_previous_cohort,
                   skip_rows=0)

#filter out excluded rows
budgeted_positions_previous_cohort <- budgeted_positions_previous_cohort %>%
 filter(is.na(`Exclude?`) | `Exclude?` != "Exclude")


#schools
schools_previous_cohort <- ers_read_sharepoint_local(folder_path = paste0(raw_data_folder_path_previous),
                   file_name_with_extension = school_file_previous_cohort,
                   skip_rows=0)
```

## 2.2 Column Mapping

```{r}

# Keep ERS names exactly as is - fill with NAs if missing
budgeted_positions_previous_cohort <- budgeted_positions_previous_cohort %>% 
  mutate(
    # Project Info
    H_record_ID  = H_record_id,
    H_partner   = case_when(
      H_partner  == "Indy" ~ "Indianapolis", 
        TRUE ~ H_partner),
    H_LEAID      = case_when(
      H_partner == "Denver"     ~ "0803360",
      H_partner == "Detroit"    ~ "2601103",
      H_partner == "Hamilton"   ~ "4701590",
      H_partner == "Palm Beach" ~ "1201500",
      H_partner == "Indy"       ~ "1804770",
      H_partner == "Jeff Co"    ~ "0804800",
      TRUE ~ "need LEAID"  # keep existing if not matched
    ),
    H_project_year = as.character(H_project_year),
    H_project      = "Budget Benchmarking 2024",

    # FTE
    D_employee_id = D_empl_id,
    D_empl_fte    = NA_real_,  # use NA_character_ if this column is text

    # account codes
    D_activity_code        = NA_character_,
    D_class_code           = NA_character_,
    D_cost_center_code     = NA_character_,
    D_fund_code            = NA_character_,
    D_funct_code           = NA_character_,
    D_dept_code            = NA_character_,
    D_location_code        = NA_character_,
    D_position_type_code   = NA_character_,
    D_position_title_code  = NA_character_,
    D_object_code          = NA_character_,
    D_program_code         = NA_character_,
    D_project_code         = NA_character_,

    # account descriptions 
    D_activity_desc        = D_activity_desc,
    D_allocation_desc      = D_allocation,
    D_class_desc           = D_class_code_desc,
    D_cost_center_desc     = D_cost_center, 
    D_fund_desc            = D_fund_code_desc,
    D_funct_desc           = D_funct_code_desc,
    D_dept_desc            = D_dept_code_desc,
    D_location_desc        = D_location_desc,
    D_position_type_desc   = NA_character_, 
    D_position_title_desc  = D_position_title,
    D_object_desc          = D_object_code_desc,
    D_program_desc         = D_program_desc,
    D_project_desc         = D_project_code_desc,

    # FTE calcs (kept as-is)
    M_fte_adjusted = as.numeric(M_fte_adjusted),
    M_fte_500      = as.numeric(M_fte_500),

    # coding fields
    C_operating_status_final = C_op_status,
    C_reporting_final        = C_reporting,
    C_position_type_final    = C_position_type_final,
    C_function_final         = C_function_final,
    C_student_type_final     = C_student_type_final,
    C_prek_expense_final     = C_prek_final,
    C_sharing_final          = C_sharing_final,

    .keep = "none"
  ) %>%
  select(where(~ !all(is.na(.x))))

```

## 2.3 Aggregate School Metrics

```{r}

school_metrics_aggregated_previous <- schools_previous_cohort |>
  group_by(H_partner) |>
  summarise(
    H_district_enrollment = sum(D_enrollment_total[Alternative_flag == "Neighborhood"], na.rm = TRUE),
    H_schools_enrollment = sum(D_enrollment_total[Alternative_flag == "Neighborhood"], na.rm = TRUE),
    H_gen_ed_enrollment = sum(D_gen_ed_enrollment[Alternative_flag == "Neighborhood"], na.rm = TRUE),
    H_special_ed_enrollment = sum(D_swd_total[Alternative_flag == "Neighborhood"], na.rm = TRUE),
    H_frl_enrollment = sum(D_frl_total[Alternative_flag == "Neighborhood"], na.rm = TRUE),
    H_ell_enrollment = sum(D_ell_total[Alternative_flag == "Neighborhood"], na.rm = TRUE),
    H_school_count = n_distinct(D_location_code[Alternative_flag == "Neighborhood"])
  )
```

## 2.4 Join to Agg School to Position Data

```{r}

# Standardize H_partner names in school_metrics_aggregated_previous
school_metrics_aggregated_previous <- school_metrics_aggregated_previous |>
  mutate(H_partner = trimws(H_partner),
         H_partner = case_when(H_partner == "Indy" ~ "Indianapolis", 
                               TRUE ~ H_partner))


budgeted_positions_previous_cohort <- budgeted_positions_previous_cohort %>%
  left_join(school_metrics_aggregated_previous, 
            by = "H_partner")
```

# Part 3: Clean Up Current Cohort Data

## 3.1 Bring in Budget Data

```{r}
#brings in 2025 data and stack it all
budgeted_positions_list <- district_list %>%
  set_names() %>%
  map(~ ers_read_sharepoint_local(
    folder_path = raw_data_folder_path_current,
    file_name_with_extension = paste0(.x, "_Homebase.xlsx"),
    sheet_name = "Budgeted Positions",
    skip_rows = 5
  ) %>%
    mutate(
      district     = .x,
      H_record_ID  = as.character(H_record_ID)  #  force character
      
    ))


# check one
names(budgeted_positions_list)
glimpse(budgeted_positions_list$Boulder)

```

## 3.2 Stack Current Cohort Data

```{r}
#stack the Data
budgeted_positions_current_cohort <- bind_rows(budgeted_positions_list)

budgeted_positions_current_cohort <- budgeted_positions_current_cohort %>% 
  mutate(
    D_empl_fte = as.numeric(D_empl_fte),
    M_fte_adjusted = as.numeric(M_fte_adjusted)
  )

```

## 3.3 Clean up column Names

```{r}
# rename columns
budgeted_positions_current_cohort <- budgeted_positions_current_cohort %>% 
  rename(H_partner = H_district)
```

## 3.4 Bring in Current School data

```{r}

safe_read_sharepoint <- possibly(
  function(...) ers_read_sharepoint(...),
  otherwise = tibble() # returns empty tibble if file not found
)

#brings in current school and stack it all

schools_list <- school_file_list %>%
  set_names() %>%
  map(~ ers_read_sharepoint_local(
    folder_path = "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2025%20Cohort/1.%20Data%20&%20Analysis%20-%20Secure/1.%20Data%20Processed",
    file_name_with_extension = paste0(.x, "_schools_processed.csv")
  ) %>%
    mutate(
      district     = .x,
      H_record_ID  = as.character(H_record_ID)  #  force character
    ))

# check one
names(schools_list)
glimpse(schools_list$Boulder)
```

## 3.6 Stack the Schools Data

```{r}
#update data types
schools_list <- map(schools_list, ~ .x %>%
  mutate(
    across(
      .cols = setdiff(names(.x), names(.x)[str_detect(names(.x), "_pct|_total|_gen")]),
      as.character
    )
  )
)

#stack it
schools_current_cohort <- bind_rows(schools_list)

#clean up school flag
schools_current_cohort<- schools_current_cohort %>%
mutate(C_school_flag = case_when(C_school_flag == "Include: ALL" ~ "Include: All",
                                 TRUE ~ C_school_flag)) %>%
rename(H_partner = H_district)


```

## 3.7 Aggregate School Metrics

```{r}

school_metrics_aggregated <- schools_current_cohort |>
  group_by(H_partner) |>
  summarise(
    H_district_enrollment = sum(D_enrollment_total[C_school_flag != "Exclude"], na.rm = TRUE),
    H_schools_enrollment = sum(D_enrollment_total[C_school_flag == "Include: All"], na.rm = TRUE),
    H_gen_ed_enrollment = sum(C_gen_ed[C_school_flag == "Include: All"],na.rm = TRUE),
    H_special_ed_enrollment = sum(D_swd_total[C_school_flag == "Include: All"], na.rm = TRUE),
    H_frl_enrollment = sum(D_frl_total[C_school_flag == "Include: All"], na.rm = TRUE),
    H_ell_enrollment = sum(D_ell_total[C_school_flag == "Include: All"], na.rm = TRUE),
    H_school_count = n_distinct(D_location_code[C_school_flag == "Include: All"])
  )

```

## 3.7.1 Special Metrics for Indy

```{r}

indy_data <- schools_current_cohort %>% filter(
  H_partner == "Indianapolis"
)

indy_enrollment_total <- indy_data %>% 
  filter(C_school_flag_reason != "Exclude") %>% 
  summarise(total_enrollment = sum(D_enrollment_total)) %>% 
  pull(total_enrollment)

indy_enrollment_lea <- indy_data %>% 
  filter(C_school_flag_reason %in% c("Neighborhood", "In-LEA Innovation")) %>% 
  summarise(total_enrollment = sum(D_enrollment_total)) %>% 
  pull(total_enrollment)

indy_enrollment_buildings <- indy_data %>% 
  filter(C_school_flag_reason %in% c("Neighborhood", "In-LEA Innovation", "Buildings")) %>% 
  summarise(total_enrollment = sum(D_enrollment_total)) %>% 
  pull(total_enrollment)


indy_schools <- indy_data %>% 
  filter(C_school_flag_reason %in% c("Neighborhood", "In-LEA Innovation")) %>% 
  summarise(total_count = n_distinct(D_location_code)) %>% 
  pull(total_count)


```

## 3.8 Merge Schools Data with Position Data

```{r}
budgeted_positions_current_cohort <- budgeted_positions_current_cohort %>%
  left_join(school_metrics_aggregated, 
            by = "H_partner")
         
```

# Part 4: Combine Years

## 4.1 Stack All Years!

```{r}
#clean up record id data type
budgeted_positions_previous_cohort <-  budgeted_positions_previous_cohort %>%
    mutate(H_record_ID = paste0(as.character(H_record_ID),H_project_year))

#do that stack
budgeted_positions_all_years <- bind_rows(budgeted_positions_current_cohort,budgeted_positions_previous_cohort)
```

```{r}

indy_2024 <- budgeted_positions_all_years %>% 
  filter(H_partner == "Indianapolis") %>% 
  filter(H_project_year == "2024")

indy_2024 <- indy_2024 %>%
    mutate(M_fte_500 = M_fte_500,
           M_fte_10K = M_fte_500 * 20,
           M_fte_per_school =  M_fte_adjusted/H_school_count,
           M_fte_per_student =  M_fte_500 / 500,
           M_fte_gen_ed = M_fte_adjusted/H_gen_ed_enrollment,
           M_fte_sped =  M_fte_adjusted/H_special_ed_enrollment)


```

## 4.2 Calculate FTE Metrics

```{r}
budgeted_positions_all_years <- budgeted_positions_all_years %>%
    mutate(M_fte_500 = M_fte_adjusted/H_schools_enrollment *500,
           M_fte_10K = M_fte_adjusted/H_district_enrollment *10000,
           M_fte_per_school =  M_fte_adjusted/H_school_count,
           M_fte_per_student =  M_fte_adjusted/H_schools_enrollment,
           M_fte_gen_ed = M_fte_adjusted/H_gen_ed_enrollment,
           M_fte_sped =  M_fte_adjusted/H_special_ed_enrollment)
```



## 4.3 Bring in Use Crosswalk

```{r}
budgeted_positions_all_years <- budgeted_positions_all_years %>%
#select(-C_use) %>%
mutate(C_function_final = case_when(C_function_final == "Curriculum & Development" ~ "Curriculum Development",
                                    TRUE ~ C_function_final))
#bring data dictionary
bb_data_dictionary <- ers_read_sharepoint_local(folder_path = data_dictionary_folder_path,
                   file_name_with_extension = bb_data_dictionary_file,
                   skip_rows=4)
```

```{r}
#left join
budgeted_positions_all_years <- budgeted_positions_all_years %>%
  left_join(bb_data_dictionary, by = c("C_function_final" = "Accepted Value")) %>%
  select(-Column) %>%
    rename(C_use = Use)
```

```{r}
indy_2024 <- indy_2024 %>%
  left_join(bb_data_dictionary, by = c("C_function_final" = "Accepted Value")) %>%
  select(-Column) %>%
    rename(C_use = Use)

```


### 4.2.1 Calculate metrics for Indy

```{r}

indy_budget_data <- budgeted_positions_all_years %>% 
  filter(H_partner == "Indianapolis") %>% 
  filter(H_project_year == "2025")

indy_budget_data <- indy_budget_data %>% 
  mutate(M_fte_500 = case_when(
    C_use %in% c("Instruction") ~ M_fte_adjusted / indy_enrollment_lea * 500,
    C_use %in% c("Pupil Services & Enrichment", "ISPG") ~ M_fte_adjusted / indy_enrollment_lea * 500,
    C_function_final %in% c("School Administration") ~ M_fte_adjusted / indy_enrollment_lea * 500,
    C_function_final %in% c("School Supervision") ~ M_fte_adjusted / indy_enrollment_lea * 500,
    
    C_function_final %in% c("Facilities & Maintenance", "Security & Safety") ~ M_fte_adjusted / indy_enrollment_buildings * 500,
    C_function_final %in% c("Research & Accountability", "Student Assignment") ~ M_fte_adjusted / indy_enrollment_buildings * 500,
    C_function_final %in% c("Human Resources", "Data Processing & Information Services", "Legal") ~ M_fte_adjusted / indy_enrollment_buildings * 500,
    
    C_function_final %in% c("Food Services") ~ M_fte_adjusted / indy_enrollment_buildings * 500,
    
    C_function_final %in% c("Student Transportation") ~ M_fte_adjusted / indy_enrollment_buildings * 500,
    
    
    C_function_final %in% c("Governance", "Communications") ~ M_fte_adjusted / indy_enrollment_total * 500,
    C_function_final %in% c("Finance, Budget, Purchasing & Distribution", "Facilities Planning") ~ M_fte_adjusted / indy_enrollment_total * 500,
    
    # Flag for missing functions
    TRUE ~ -1
  )
  
  )


indy_budget_data <- indy_budget_data %>% 
  mutate(M_fte_10K = M_fte_500 * 20,
         M_fte_per_student = M_fte_500 / 500,
         M_fte_per_school = M_fte_adjusted / indy_schools
  )



```

```{r}


without_indy <- budgeted_positions_all_years %>% 
  filter(H_partner != "Indianapolis")


budgeted_positions_all_years <- rbind(without_indy, indy_budget_data, indy_2024)

```

## 4.5 Create Exclusion Flag

```{r}
#exclude flag
budgeted_positions_all_years <- budgeted_positions_all_years |>
  mutate(
    C_exclusion_flag = case_when(
      C_operating_status_final != "PreK-12 Operating" ~ "Exclude",
      C_prek_expense_final == "PreK" ~ "Exclude",
      C_use == "O&M" ~ "Exclude",
      C_position_type_final == "Substitute" ~ "Exclude",
      TRUE ~ "Include"
    ),

 #exclude flag reason
    C_exclusion_flag_reason = case_when(
      C_operating_status_final != "PreK-12 Operating" ~ "Not PreK-12 Operating",
      C_prek_expense_final == "PreK" ~ "PreK Expense",
      C_use == "O&M" ~ "O&M Use",
      C_position_type_final == "Substitute" ~ "Substitute",
      TRUE ~ "Included"
    )
  )
budgeted_positions_all_years |> count(C_exclusion_flag_reason)
```

# Part 5: Normalize Text Columns

## 5.1 Create the Function

```{r}
#this functions removes extra spaces, weird characters, etc. so that text will be exported cleanly
clean_string <- function(x, flatten_newlines = TRUE) {
  y <- suppressWarnings(as.character(x))
  y <- str_replace_all(y, "\u00A0", " ")   # NBSP
  y <- str_replace_all(y, "\u200B", "")    # zero-width space
  y <- stringi::stri_replace_all_regex(y, "[\\p{Cc}&&[^\\t\\n]]", "")
  if (flatten_newlines) y <- str_replace_all(y, "\\r?\\n+", " ")
  y <- str_squish(y)
  y
}

clean_text_cols <- function(df, flatten_newlines = TRUE) {
  df %>%
    mutate(across(
      where(is.character),
      ~ clean_string(.x, flatten_newlines = flatten_newlines)
    ))
}

normalize_keys <- function(df,
                           partner_col = "H_partner",
                           id_cols = c("H_partner", "H_LEAID")) {
  out <- df
  if (partner_col %in% names(out)) {
    out[[partner_col]] <- toupper(clean_string(out[[partner_col]]))
  }
  for (c in id_cols[id_cols %in% names(out)]) {
    out[[c]] <- clean_string(out[[c]])
  }
  out
}

####
prep_for_power_query <- function(df,
                                 partner_col = "H_partner",
                                 id_cols = c("H_partner", "H_LEAID"),
                                 flatten_newlines = TRUE) {
  df %>%
    clean_text_cols(flatten_newlines = flatten_newlines) %>%
    normalize_keys(partner_col = partner_col, id_cols = id_cols)
}
```

## 5.2 Run the Function

```{r}
budgeted_positions_clean <- budgeted_positions_all_years %>%
  prep_for_power_query(partner_col = "H_partner",
                       id_cols = c("H_partner", "H_LEAID"))

schools_current_cohort <- schools_current_cohort %>%
  prep_for_power_query(partner_col = "H_partner",
                       id_cols = c("H_partner", "H_LEAID"))

```

## 5.3 Merge Schools Data

```{r}

# Merge schools onto budgeted positions table
 budgeted_positions_merge<- budgeted_positions_clean |>
  left_join(schools_current_cohort %>%
            select(H_partner, 
                   H_project_year, 
                   D_location_code,
                   D_enrollment_total,
                   C_school_flag,
                   C_school_flag_reason,
                   C_school_level,
                   C_school_level_mapped,
                   C_school_size,
                   M_frl_quartile_flag_sized
                  )
            , by = c("H_partner", "H_project_year", "D_location_code"))

#rename columns
budgeted_positions_merge<-budgeted_positions_merge %>%
rename(S_enrollment_total = D_enrollment_total,
       S_school_level = C_school_level,
       S_school_level_mapped = C_school_level_mapped,
       S_school_size = C_school_size,
       S_school_poverty = M_frl_quartile_flag_sized)

#clean up columns
budgeted_positions_merge<-budgeted_positions_merge %>%
  mutate(
    S_school_level = case_when(is.na(S_school_level) ~ "Non-School", TRUE ~ as.character(S_school_level)),
    S_school_level_mapped = case_when(is.na(S_school_level_mapped) ~ "Non-School", TRUE ~ as.character(S_school_level_mapped)),
    S_school_size = case_when(is.na(S_school_size) ~ "Non-School", TRUE ~ as.character(S_school_size)),
    S_school_poverty = case_when(is.na(S_school_poverty) ~ "Non-School", TRUE ~ as.character(S_school_poverty)),
      C_school_flag = case_when(is.na(C_school_flag) ~ "Non-School", TRUE ~ as.character(C_school_flag)),
        C_school_flag_reason = case_when(is.na(C_school_flag_reason) ~ "Non-School", TRUE ~ as.character(C_school_flag_reason)),
  )


# Anti-join: D_location_codes in budgeted_positions_clean not matched in schools_previous_cohort
unmatched_in_positions <- budgeted_positions_merge |>
  filter(C_reporting_final == "School" & H_project_year == "2025")|>
  anti_join(schools_current_cohort, by = c("H_partner", "H_project_year", "D_location_code")) |>
  select(D_location_code, D_location_desc, H_partner,C_exclusion_flag) |>
  distinct()


# Anti-join: D_location_codes in schools_previous_cohort not matched in budgeted_positions_clean
unmatched_in_school <- schools_current_cohort |>
  anti_join(budgeted_positions_merge, by = c("H_partner", "H_project_year", "D_location_code")) |>
  select(D_location_code, D_location_desc, H_partner, C_school_flag_reason, C_school_flag)

# Combine both sets of unmatched codes
unmatched_codes <- bind_rows(
  unmatched_in_positions |> mutate(source = "budgeted_positions_merge"),
  unmatched_in_school |> mutate(source = "schools_current_cohort")
)

unmatched_codes |> glimpse()

nrow(budgeted_positions_clean)
nrow(budgeted_positions_merge)
ncol(budgeted_positions_clean)
ncol(budgeted_positions_merge)

```

# Part 6: Final Clean ups

## 6.1 Coding Fixes

### Detroit fixes

```{r}

#Detroit Location Fixes
budgeted_positions_merge <- budgeted_positions_merge %>%
  mutate(D_location_code = case_when(H_partner == "DETROIT"  & D_location_desc == "Pulaski Elementary-Middle" ~ "031300",
    TRUE ~ D_location_code
  ))
```

### Denver Data Fixes

```{r}
#Move Denver Data Processing People/Clean Up to match 2024
budgeted_positions_merge <- budgeted_positions_merge %>%
  mutate(C_sharing_final = case_when(H_partner == "DENVER" & H_project_year == "2025" & D_position_title_desc == "PARTNER, SCHOOL TECHNOLOGY" ~ C_sharing_final,
                                     TRUE ~ C_sharing_final),
        C_function_final = case_when(H_partner == "DENVER" & H_project_year == "2025" & D_position_title_desc %in% c("Specialis II School Technology", "SPECIALIST II, SCHOOL TECHNOLOGY") ~ "Data Processing & Information Services",
                                     TRUE ~ C_function_final),
        C_use = case_when(H_partner == "DENVER" & H_project_year == "2025" & D_position_title_desc %in% c("Specialis II School Technology", "SPECIALIST II, SCHOOL TECHNOLOGY") ~ "Business Services",
                                     TRUE ~ C_use))

#check the changes 
budgeted_positions_merge %>%
  filter(H_partner == "DENVER" & H_project_year == "2025" & D_position_title_desc %in% c("Specialis II School Technology", "SPECIALIST II, SCHOOL TECHNOLOGY")) %>%
  group_by(D_position_title_desc, C_function_final, C_use, C_sharing_final) %>%
  summarize(total_fte = sum(M_fte_adjusted, na.rm = TRUE)) %>%
  distinct()


#update Occupational Therapist to "Other" for DETROIT
budgeted_positions_merge <- budgeted_positions_merge %>%
  mutate(C_function_final = case_when(H_partner == "DENVER" & H_project_year == "2024" &
                                      C_function_final == "Development & Fundraising"  ~  "Finance, Budget, Purchasing & Distribution",
                                     TRUE ~ C_function_final))

  

#check the changes
budgeted_positions_merge %>%
  filter(H_partner == "DENVER" & H_project_year == "2024" & C_function_final == "Development & Fundraising") %>%
  group_by(C_function_final, C_use) %>%
  summarize(total_fte = sum(M_fte_adjusted, na.rm = TRUE)) %>%
  distinct() %>%
  print(n = Inf)

```

### Detroit Data Fixes

```{r}
#update Occupational Therapist to "Other" for DETROIT
budgeted_positions_merge <- budgeted_positions_merge %>%
  mutate(C_position_type_final = case_when(H_partner == "DETROIT" & 
                                      C_function_final == "Human Resources" & 
                                      C_position_type_final == "Occupational Therapist" ~ "Other",
                                     TRUE ~ C_position_type_final))


```

### Guidance Counselor Fixes

```{r}
#Guidance Counselor Fixes
budgeted_positions_merge <- budgeted_positions_merge %>%
  mutate(C_function_final = case_when(C_position_type_final == "Guidance Counselor" & H_project_year == "2025" & S_school_level_mapped == "ES" ~ "Social & Emotional",
  C_position_type_final == "Guidance Counselor" & H_project_year == "2025" & S_school_level_mapped == "SS" ~ "Career & Academic Counseling",
                                     TRUE ~ C_function_final))

#check the changes
budgeted_positions_merge %>%
  filter(C_position_type_final == "Guidance Counselor" & H_project_year == "2025") %>%
  group_by(D_position_title_desc, C_function_final, C_use, C_sharing_final, S_school_level_mapped) %>%
  summarize(total_fte = sum(M_fte_adjusted, na.rm = TRUE)) %>%
  distinct() %>%
  print(n = Inf)
```

### Shared Services for OT, PT, and Psychs

```{r}

#update coding
budgeted_positions_merge <- budgeted_positions_merge %>%
  mutate(C_sharing_final = case_when(C_position_type_final %in% c("Occupational Therapist", 
                                                                  "Physical Therapist", 
                                                                  "Psychologist") ~ "Shared Services",
                                     TRUE ~ C_sharing_final))

#update psychologists to Social & Emotional function
budgeted_positions_merge <- budgeted_positions_merge %>%
  mutate(C_function_final = case_when(C_position_type_final == "Psychologist" ~ "Social & Emotional",
                                     TRUE ~ C_function_final))


#check the changes
budgeted_positions_merge %>%
  filter(C_position_type_final %in% c("Occupational Therapist", 
                                      "Physical Therapist", 
                                      "Psychologist")) %>%
  group_by(C_position_type_final, C_function_final, C_use, C_sharing_final) %>%
  summarize(total_fte = sum(M_fte_adjusted, na.rm = TRUE)) %>%
  distinct()

```


## 5.6 Clean up table and Reorder

```{r}
#select columns to keep
budgeted_positions_final <- budgeted_positions_merge %>% 
  select(
    # Project Info
    H_record_ID ,
    H_partner,
    H_LEAID,
    H_project_year,
    H_project , 
    
    #employee
    D_employee_id,
    D_empl_fte ,

   # account descriptions 
    D_activity_desc,
    D_allocation_desc,
    D_class_desc,
    D_cost_center_desc, 
    D_fund_desc,
    D_funct_desc,
    D_dept_desc,
    D_location_code,
    D_location_desc,
    D_position_type_desc, 
    D_position_title_desc,
    D_object_desc,
    D_program_desc,
    D_project_desc,

    # FTE calcs (kept as-is)
    M_fte_adjusted,

    # coding fields
    C_operating_status_final,
    C_reporting_final,
    C_position_type_final,
    C_use,
    C_function_final ,
    C_student_type_final,
    C_prek_expense_final,
    C_sharing_final,

    #school Helpers
    H_district_enrollment,
    H_schools_enrollment,
    H_gen_ed_enrollment,
    H_special_ed_enrollment,
    H_frl_enrollment,
    H_ell_enrollment,
    H_school_count,

    #FTE per Student calcs
    M_fte_500,
    M_fte_10K,
    M_fte_per_school,
    M_fte_per_student,
    M_fte_gen_ed,
    M_fte_sped,

    #School Level Columns
    S_enrollment_total,
    S_school_level,
    S_school_level_mapped,
    S_school_size,
    S_school_poverty,

     #exclusion columns
    C_school_flag,
    C_school_flag_reason,
    C_exclusion_flag,
    C_exclusion_flag_reason
  )
```

# Part 6: Export Stacked Data

```{r}
processed_data_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2025%20Cohort/1.%20Data%20&%20Analysis%20-%20Secure/4.%20STACKED%20DATA"

#budgeted positions
ers_write_sharepoint(data = budgeted_positions_final,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = "budgeted_positions_all_years.csv")

#unmatched schools
ers_write_sharepoint(data = unmatched_codes,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = "unmatched_school_locations.csv")

#schools file
ers_write_sharepoint(data = schools_current_cohort,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = "schools_current_cohort.csv")
```

```{r}

denver <- budgeted_positions_all_years %>% 
  filter(H_partner == "Denver") %>% 
  select(c("D_position_title_desc", "C_function_manual", "C_function_1", "C_function_final", "C_sharing_manual", "C_sharing_final"))



```
## Analysis Check

### Position Type and Function Discrepancies Across Years

```{r}

# Filter for Detroit and included positions
districts_to_check <- c("DENVER", "INDIANAPOLIS", "HAMILTON")

discrepancies_all <- lapply(districts_to_check, function(district) {
  budgeted_positions_final %>%
    filter(H_partner == district, C_exclusion_flag == "Include") %>%
    group_by(D_position_title_desc) %>%
    summarize(
      position_type_2024 = paste(unique(C_position_type_final[H_project_year == "2024"]), collapse = "; "),
      position_type_2025 = paste(unique(C_position_type_final[H_project_year == "2025"]), collapse = "; "),
      function_2024 = paste(unique(C_function_final[H_project_year == "2024"]), collapse = "; "),
      function_2025 = paste(unique(C_function_final[H_project_year == "2025"]), collapse = "; "),
      total_fte_2024 = sum(M_fte_adjusted[H_project_year == "2024"], na.rm = TRUE),
      total_fte_2025 = sum(M_fte_adjusted[H_project_year == "2025"], na.rm = TRUE),
      .groups = "drop"
    ) %>%
    filter(
      position_type_2024 != position_type_2025 |
      function_2024 != function_2025
    ) %>%
    filter(
      !position_type_2024 %in% c("Excluded", "", NA),
      !position_type_2025 %in% c("Excluded", "", NA),
      !function_2024 %in% c("Excluded", "", NA),
      !function_2025 %in% c("Excluded", "", NA)
    ) %>%
    mutate(district = district)
})

discrepancies_all <- bind_rows(discrepancies_all)
discrepancies_all
```

### Sharing discrepancies across years

```{r}
sharing_discrepancies <- budgeted_positions_final %>%
  filter(H_partner %in% districts_to_check, C_exclusion_flag == "Include", C_reporting_final == "Non-School") %>%
  group_by(H_partner, D_position_title_desc) %>%
  summarize(
    function_2024 = paste(unique(C_function_final[H_project_year == "2024"]), collapse = "; "),
    function_2025 = paste(unique(C_function_final[H_project_year == "2025"]), collapse = "; "),
    sharing_2024 = paste(unique(C_sharing_final[H_project_year == "2024"]), collapse = "; "),
    sharing_2025 = paste(unique(C_sharing_final[H_project_year == "2025"]), collapse = "; "),
    total_fte_2024 = sum(M_fte_adjusted[H_project_year == "2024"], na.rm = TRUE),
    total_fte_2025 = sum(M_fte_adjusted[H_project_year == "2025"], na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(sharing_2024 != sharing_2025) %>%
  filter(
    !sharing_2024 %in% c("Excluded", "", NA),
    !sharing_2025 %in% c("Excluded", "", NA)
  )

  sharing_discrepancies <- bind_rows(sharing_discrepancies)
sharing_discrepancies
```

### Columbus & Boulder Deep Dive

```{r}
#filter to columbus data and exp
```

### export analysis

```{r}
ers_write_sharepoint(data = discrepancies_all,
                     folder_path = "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2025%20Cohort/1.%20Data%20&%20Analysis%20-%20Secure/5.%20Cohort-Wide%20Checks",
                     file_name_with_extension = "position_discrepancies.csv")

ers_write_sharepoint(data = sharing_discrepancies,
                     folder_path = "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2025%20Cohort/1.%20Data%20&%20Analysis%20-%20Secure/5.%20Cohort-Wide%20Checks",
                     file_name_with_extension = "function_sharing_discrepancies.csv")

```

```{r}

sapply(budgeted_positions_final, class)


```