---
title: "01_bb_validation"
author: "Jenny Katz"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3   # Customize the depth of the table of contents for HTML document
    toc_float:
      collapsed: true    # Start with the table of contents expanded
      smooth_scroll: true  # Enable smooth scrolling
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "knitted files", output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 0: Read Libraries and connect to Sharepoint

```{r}
erstools::ers_load_packages() 
options(scipen = 99) #prevent any numbers from showing up as scientific notation
```

```{r}
#set sharepoint site URL for access to our org-wide files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"

#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)

#bring in pub available data
nces_data <- read_excel("NCES Data for BB Districts.xlsx")
```

# Part 1: Update Variables

```{r}

#1. folder path to data
raw_data_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2025%20Cohort/1.%20Data%20&%20Analysis%20-%20Secure/0.%20Data%20Received/"

#2. district folder
district_folder <-"Cleveland"

#3. file names 
budgeted_positions_file <- "Employee Information.xlsx"
school_file <- "School Information - Characteristics.xlsx"
enrollment_file <-"Student Information - Enrollment.xlsx"
location_crosswalk_file <-"location_crosswalk_(ers_created).csv"

#4.District & Project INfo
LEAID = "3904378" #Open up nces_data to find
project_year = "2025"
project = "Budget Benchmarking 2025"

```

# Part 2: Bring in Data

## Budgeted Positions

```{r}
#4.Bring in files 
#budgeted positions
budgeted_positions_raw <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = budgeted_positions_file, skip_rows = 5)

```

## Location Crosswalks

```{r}
location_crosswalk <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = location_crosswalk_file)
```

# Part 3: Clean Up Budgeted Positions

## 3.0 Column Mapping

```{r}

#    # Keep ERS names exactly as is - fill with NAs if missing
library(dplyr)
budgeted_positions_raw <- budgeted_positions_raw %>% 
  mutate(

    #Project INfo
    H_record_ID = row_number(),
    H_district = district_folder,
    H_LEAID = LEAID,
    H_project_year = project_year,
    H_project = project,
    
    #FTE
    D_employee_id = `Employee ID`, #critical
    D_empl_fte = `FTE`, #critical
   
    
    #account codes
    D_activity_code     = NA_character_,
    D_allocation_code = NA_character_,
    D_class_code   = NA_character_,
    D_cost_center_code  = NA_character_,
    D_fund_code    = NA_character_, #critical
    D_funct_code   = NA_character_,
    D_dept_code         = NA_character_,
    D_location_code      = "TO FIX", #critical
    D_position_type_code    = NA_character_,
    D_position_title_code    = `Position ID`,
    D_object_code  = NA_character_,
    D_program_code = NA_character_,
    D_project_code = NA_character_,
    
    #account descriptions
    D_activity_desc     = NA_character_,
    D_allocation_desc  = NA_character_,
    D_class_desc   = NA_character_,
    D_cost_center_desc  = `Cost Center`,
    D_fund_desc    = NA_character_,
    D_funct_desc   = NA_character_,
    D_dept_desc    = `Cost Center Hierarchy`, 
    D_location_desc = `Location`, #critical
    D_position_type_desc     = `Job Family Groups`, 
    D_position_title_desc    = `Job Profile`, #critical
    D_object_desc  = NA_character_,
    D_program_desc = NA_character_,
    D_project_desc = NA_character_,
  
    
   #other disrict specific fields to add... update as needed
    D_other_1 =  `Created Moment`,
    D_other_2 =  `Unions All`,
    D_other_3 =  `Position Status`,
    D_other_4 =  NA_character_,
    
     .keep = "none"
   
  ) %>%
 select(where(~ !all(is.na(.x))))

```

```{r}

budgeted_positions_raw <- budgeted_positions_raw %>% 
  filter(D_position_title_code != "Grand Total")


```

## bring in location crosswalk

```{r}

#add locations by detailed location desc
budgeted_positions_raw <- budgeted_positions_raw %>%
mutate(H_location_desc_clean = str_trim(
      str_replace(D_cost_center_desc, "^\\d{4}CC\\s*", "")
      ),
  D_location_code = case_when(
  H_location_desc_clean == "Facing History New Tech High School" ~ "0174CC",
  H_location_desc_clean == "Cleveland Remote High School" ~ "0111CC",
  H_location_desc_clean == "Davis Aerospace and Maritime High School" ~ "0326CC",
  D_location_desc == "Adlai E Stevenson" ~ "0004CC",
  D_location_desc == "Albert B Hart" ~ "0184CC",
  D_location_desc == "Alfred A Benesch" ~ "0076CC",
  D_location_desc == "Almira" ~ "0012CC",
  D_location_desc == "Andrew J Rickoff" ~ "0016CC",
  D_location_desc == "Anton Grdina" ~ "0021CC",
  D_location_desc == "Artemus Ward" ~ "0023CC",
  D_location_desc == "Bard High School Early College" ~ "0209CC",
  D_location_desc == "Benjamin Franklin" ~ "0036CC",
  D_location_desc == "Bolton" ~ "0041CC",
  D_location_desc == "Buhrer Dual Language" ~ "0064CC",
  D_location_desc == "Campus International (K-8)" ~ "0102CC",
  D_location_desc == "Campus International HS" ~ "0282CC",
  D_location_desc == "Charles Mooney-Facing History New Tech High School" ~ "0080CC",
  D_location_desc == "Charles Dickens" ~ "0077CC",
  D_location_desc == "Clara E Westropp" ~ "0090CC",
  D_location_desc == "Clark" ~ "0088CC",
  D_location_desc == "Davis/Digital Arts Campus" ~ "0208CC",
  D_location_desc == "Cleveland School of the Arts" ~ "0801CC",
  D_location_desc == "Collinwood High School" ~ "0096CC",
  D_location_desc == "Daniel E Morgan" ~ "0109CC",
  D_location_desc == "Denison" ~ "0112CC",
  D_location_desc == "Dike School of the Arts" ~ "0124CC",
  D_location_desc == "Douglas MacArthur Girls Leadership Academy" ~ "0130CC",
  D_location_desc == "East Clark" ~ "0148CC",
  D_location_desc == "East Tech/New Tech East" ~ "0161CC",
  D_location_desc == "Euclid Park" ~ "0168CC",
  D_location_desc == "Franklin D Roosevelt" ~ "0172CC",
  D_location_desc == "Garfield" ~ "0188CC",
  D_location_desc == "Garrett Morgan" ~ "0802CC",
  D_location_desc == "Garrett Morgan School of Leadership and Innovation" ~ "0030CC",
  D_location_desc == "George Washington Carver" ~ "0198CC",
  D_location_desc == "Ginn Academy" ~ "0805CC",
  D_location_desc == "Glenville High School" ~ "0220CC",
  D_location_desc == "Halle" ~ "0224CC",
  D_location_desc == "Hannah Gibbons" ~ "0229CC",
  D_location_desc == "Harvey Rice" ~ "0240CC",
  D_location_desc == "John Adams High School" ~ "0268CC",
  D_location_desc == "JFK" ~ "0285CC",
  D_location_desc == "Joseph M Gallagher" ~ "0279CC",
  D_location_desc == "Kenneth W Clement Boys Leadership Academy" ~ "0297CC",
  D_location_desc == "Louisa May Alcott" ~ "0340CC",
  D_location_desc == "Luis Munoz Marin" ~ "0328CC",
  D_location_desc == "Marion C Seltzer" ~ "0353CC",
  D_location_desc == "Marion-Sterling" ~ "0354CC",
  D_location_desc == "Mary B Martin" ~ "0347CC",
  D_location_desc == "Mary Church Terrell School" ~ "0338CC",
  D_location_desc == "Mary M Bethune" ~ "0350CC",
  D_location_desc == "Max S Hayes High School" ~ "0349CC",
  D_location_desc == "MC2STEM Cleveland State University" ~ "0035CC",
  D_location_desc == "MC2STEM Great Lakes Science Center" ~ "0035CC",
  D_location_desc == "MC2STEM Tri-C" ~ "0035CC",
  D_location_desc == "Memorial" ~ "0027CC",
  D_location_desc == "Miles" ~ "0368CC",
  D_location_desc == "Miles Park" ~ "0372CC",
  D_location_desc == "Mound" ~ "0396CC",
  D_location_desc == "Nathan Hale" ~ "0411CC",
  D_location_desc == "Natividad Pagan International Newcomers Academy" ~ "0537CC",
  D_location_desc == "New Tech West" ~ "0060CC",
  D_location_desc == "Oliver H Perry" ~ "0428CC",
  D_location_desc == "Orchard" ~ "0436CC",
  D_location_desc == "Paul L Dunbar" ~ "0451CC",
  D_location_desc == "Riverside" ~ "0485CC",
  D_location_desc == "Robert H Jamison" ~ "0482CC",
  D_location_desc == "Robinson G Jones" ~ "0487CC",
  D_location_desc == "Scranton" ~ "0500CC",
  D_location_desc == "Stephanie Tubbs Jones School" ~ "0448CC",
  D_location_desc == "Stonebrook-White" ~ "0376CC",
  D_location_desc == "Stonebrook Montessori" ~ "0376CC",
  D_location_desc == "Sunbeam" ~ "0532CC",
  D_location_desc == "Tremont Montessori" ~ "0544CC",
  D_location_desc == "Valley View Boys Leadership Academy" ~ "0550CC",
  D_location_desc == "Wade Park" ~ "0556CC",
  D_location_desc == "Warner Girls Leadership Academy" ~ "0572CC",
  D_location_desc == "Waverly" ~ "0596CC",
  D_location_desc == "Whitney M Young" ~ "0078CC",
  D_location_desc == "Wilbur Wright" ~ "0616CC",
  D_location_desc == "William Cullen Bryant" ~ "0622CC",
  D_location_desc == "William Rainey Harper" ~ "0621CC",
  D_location_desc == "Willson" ~ "0624CC",

  ##now update ased on cost center
  D_cost_center_desc == "0159CC Cleveland Early College High School" ~ "0159CC",
  D_cost_center_desc == "0326CC Davis Aerospace and Maritime High School" ~ "0326CC",
  D_cost_center_desc == "0111CC Cleveland Remote High School" ~ "0111CC",
  D_cost_center_desc == "0110CC Cleveland Remote School K-8" ~ "0110CC",
  D_cost_center_desc == "0692CC Cleveland School of Architecture & Design" ~ "0692CC",
  D_cost_center_desc == "0691CC Cleveland School of Science and Medicine" ~ "0691CC",
  D_cost_center_desc == "0300CC John Marshall School of Civic & Business Leadership" ~ "0300CC",
 D_cost_center_desc == "0298CC John Marshall School of Engineering" ~ "0298CC",
 D_cost_center_desc == "0299CC John Marshall School of Information Technology" ~ "0299CC",
 D_cost_center_desc == "0333CC Lincoln-West School of Global Studies" ~ "0333CC",
 D_cost_center_desc == "0334CC Lincoln-West School of Science & Health" ~ "0334CC",
 D_cost_center_desc == "0245CC Rhodes College and Career Academy" ~ "0245CC",
 D_cost_center_desc == "0243CC Rhodes Academy of Environmental Studies" ~ "0243CC",
      TRUE ~ D_location_desc
    ),
#now clean up the location descriptions
  D_location_desc = case_when(D_cost_center_desc %in% c("0159CC Cleveland Early College High School" ,
  "0326CC Davis Aerospace and Maritime High School" ,
 "0111CC Cleveland Remote High School",
 "0110CC Cleveland Remote School K-8" ,
  "0692CC Cleveland School of Architecture & Design",
 "0691CC Cleveland School of Science and Medicine" ,
  "0300CC John Marshall School of Civic & Business Leadership",
  "0298CC John Marshall School of Engineering" ,
 "0299CC John Marshall School of Information Technology" ,
"0333CC Lincoln-West School of Global Studies" ,
 "0334CC Lincoln-West School of Science & Health",
 "0245CC Rhodes College and Career Academy",
 "0243CC Rhodes Academy of Environmental Studies") ~ H_location_desc_clean,
      TRUE ~ D_location_desc
    )
)
#check unmatched locations
budgeted_positions_raw %>%
select(D_location_desc, D_location_code) %>%
distinct()
```

## 3.1 Manual Checks

-   Look at Data Validation table in OneNote and complete (right year, detailed enough, etc.)

## 3.2 Review Table Structure

```{r}
#check number of columns/rows. 
#does this seem right? 
#ALso check for missing data and data types within the file
ers_column_info(budgeted_positions_raw)
```

## 3.3. Remove Blank Rows/Columns

This code block deletes any rows or columns made up fully of blank values ("", NA," ", etc.). These rows and columns wouldn't give you any additional info so it's best practice to take them out early on.

```{r}

#pre-check
ncol(budgeted_positions_raw)
nrow(budgeted_positions_raw)

#remove empty rows and columns
budgeted_positions_raw <- remove_empty(budgeted_positions_raw,
                                 which = c("rows", "cols"))

#post-check
nrow(budgeted_positions_raw)
ncol(budgeted_positions_raw)

```

## 3.4 Check for Duplicates/Remove

If any rows are repetitive or not required for your analysis you can remove them. For example - you have multiple years of data but only need one.

```{r}

#Check for duplicates
budgeted_positions_raw %>%
  filter(!is.na(D_employee_id)) %>% 
 get_dupes(-H_record_ID)

```

## 3.5 Review and modify column data types

Review the data type of each column and modify if needed. Refer to the Data Dictionary for the correct data type for each column. For more information on different data types in R and why it matters, please refer to this: <https://app.tettra.co/teams/ersknowledge/pages/32-r-data-types>

```{r}
#review data type of each column
#ALL Columns except FTE and record ID, should be charachter (this will make joining across tables easier)
col_info <- data.frame(data_type = sapply(budgeted_positions_raw, class))
print(col_info)



#change data type of columns if needed (below is an example to convert to numeric values) - common functions are as.character, as.numeric, as.integer, as.Date

#Created Moment set to charactet
budgeted_positions_raw$D_other_1 <- as.character(budgeted_positions_raw$D_other_1)

```

# Part 4: FTE Adjustments

Employees may have multiple positions or may work in multiple locations and have a FTE of 1 for each position. To get accurate FTE for each employee, project teams should adjust for this and split the FTE across positions/locations.

## 4.0 Compare FTEs to NCES

```{r}
#compare total FTE in budgeted positions with NCES data
public_data_comp_check <- nces_data %>%
  left_join((budgeted_positions_raw %>%
                group_by(H_LEAID) %>%
              summarise(M_total_district_fte = sum(D_empl_fte, na.rm = TRUE))),
              by = c("LEAID" = "H_LEAID")) %>%
  mutate(M_fte_difference = M_total_district_fte  - `Total Staff FTE`)

```

## 4.1 Investigate FTES

First investigate employees whose FTE sums over 1, so we can investigate why and make sure it's not a larger issue with data that we need to go back to district for

```{r}

#count of employees who have an sum of FTE over 1
# 1) compute per-employee adjusted FTE 
budgeted_positions_raw <- budgeted_positions_raw %>%
  group_by(D_employee_id) %>%
  mutate(H_fte_sum = sum(D_empl_fte, na.rm = TRUE)) %>%
  ungroup() 


#2) Review employees whose fte sums > 1
budgeted_positions_raw %>%
  select(D_employee_id,D_position_title_desc,D_location_desc,H_fte_sum) %>%
  filter(H_fte_sum >1)

#
```

## 4.2 Adjust FTEs

If issue is not that big, that we can adjust the employee FTE Down

```{r}
#NOTE:  DO NOT ADJUST VACANT FTE DOWN IF THEY HAVE THE SAME EMPLOYEE ID (NA,0, 1, blanks)

budgeted_positions_raw <- budgeted_positions_raw %>%
  mutate(M_fte_adjusted = round(D_empl_fte, digits =  2))

budgeted_positions_raw %>%
  select(D_employee_id,D_position_title_desc,D_location_desc,M_fte_adjusted) %>%
  filter(M_fte_adjusted >1)

budgeted_positions_raw <- budgeted_positions_raw %>%
  group_by(D_employee_id) %>%
  mutate(
    # total FTE across same job title & ID
    total_fte = sum(M_fte_adjusted, na.rm = TRUE),
    # count of rows in that group
    n_rows = n(),
    # adjust only if total_fte > 1, job titles are the same,
    # and IDs are not in the excluded list
    M_fte_adjusted = case_when(
      !(is.na(D_employee_id)) &
        total_fte > 1 ~ M_fte_adjusted / total_fte,
      TRUE ~ M_fte_adjusted
    )
  ) %>%
  ungroup() %>%
  select(-H_fte_sum, -total_fte, -n_rows)



# Some checks 
  
print(sum(budgeted_positions_raw$D_empl_fte, na.rm = TRUE))

print(sum(budgeted_positions_raw$M_fte_adjusted, na.rm = TRUE))

#2) Double Check that no adjusted FTES are greater than 1
budgeted_positions_raw %>%
  select(D_employee_id,D_position_title_desc,D_location_desc,M_fte_adjusted) %>%
  filter(M_fte_adjusted >1)
```

# Part 6. Clean up Schools file

## 6.1 Bring in School Data

```{r}

#schools files
school_raw <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = school_file,
                   skip_rows = 1)

#enrollment file
enrollment_raw <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = enrollment_file)


```

## CLEVELAND ONLY -MERGE

```{r}
#check umatched rows before merge

unmatched_rows <- school_raw |> anti_join(enrollment_raw, by = c("Code" = "Cost Center"))

# Show the number of unmatched rows
nrow(unmatched_rows)

# Preview unmatched rows
head(unmatched_rows)


#merge enrollment and school
school_raw <- enrollment_raw |>
  left_join(
    school_raw |>
      select(Code, Type),
    by = c("Cost Center" = "Code")
  )

#fine the two schools that don't match are rolled up into MC2STEM so its okay
```

## 6.2 Map District Col to ERS Columns

```{r}
#changing column names and removing unneeded columns
#replace the column names after the "=" with column names from your data.You can press tab after the = to select column names available in the data

school_raw <- school_raw %>% 
  mutate(
    
    #Project INfo
    H_record_ID = row_number(),
    H_district = district_folder,
    H_LEAID = LEAID,
    H_project_year = project_year,
    H_project = project,
    
    #school info
    D_location_code = `Cost Center`,
    D_location_desc = `school name`,
    D_school_level = `school level`,
    D_school_type = `Type`,
    D_grade_span = `gradespan`,
    # These are actually NA's, replacing with 0's so the code wil run
    D_enrollment_total = case_when(is.na(total_enrollment) ~ NA_real_,
                                   is.na(pk) ~ total_enrollment,
                                   TRUE ~ total_enrollment - pk),
    D_preK_enrollment = pk,
    D_ell_total = `Total ELL`,
    D_frl_total =   case_when(is.na(total_enrollment) ~ NA_real_,
                                   is.na(pk) ~ total_enrollment,
                                   TRUE ~ total_enrollment - pk),
    D_swd_total = `Total SPED All`,
    D_self_contained = `Total Locked`,
    
    
     .keep = "none"
  )

  

```

## 6.3 Update Data Type

```{r}
#review data type of each column
#ALL Columns except FTE and record ID, should be charachter (this will make joining across tables easier)
col_info <- data.frame(data_type = sapply(school_raw, class))
print(col_info)



#change data type of columns if needed (below is an example to convert to numeric values) - common functions are as.character, as.numeric, as.integer, as.Date

#budgeted_positions_raw$D_empl_FTE <- as.numeric(budgeted_positions_raw$D_empl_FTE)
```

## 6.4 Review School Structure

```{r}
#check number of columns/rows. 
#does this seem right? 
#ALso check for missing data and data types within the file
ers_column_info(school_raw)
```

## 6.5 Check /Remove Duplicate School

### a. Investigate Dups

```{r}

#Check for duplicates
school_raw %>%
 get_dupes(-H_record_ID)

#find shools with the same school ID
duplicates_by_id <- school_raw  %>%
  get_dupes(D_location_code)

```

### b. Remove Duplicates

```{r}
# No duplicates
```

## 6.6 Add Coding Columns

### a. School Level

```{r}
#view school level
school_raw %>%
  group_by(D_school_level) %>%
  count()

#Update for your district's values 
#No purely MS values in Cleveland
school_raw <-school_raw %>%
  mutate(C_school_level = case_when(D_school_level == "K8"	~ "ES",
                                    D_school_level == "HS" ~ "HS",
                                    TRUE ~ "Other"))
#check coding
school_raw %>%
  group_by(D_school_level,C_school_level) %>%
  count()

```

### b. Alternative School Flag

```{r}
# Will update crosswalks later, but so far these all look like traditional schools
#NOTE THE MC2STEMS were NOt included in enrollment files nor was cleveland remote 

# Maybe return to: Bard HS Early College, Cleveland Early College High School, New Tech West/East

library(dplyr)

school_raw <- school_raw |>
  mutate(
    C_school_flag = case_when(
      D_location_desc == "Natividad Pagan International Newcomers Academy" ~ "Include: District-Wide Only",
      D_location_desc == "Sunbeam" ~ "Include: District-Wide Only",
      D_location_desc == "MC2STEM" ~ "Include: District-Wide Only",
        D_location_desc == "MC2STEM Great Lakes Science Center" ~ "Include: District-Wide Only",
      D_location_desc == "MC2STEM Tri-C" ~ "Include: District-Wide Only",
      D_location_desc == "MC2STEM Cleveland State University" ~ "Include: District-Wide Only",
      D_location_desc == "Cleveland Metro Remote School K-8" ~ "Include: District-Wide Only",
      D_location_desc == "Cleveland Metro Remote School HS" ~ "Include: District-Wide Only",
      TRUE ~ "Include: All"
    )
  )
school_raw <- school_raw %>%
  mutate(
    C_school_flag_reason = case_when(
      D_location_desc == "Natividad Pagan International Newcomers Academy" ~ "Alternative",
      D_location_desc == "Sunbeam" ~ "Alternative",
      D_location_desc == "MC2STEM Great Lakes Science Center" ~ "Alternative",
      D_location_desc == "MC2STEM Tri-C" ~ "Alternative",
      D_location_desc == "MC2STEM Cleveland State University" ~ "Alternative",
      D_location_desc == "MC2STEM" ~ "Alternative",
      D_location_desc == "Cleveland Metro Remote School K-8" ~ "Alternative",
      D_location_desc == "Cleveland Metro Remote School HS" ~ "Alternative",
      TRUE ~ "Neighborhood"
    )
  )

#check coding
school_raw %>%
  group_by(C_school_flag) %>%
  count()

```

### c. Enrollment Characteristics

```{r}

#Update for your district's values 
school_raw <-school_raw %>%
  mutate(C_school_level_mapped = case_when(C_school_level == "ES"	~ "ES",
                                    C_school_level == "HS"	~ "SS",
                                    TRUE ~ "Other"))

school_raw %>%
  group_by(C_school_level,C_school_level_mapped) %>%
  count()


#enrollment data not available for Cleveland
# Keeping the code for now, but all these columns will be NA

school_raw <- school_raw %>% 
  mutate(C_school_size = ifelse(C_school_level == "ES", ifelse(D_enrollment_total < 350, "Small", "Not Small"), ifelse(D_enrollment_total < 500, "Small", "Not Small")))


school_raw <- school_raw %>%
  mutate(C_gen_ed = D_enrollment_total - D_self_contained,
         M_ell_pct = D_ell_total / D_enrollment_total,
         M_swd_pct = as.numeric(D_swd_total) / D_enrollment_total,
         M_frl_pct = D_frl_total / D_enrollment_total)


# Adding in sanity cleanups - you should manually check any huge variations
school_raw <- school_raw %>%
  mutate(M_ell_pct = ifelse(is.na(M_ell_pct), 0, ifelse(M_ell_pct > 1, 1, M_ell_pct)),
         M_swd_pct = ifelse(is.na(M_swd_pct), 0, ifelse(M_swd_pct > 1, 1, M_swd_pct)),
         M_frl_pct = ifelse(is.na(M_frl_pct), 0, ifelse(M_frl_pct > 1, 1, M_frl_pct)))

```

```{r}

# Create all levels quartiles flags
school_raw <- school_raw %>%
  group_by(C_school_flag) %>%
  mutate(
    M_swd_quartile_flag = case_when(
      M_swd_pct <= quantile(M_swd_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_swd_pct > quantile(M_swd_pct, 0.25, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.5, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag) %>%
  mutate(
    M_ell_quartile_flag = case_when(
      M_ell_pct <= quantile(M_ell_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_ell_pct > quantile(M_ell_pct, 0.25, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.5, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag) %>%
  mutate(
    M_frl_quartile_flag = case_when(
      M_frl_pct <= quantile(M_frl_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_frl_pct > quantile(M_frl_pct, 0.25, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.5, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


```

```{r}

# Create all levels quartiles flags
school_raw <- school_raw %>%
  group_by(C_school_flag, C_school_level_mapped) %>%
  mutate(
    M_swd_quartile_flag_sized = case_when(
      M_swd_pct <= quantile(M_swd_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_swd_pct > quantile(M_swd_pct, 0.25, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.5, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag, C_school_level_mapped) %>%
  mutate(
    M_ell_quartile_flag_sized = case_when(
      M_ell_pct <= quantile(M_ell_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_ell_pct > quantile(M_ell_pct, 0.25, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.5, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag, C_school_level_mapped) %>%
  mutate(
    M_frl_quartile_flag_sized = case_when(
      M_frl_pct <= quantile(M_frl_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_frl_pct > quantile(M_frl_pct, 0.25, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.5, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


```

```{r}

# Validating the quartiles
make_quartile_summary <- function(data, var, flag_var) {
  cutoffs <- quantile(data[[var]], probs = c(0.25, 0.5, 0.75), na.rm = TRUE)
  
  counts <- data %>%
    count(!!sym(flag_var)) %>%
    rename(quartile = !!sym(flag_var), count = n)
  
  tibble(
    metric = var,
    q25 = cutoffs[1],
    q50 = cutoffs[2],
    q75 = cutoffs[3]
  ) %>%
    crossing(counts)
}

# build tables for each metric (only looking at traditional campuses)

## Switch filters to small, not-small as needed
data_testing <- school_raw %>% 
  filter(C_school_flag == "Include: All") %>% 
  filter(C_school_size == "Not Small")

frl_summary <- make_quartile_summary(data_testing, "M_frl_pct", "M_frl_quartile_flag")
swd_summary <- make_quartile_summary(data_testing, "M_swd_pct", "M_swd_quartile_flag")
ell_summary <- make_quartile_summary(data_testing, "M_ell_pct", "M_ell_quartile_flag")

# combine into one concise table
quartile_summary <- bind_rows(frl_summary, swd_summary, ell_summary)

print(quartile_summary)

```

## 6.7 Compare Enrollment to NCES

```{r}
public_data_comp_check <- public_data_comp_check %>%
  left_join(
    (school_raw %>%
       group_by(H_LEAID) %>%
       summarise(M_total_district_enrollment = sum(D_enrollment_total, na.rm = TRUE))),
    by = c("LEAID" = "H_LEAID")
  ) %>%
  mutate(
    M_enrollment_difference = M_total_district_enrollment - `Total Enrollment`
  )
```

# Part 7: Position + Schools Crosswalk

### b. Join Position & School Tables

```{r}
#Create a Detailed location crosswalk to update D_location_code in BUdgeted Positions File
budgeted_positions_raw <- budgeted_positions_raw %>%
  mutate(H_detailed_location_desc_helper = case_when(
    str_detect(D_location_desc, "Admin") ~ "Administration",
    TRUE ~ D_location_desc
  ))
```

```{r}

#Create Crosswalk 
location_crosswalk <- bind_rows(
   # --- Part 1: Prepare the Budget Positions Table ---
 budgeted_positions_raw %>%
  filter(!is.na(D_location_code)) %>%
  count(D_location_code, D_location_desc, name = "row_count") %>% 
  mutate( table = "Position") ,

  # --- Part 2: Prepare the Schools Table ---
  school_raw %>%
    count(D_location_code, D_location_desc, name = "row_count") %>% 
    mutate(table = "School")) %>%
    
    
   #Keep only the first school name for each ID ---
   #Keep only the first school name for each ID ---
  group_by(D_location_code) %>%
  arrange(desc(table)) %>%
  mutate(D_location_desc = first(D_location_desc)) %>%
  ungroup() %>%

 #remove duplicates
 group_by(D_location_code, D_location_desc, table) |>
  summarize(row_count = sum(row_count, na.rm = TRUE), .groups = "drop") %>%

  # --- Part 3: Pivot the Combined Table ---
  pivot_wider(
    names_from = table,
    values_from = row_count,
    values_fill = 0
  ) %>%
  arrange(D_location_desc) 




#bring in Total FTE
location_crosswalk <- location_crosswalk %>%
  left_join(
    budgeted_positions_raw %>%
      group_by(D_location_code) %>%
      summarise(M_total_FTE = sum(M_fte_adjusted)) %>%
      filter(M_total_FTE > 0) %>%
      filter(!is.na(D_location_code)) %>%
      select(D_location_code, M_total_FTE),
    by = "D_location_code"
  )

#Bring in Total Enrollment 
location_crosswalk <- location_crosswalk %>%
  left_join(
    school_raw %>%
      select(D_location_code, D_enrollment_total,D_school_type),
    by = "D_location_code"
  ) %>%
  mutate(M_fte_per_500_students = M_total_FTE / D_enrollment_total * 500) %>%
  arrange(desc(School))
  



#flag issues
location_crosswalk <- location_crosswalk %>%
  mutate(C_location_issue_flag = case_when(Position < 1 & School == 1 ~"Not In Position File",
                                         School == 0 & Position > 0 ~ "Not In School File",
                                         M_fte_per_500_students <40 | M_fte_per_500_students > 90 ~ "Low/High FTE per 500",
                                         D_enrollment_total <10 & D_enrollment_total > 0 ~ "Low Enrollment",
                                          TRUE ~ "Okay"))
 print(location_crosswalk)   
 

 positions_locations <- budgeted_positions_raw %>%
  filter(!is.na(D_location_code)) %>%
  count(D_location_code, D_location_desc, name = "row_count") %>% 
  mutate( table = "Position") 
```

### c. Fix Location Codes

Note this will have to be updated for each individual districts

```{r}

# Will validate afterward, but:

# Note Newton D Bauer School of the Arts closed, only appears in the budget file
# budgeted_positions_raw <- budgeted_positions_raw %>%
 #  mutate(
   #  D_location_code = case_when(
  #     D_location_code == "0040CC New Tech East" ~ "0161CC East Tech High School",
  #     D_location_code == "0276CC John Adams High School" ~ "0268CC John Adams College and # Career Academy",
   #    # Had to make a judgement call here
   #    D_location_code == "0330CC Lincoln-West High School" ~ "0333CC Lincoln-West School of Global Studies",
   #    TRUE ~ D_location_code
   #  )
 #  )



```

# Part 8: Machine Learning

## 8.0 Set-Up

```{r}

library(tm)
library(textstem) #for turning words into their stem form
library(randomForest) #for random forest model
library(ranger) #for random forest model
library(xgboost) #for random forest model

#set sharepoint site URL for access to our org-wide and external files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)

client_work_drive <- orgfiles_site$get_drive("Client Work")

# Modified read_sharepoint logic for RDS files
read_sharepoint_ers_obj <- function(drive_name = client_work_drive, 
                                folder_path, 
                                file_name_with_extension, 
                                sheet_name = NULL, 
                                skip_rows = 0,
                                na_values = c("","N/A","NA","Missing")) {
  # Construct the full data path
  data_path <- file.path(folder_path, file_name_with_extension)
  
  file_extension <- paste0(".", tools::file_ext(file_name_with_extension))
  
  # Create a temporary file with the appropriate file extension
  temp_file <- tempfile(fileext = file_extension)
  
  # Download the file from the specified drive to the temporary file
  drive_name$download_file(data_path, dest = temp_file)
  
  # Read the downloaded file based on its file extension
  data_loaded <- readRDS(temp_file)
  
  # Return the loaded data
  return(data_loaded)
}



district_personnel <- budgeted_positions_raw

training_bag_columns_position <- ers_read_sharepoint(folder_path = "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/1.%20BB%20Resources%20&%20Admin/3.%20BB%20Methodology%20&%20Guidance/Machine%20Learning%20Models",
  file_name_with_extension =  "position_type_training_bag_columns_9.24.24.csv")

model_position <- read_sharepoint_ers_obj(folder_path = "Other/Budget Benchmarking/1. BB Resources & Admin/3. BB Methodology & Guidance/Machine Learning Models", file_name_with_extension =  "position_type_model_9.24.24.rds")


training_bag_columns_function <- ers_read_sharepoint(folder_path = "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/1.%20BB%20Resources%20&%20Admin/3.%20BB%20Methodology%20&%20Guidance/Machine%20Learning%20Models",
  file_name_with_extension =  "function_training_bag_columns_9.24.24.csv")

model_function <- read_sharepoint_ers_obj(folder_path = "Other/Budget Benchmarking/1. BB Resources & Admin/3. BB Methodology & Guidance/Machine Learning Models", file_name_with_extension =  "function_model_9.24.24.rds")

```

## 8.1 Modify Data

```{r}

# Add a reporting proxy. To train the ML model, we need a reporting coding, in leiu of a complete set of coding, we can use any proxy to get the job done. In Indy they had a similar column structure that we could use, otherwise you can use whether or not there is a match in the school's file as a nice proxy

district_personnel <- district_personnel %>%
  mutate(
    C_reporting = case_when(
      str_detect(D_location_desc, "Administration") ~ "Non-School",
      TRUE ~ "School"
    )
  )


district_personnel <- district_personnel %>%
  rename(`D_position_title` = `D_position_title_desc`)



```

## 8.2 Load Functions

```{r}

# All functions

#building a prefix function that allows me to add a tag into all words to mark which variable they come from
add_prefix_to_words <- function(string, prefix) {
  # Split the string into words
  words <- unlist(strsplit(string, " "))
  
  # Add prefix to each word
  prefixed_words <- paste0(prefix, words)
  
  # Rejoin the words into a single string
  prefixed_string <- paste(prefixed_words, collapse = " ")
  
  return(prefixed_string)
}

#subset to columns needed (district_data_concat is combination of all fields relevant for prediction and standardized to only regular characters)
prediction_data_prep <- function(data,
                                 predicted_value) {
  data %>%
  mutate(D_position_title = str_replace_all(D_position_title,
                                           "[^[:alnum:]]",
                                           " "),
         D_location_desc = str_replace_all(D_location_desc,
                                           "[^[:alnum:]]",
                                           " "),
         C_reporting = str_replace_all(C_reporting,
                                           "[^[:alnum:]]",
                                           " "),
         D_position_title = sapply(D_position_title,
                                 add_prefix_to_words,
                                 "pt"),
         D_location_desc = sapply(D_location_desc,
                                  add_prefix_to_words,
                                  "ld"),
         C_reporting = sapply(C_reporting,
                              add_prefix_to_words,
                              "rp"),
         district_data_concat = paste(D_position_title,
                                      D_location_desc,
                                      C_reporting,
                                      sep = " ")) %>%
  select(district_data_concat)
}


# Warnings here are fine
bag_prep <- function(data) {
  Corpus(VectorSource(data$district_data_concat)) %>%
  tm_map(content_transformer(tolower)) %>% 
  tm_map(removePunctuation) %>% 
  tm_map(removeNumbers) %>% 
  tm_map(removeWords, stopwords("english")) %>% 
  tm_map(stripWhitespace) %>% 
  DocumentTermMatrix() %>%
  as.matrix() %>% 
  as.data.frame()
}

#adding key terms if they're not already in data (so that select doesn't break)
add_each_col <- function(dataset, col_name) {
  if (!(col_name %in% names(dataset))) {
    dataset <- dataset %>% 
      mutate(!!col_name := 0)
  }
  return(dataset)
}

add_all_cols <- function(dataset) {
  dataset %>% 
  add_each_col("pttchr") %>% 
  add_each_col("pttch") %>% 
  add_each_col("ptasst") %>% 
  add_each_col("ptpara") %>% 
  add_each_col("ptsvcs") %>% 
  add_each_col("ptmgr") %>% 
  add_each_col("ptcoor") %>% 
  add_each_col("ptcoord") %>% 
  add_each_col("ptteacher") %>% 
  add_each_col("ptassistant") %>% 
  add_each_col("ptparaprofessional") %>% 
  add_each_col("ptservices") %>% 
  add_each_col("ptmanager") %>% 
  add_each_col("ptcoordinator")
}

fix_abbrevs <- function(data) {
  data %>% 
  mutate(ptteacher = ifelse(pttchr == 1 | pttch == 1 | ptteacher == 1, 1, 0),
         ptassistant = ifelse(ptasst == 1 | ptassistant == 1, 1, 0),
         ptparaprofessional = ifelse(ptpara ==1 | ptparaprofessional ==1, 1, 0),
         ptservices = ifelse(ptservices == 1 | ptsvcs == 1, 1, 0),
         ptmanager = ifelse(ptmanager == 1 | ptmgr == 1, 1, 0),
         ptcoordinator=ifelse(ptcoord==1|ptcoor==1|ptcoordinator==1, 1, 0)) %>% 
  select(-c(pttchr,
            pttch,
            ptasst,
            ptpara,
            ptsvcs,
            ptmgr,
            ptcoord,
            ptcoor))
}


#removing words that weren't present in training data

align_bags <- function(list_of_cols, test_data) {

  non_training_words <- setdiff(colnames(test_data),
                                list_of_cols$`.`)

  adjusted_test_data <- test_data %>%
    select(-any_of(non_training_words))

  return(adjusted_test_data)
}

add_missing_cols <- function(list_of_cols, test_df) {

  missing_columns <- setdiff(list_of_cols$`.`,
                             names(test_df))

  zero_values_df <- as_tibble(
    matrix(0, ncol = length(missing_columns),
           nrow = nrow(test_df),
           dimnames = list(NULL, missing_columns))
  )

  adjusted_test_df <- bind_cols(test_df,
                                zero_values_df)

  return(adjusted_test_df)
}


prediction_data_prep_exp <- function(data) {
  data %>%
  mutate(D_position_title_h = str_replace_all(D_position_title,
                                           "[^[:alnum:]]",
                                           " "),
         D_location_desc_h = str_replace_all(D_location_desc,
                                           "[^[:alnum:]]",
                                           " "),
         C_reporting_h = str_replace_all(C_reporting,
                                           "[^[:alnum:]]",
                                           " "),
         D_position_title_h = sapply(D_position_title_h,
                                 add_prefix_to_words,
                                 "pt"),
         D_location_desc_h = sapply(D_location_desc_h,
                                  add_prefix_to_words,
                                  "ld"),
         C_reporting_h = sapply(C_reporting_h,
                              add_prefix_to_words,
                              "rp"),
         district_data_concat = paste(D_position_title_h,
                                      D_location_desc_h,
                                      C_reporting_h,
                                      sep = " ")) %>% 
    select(-c(D_location_desc_h, D_position_title_h, C_reporting_h))
}




```

```{r}

# ERS content functions


turn_position_type_into_values <- function(dataset) {
  dataset %>% 
    mutate(C_position_type = case_when(
      C_position_type == 0 ~ "Substitute",
      C_position_type == 1 ~ "Teacher",
      C_position_type == 2 ~ "Other",
      C_position_type == 3 ~ "Nurse",
      C_position_type == 4 ~ "TA",
      C_position_type == 5 ~ "Sec/Clerk/Other Admin",
      C_position_type == 6 ~ "Coordinator/Manager",
      C_position_type == 7 ~ "Principal",
      C_position_type == 8 ~ "Speech Therapist",
      C_position_type == 9 ~ "Guidance Counselor",
      C_position_type == 10 ~ "Custodian",
      C_position_type == 11 ~ "Psychologist",
      C_position_type == 12 ~ "Instructional Coach",
      C_position_type == 13 ~ "Vice Principal",
      C_position_type == 14 ~ "Social Worker",
      C_position_type == 15 ~ "School Monitor/Security",
      C_position_type == 16 ~ "Club Advisor/Coach",
      C_position_type == 17 ~ "Librarian",
      C_position_type == 18 ~ "(Exec) Director",
      C_position_type == 19 ~ "Non-Position",
      C_position_type == 20 ~ "Nurse Aide",
      C_position_type == 21 ~ "Area Officers",
      C_position_type == 22 ~ "Occupational Therapist",
      C_position_type == 23 ~ "Physical Therapist",
      C_position_type == 24 ~ "Instructional Coach Supervisors",
      TRUE ~ NA
    ))
}


turn_function_into_values <- function(dataset) {
  dataset %>% 
    mutate(C_function = case_when(
      C_function == 0 ~ "Other Non-Compensation",
      C_function == 1 ~ "Teacher Compensation",
      C_function == 2 ~ "Aides Compensation",
      C_function == 3 ~ "Substitute Compensation",
      C_function == 4 ~ "Library & Media",
      C_function == 5 ~ "Instructional Materials & Supplies",
      C_function == 6 ~ "Other Compensation",
      C_function == 7 ~ "Extended Time & Tutoring",
      C_function == 8 ~ "Enrichment",
      C_function == 9 ~ "Social & Emotional",
      C_function == 10 ~ "Physical Health & Services",
      C_function == 11 ~ "Career & Academic Counseling",
      C_function == 12 ~ "Parent & Community Relations",
      C_function == 13 ~ "Professional Growth",
      C_function == 14 ~ "Curriculum Development",
      C_function == 15 ~ "Recruitment",
      C_function == 16 ~ "Special Population Program Management & Support",
      C_function == 17 ~ "Governance",
      C_function == 18 ~ "School Supervision",
      C_function == 19 ~ "School Administration",
      C_function == 20 ~ "Research & Accountability",
      C_function == 21 ~ "Communications",
      C_function == 22 ~ "Student Assignment",
      C_function == 23 ~ "Facilities & Maintenance",
      C_function == 24 ~ "Security & Safety",
      C_function == 25 ~ "Food Services",
      C_function == 26 ~ "Student Transportation",
      C_function == 27 ~ "Utilities",
      C_function == 28 ~ "Human Resources",
      C_function == 29 ~ "Finance, Budget, Purchasing & Distribution",
      C_function == 30 ~ "Data Processing & Information Services",
      C_function == 31 ~ "Facilities Planning",
      C_function == 32 ~ "Development & Fundraising",
      C_function == 33 ~ "Legal",
      C_function == 34 ~ "Insurance",
      TRUE ~ "Other Non-Compensation"))
}


```

## 8.3 Train Position Type Model

```{r}


district_personnel_position_type <- prediction_data_prep(district_personnel,
                                                         "C_position_type")

#making all district dataframes a single row per combination of values
district_personnel_position_type <- district_personnel_position_type %>% 
  group_by(district_data_concat) %>% 
  summarize()

# Prep the bags
district_personnel_position_type_test_bag <- bag_prep(district_personnel_position_type)

# Clean abreviations and columns
district_personnel_position_type_test_bag <- add_all_cols(district_personnel_position_type_test_bag)
district_personnel_position_type_test_bag <- fix_abbrevs(district_personnel_position_type_test_bag)


district_personnel_position_type_test_bag <- align_bags(training_bag_columns_position,
                                                        district_personnel_position_type_test_bag)

# Identify and add missing columns in the test data
district_personnel_position_type_test_bag <- add_missing_cols(training_bag_columns_position,                              district_personnel_position_type_test_bag)

set.seed(223)

#ordering columns the exact same order across training and test

  ## Get the column names except "C_position_type"
column_sorting <- sort(setdiff(training_bag_columns_position$'.', "C_position_type"))

new_column_order_test <- c(column_sorting)

position_type_district_test_bag <- district_personnel_position_type_test_bag[, new_column_order_test]



# Make predictions on the test data
rf_position_type_predictions_district <- predict(model_position,
                                                 as.matrix(position_type_district_test_bag))

#turning the predictions into a probability dataframe for each row in original data
probability_matrix_district_position_type <- data.frame(
  row_number = rep(c(1:nrow(district_personnel_position_type)),
                   each = 25),
  C_position_type = rep(c(0:24), nrow(district_personnel_position_type)),
  confidence = rf_position_type_predictions_district)

#shifting back to character values for position type
probability_matrix_district_position_type <-
  turn_position_type_into_values(probability_matrix_district_position_type)

#filtering to row with highest prediction value
probability_matrix_district_position_type <- probability_matrix_district_position_type %>% 
  group_by(row_number) %>% 
  filter(confidence == max(confidence))

#adding prediction and confidence level to original data
district_personnel_position_type <- district_personnel_position_type %>% 
  ungroup() %>% 
  mutate(predicted_value = probability_matrix_district_position_type$C_position_type,
         confidence = probability_matrix_district_position_type$confidence)

## Cleaning up the original data
district_data_export_position <- prediction_data_prep_exp(district_personnel)

district_data_export_position <- district_data_export_position %>% 
  left_join(district_personnel_position_type %>% select(district_data_concat, predicted_value, confidence),
            by = "district_data_concat")


corrected_predictions <- district_data_export_position$predicted_value

corrected_predictions[corrected_predictions == "Teacher"] <- "temp_teacher"
corrected_predictions[corrected_predictions == "Other"] <- "Teacher"
corrected_predictions[corrected_predictions == "Substitute"] <- "Other"
corrected_predictions[corrected_predictions == "temp_teacher"] <- "Substitute"

district_data_export_position$predicted_value <- corrected_predictions

```

## 8.4 Train Function Model

```{r}

district_personnel_function <- prediction_data_prep(district_personnel,
                                                         "C_function")

#making all district dataframes a single row per combination of values
district_personnel_function <- district_personnel_function %>% 
  group_by(district_data_concat) %>% 
  summarize()

# Prep the bags
district_personnel_function_test_bag <- bag_prep(district_personnel_function)

# Clean abreviations and columns
district_personnel_function_test_bag <- add_all_cols(district_personnel_function_test_bag)
district_personnel_function_test_bag <- fix_abbrevs(district_personnel_function_test_bag)

district_personnel_function_test_bag <- align_bags(training_bag_columns_function,
                                                        district_personnel_function_test_bag)
# Identify and add missing columns in the test data
district_personnel_function_test_bag <- add_missing_cols(training_bag_columns_function,                              district_personnel_function_test_bag)


#setting seed for replicability
set.seed(223)

#ordering columns the exact same order across training and test

  ## Get the column names except "C_position_type"
column_sorting <- sort(setdiff(training_bag_columns_function$'.', "C_function"))

new_column_order_test <- c(column_sorting)

function_district_test_bag <- district_personnel_function_test_bag[, new_column_order_test]


# Make predictions on the test data
rf_function_predictions_district <- predict(model_function,
                                                 as.matrix(function_district_test_bag))

#turning the predictions into a probability dataframe for each row in original data
probability_matrix_district_function <- data.frame(
  row_number = rep(c(1:nrow(district_personnel_function)),
                   each = 35),
  C_function = rep(c(0:34), nrow(district_personnel_function)),
  confidence = rf_function_predictions_district)

#shifting back to character values for position type
probability_matrix_district_function <-
  turn_function_into_values(probability_matrix_district_function)

#filtering to row with highest prediction value
probability_matrix_district_function <- probability_matrix_district_function %>% 
  group_by(row_number) %>% 
  filter(confidence == max(confidence))

#adding prediction and confidence level to original data
district_personnel_function <- district_personnel_function %>% 
  ungroup() %>% 
  mutate(predicted_value = probability_matrix_district_function$C_function,
         confidence = probability_matrix_district_function$confidence)


district_data_export_function <- prediction_data_prep_exp(district_personnel)

district_data_export_function <- district_data_export_function %>% 
  left_join(district_personnel_function %>% select(district_data_concat, predicted_value, confidence),
            by = "district_data_concat")

```

## 8.5 Merging predictions together

```{r}

## Change back any columns you changed in 8.1

district_data_export_function <- district_data_export_function %>%
  rename(`D_position_title_desc` = `D_position_title`)

district_data_export_position <- district_data_export_position %>%
  rename(`D_position_title_desc` = `D_position_title`)

shared_cols <- intersect(names(budgeted_positions_raw), names(district_data_export_position))
shared_cols <- intersect(shared_cols, names(district_data_export_function))


budgeted_positions_raw <- budgeted_positions_raw %>%
  # join position predictions
  left_join(
    district_data_export_position %>%
      select(all_of(shared_cols), ML_predicted_position_type = predicted_value),
    by = shared_cols
  ) %>%
  # join function predictions
  left_join(
    district_data_export_function %>%
      select(all_of(shared_cols), ML_predicted_function = predicted_value),
    by = shared_cols
  )


```

# Part 9: Upload File to Sharepoint

## Part 9.1 Any final clean up

Add any final clean up to your file below (remove unncessary columns, update column names to be standard etc.)

```{r}
processed_data_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2025%20Cohort/1.%20Data%20&%20Analysis%20-%20Secure/1.%20Data%20Processed"

#budgeted positions
ers_write_sharepoint(data = budgeted_positions_raw,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = paste0(district_folder,"_budgeted_positions_processed.csv"))


#Schools File
ers_write_sharepoint(data = school_raw,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = paste0(district_folder,"_schools_processed.csv"))

#location crosswalk
ers_write_sharepoint(data = location_crosswalk,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = paste0(district_folder,"_location_crosswalk.csv"))
```