---
title: "01_bb_validation"
author: "Jenny Katz"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3   # Customize the depth of the table of contents for HTML document
    toc_float:
      collapsed: true    # Start with the table of contents expanded
      smooth_scroll: true  # Enable smooth scrolling
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "knitted files", output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 0: Read Libraries and connect to Sharepoint

```{r}
erstools::ers_load_packages()
options(scipen = 99) #prevent any numbers from showing up as scientific notation
```

```{r}
#set sharepoint site URL for access to our org-wide files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"

#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)

#bring in pub available data
nces_data <- read_excel("NCES Data for BB Districts.xlsx")
```

# Part 1: Update Variables

```{r}

#1. folder path to data
raw_data_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2025%20Cohort/1.%20Data%20&%20Analysis%20-%20Secure/0.%20Data%20Received/"

#2. district folder
district_folder <-"Columbus"


#3. file names 
budgeted_positions_file <- "Budget Benchmarking Data Updated.xlsx"
school_file <- "Budget Benchmarking Data Updated.xlsx"


#4.District & Project INfo
LEAID = "3904380" #Open up nces_data to find
project_year = "2025"
project = "Budget Benchmarking 2025"

```

# Part 2: Bring in Data

```{r}
#4.Bring in files 
#budgeted positions
budgeted_positions_raw <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = budgeted_positions_file, sheet_name = "Employee Accounts")


budgeted_positions_raw_original <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = budgeted_positions_file, sheet_name = "Employees")

#rules 

```

```{r}

## Looking at the different rows

fte1 <- budgeted_positions_raw %>%
  group_by(`Employee Number`) %>%
  summarise(FTE_sum_1 = sum(`CY FTE Percent`, na.rm = TRUE))

fte2 <- budgeted_positions_raw_original %>%
  group_by(`Employee Number`) %>%
  summarise(FTE_sum_2 = sum(`CY FTE Percent`, na.rm = TRUE))

# Step 2: Join them together on employee_id
fte_compare <- full_join(fte1, fte2, by = "Employee Number")

# Step 3: Identify differences
fte_diff <- fte_compare %>%
  filter(FTE_sum_1 != FTE_sum_2 | is.na(FTE_sum_1) | is.na(FTE_sum_2))

# Step 4: View results
print(fte_diff)

```

## 2.1 Loading in the Crosswalks

```{r}

chart_of_accounts <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = budgeted_positions_file, sheet_name = "Chart of Accounts")

chart_of_accounts <- chart_of_accounts %>% 
  select(c(`Seg Title`, `Seg Value`, `Seg Description`))


```

## 2.3 Merge with crosswalks

```{r}

budgeted_positions_raw <- budgeted_positions_raw %>%
  separate(`Account Details`, into = c("Fund Code", "Function Code", "Special Cost Center", "Object", "Subject Area", "OPU", "Grade Level", "Job Assignment", "Budget Supervisor", "Extra"), sep = "-")

temp <- chart_of_accounts %>% 
  filter(`Seg Title` == "Fund")

budgeted_positions_raw$Fund_Description <- temp$`Seg Description`[match(budgeted_positions_raw$`Fund Code`, temp$`Seg Value`)]

temp <- chart_of_accounts %>% 
  filter(`Seg Title` == "Function")

budgeted_positions_raw$Function_Description <- temp$`Seg Description`[match(budgeted_positions_raw$`Function Code`, temp$`Seg Value`)]

temp <- chart_of_accounts %>% 
  filter(`Seg Title` == "Special Cost Center")

budgeted_positions_raw$Cost_Center_Description <- temp$`Seg Description`[match(budgeted_positions_raw$`Special Cost Center`, temp$`Seg Value`)]

temp <- chart_of_accounts %>% 
  filter(`Seg Title` == "Object")

budgeted_positions_raw$Object_Description <- temp$`Seg Description`[match(budgeted_positions_raw$`Object`, temp$`Seg Value`)]

temp <- chart_of_accounts %>% 
  filter(`Seg Title` == "Subject Area")

budgeted_positions_raw$Subject_Description <- temp$`Seg Description`[match(budgeted_positions_raw$`Subject Area`, temp$`Seg Value`)]

temp <- chart_of_accounts %>% 
  filter(`Seg Title` == "OPU")

budgeted_positions_raw$OPU_Description <- temp$`Seg Description`[match(budgeted_positions_raw$`OPU`, temp$`Seg Value`)]

temp <- chart_of_accounts %>% 
  filter(`Seg Title` == "Grade Level")

budgeted_positions_raw$Grade_Description <- temp$`Seg Description`[match(budgeted_positions_raw$`Grade Level`, temp$`Seg Value`)]

temp <- chart_of_accounts %>% 
  filter(`Seg Title` == "Job Assignment")


budgeted_positions_raw$Job_Assignment_Description <- temp$`Seg Description`[match(budgeted_positions_raw$`Job Assignment`, temp$`Seg Value`)]

temp <- chart_of_accounts %>% 
  filter(`Seg Title` == "Budget Supervisor")

budgeted_positions_raw$Budget_Supervisor_Description <- temp$`Seg Description`[match(budgeted_positions_raw$`Budget Supervisor`, temp$`Seg Value`)]

```

# Part 3: Clean Up Budgeted Positions

## 3.0 Column Mapping

```{r}

#    # Keep ERS names exactly as is - fill with NAs if missing
budgeted_positions_raw <- budgeted_positions_raw %>% 
  mutate(

    #Project INfo
    H_record_ID = row_number(),
    H_district = district_folder,
    H_LEAID = LEAID,
    H_project_year = project_year,
    H_project = project,
    
    #FTE
    D_employee_id = `Employee Number`, #critical
    D_empl_fte = `CY FTE Percent`, #critical
   
    
    #account codes
    D_activity_code     = `Grade Level`,
    D_allocation_code = NA_character_,
    D_class_code   = `Job Assignment`,
    D_cost_center_code  = `Special Cost Center`,
    D_fund_code    = `Fund Code`, #critical
    D_funct_code   = `Function Code`,
    D_dept_code         = OPU,
    D_location_code      = Location, #critical
    D_position_type_code    = `Job Class`,
    D_position_title_code    = Position,
    D_object_code  = Object,
    D_program_code = Subject_Description,
    D_project_code = NA_character_,
    
    #account descriptions
    D_activity_desc     = Grade_Description,
    D_allocation_desc  = `Account Desc`,
    D_class_desc   = Job_Assignment_Description,
    D_cost_center_desc  = Cost_Center_Description,
    D_fund_desc    = Fund_Description, #critical
    D_funct_desc   = Function_Description,
    D_dept_desc    = OPU_Description, 
    D_location_desc = `Location Desc`, #critical
    D_position_type_desc     = `Job Class Desc`, 
    D_position_title_desc    = `Position Desc`, #critical
    D_object_desc  = Object_Description,
    D_program_desc = Subject_Description,
    D_project_desc = NA_character_,
  
    
   #other disrict specific fields to add... update as needed
    D_other_1 =  Budget_Supervisor_Description,
    D_other_2 =  `Budget Supervisor`,
    D_other_3 =  NA_character_,
    D_other_4 =  NA_character_,
    
     .keep = "none"
   
  ) %>%
 select(where(~ !all(is.na(.x))))

```

## 3.1 Manual Checks

-   Look at Data Validation table in OneNote and complete (right year, detailed enough, etc.)

## 3.2 Review Table Structure

```{r}
#check number of columns/rows. 
#does this seem right? 
#ALso check for missing data and data types within the file
ers_column_info(budgeted_positions_raw)
```

## 3.3. Remove Blank Rows/Columns

This code block deletes any rows or columns made up fully of blank values ("", NA," ", etc.). These rows and columns wouldn't give you any additional info so it's best practice to take them out early on.

```{r}

#pre-check
ncol(budgeted_positions_raw)
nrow(budgeted_positions_raw)

#remove empty rows and columns
budgeted_positions_raw <- remove_empty(budgeted_positions_raw,
                                 which = c("rows", "cols"))

#post-check
nrow(budgeted_positions_raw)
ncol(budgeted_positions_raw)

```

## 3.4 Check for Duplicates/Remove

If any rows are repetitive or not required for your analysis you can remove them. For example - you have multiple years of data but only need one.

```{r}

#Check for duplicates
budgeted_positions_raw %>%
 get_dupes(-H_record_ID)

```

## 3.5 Review and modify column data types

Review the data type of each column and modify if needed. Refer to the Data Dictionary for the correct data type for each column. For more information on different data types in R and why it matters, please refer to this: <https://app.tettra.co/teams/ersknowledge/pages/32-r-data-types>

```{r}
#review data type of each column
#ALL Columns except FTE and record ID, should be charachter (this will make joining across tables easier)
col_info <- data.frame(data_type = sapply(budgeted_positions_raw, class))
print(col_info)



#change data type of columns if needed (below is an example to convert to numeric values) - common functions are as.character, as.numeric, as.integer, as.Date


```

# Part 4: FTE Adjustments

Employees may have multiple positions or may work in multiple locations and have a FTE of 1 for each position. To get accurate FTE for each employee, project teams should adjust for this and split the FTE across positions/locations.

## 4.0 Compare FTEs to NCES

```{r}
#compare total FTE in budgeted positions with NCES data
public_data_comp_check <- nces_data %>%
  left_join((budgeted_positions_raw %>%
                group_by(H_LEAID) %>%
              summarise(M_total_district_fte = sum(D_empl_fte, na.rm = TRUE))),
              by = c("LEAID" = "H_LEAID")) %>%
  mutate(M_fte_difference = M_total_district_fte  - `Total Staff FTE`)

```

## 4.1 Investigate FTES

First investigate employees whose FTE sums over 1, so we can investigate why and make sure it's not a larger issue with data that we need to go back to district for

```{r}

#count of employees who have an sum of FTE over 1
# 1) compute per-employee adjusted FTE 
budgeted_positions_raw <- budgeted_positions_raw %>%
  group_by(D_employee_id) %>%
  mutate(H_fte_sum = sum(D_empl_fte, na.rm = TRUE)) %>%
  ungroup() 


#2) Review employees whose fte sums > 1
budgeted_positions_raw %>%
  select(D_employee_id,D_position_title_desc,D_location_desc,H_fte_sum) %>%
  filter(H_fte_sum >1)

#
```

```{r}
# Turning coaching positions to 0

budgeted_positions_raw <- budgeted_positions_raw %>%
  group_by(D_employee_id) %>%
  mutate(H_fte_sum = sum(D_empl_fte, na.rm = TRUE)) %>%
  ungroup() 


budgeted_positions_raw <- budgeted_positions_raw %>%
  mutate(
    # Clean up whitespace and tabs
    D_position_title_desc = str_squish(D_position_title_desc),
    
    # Set FTE = 0 for certain positions when total FTE > 1
    D_empl_fte = if_else(
      H_fte_sum > 1 & str_detect(
        D_position_title_desc,
        regex("\\b(VARSITY|VOLLEYBALL|SOCCER|FOOTBALL|X COUNTRY|TENNIS|STRENGTH & CONDITIONING)\\b",
              ignore_case = TRUE)
      ),
      0,
      D_empl_fte
    )
  )


```

## 4.2 Adjust FTEs for the duplicated position numbers

```{r}

budgeted_positions_raw <- budgeted_positions_raw %>%
  group_by(D_employee_id, D_position_title_code) %>%
  mutate(
    # total FTE across same job title & ID
    total_fte = sum(D_empl_fte, na.rm = TRUE),
    # count of rows in that group
    n_rows = n(),
    # adjust only if total_fte > 1, job titles are the same,
    # and IDs are not in the excluded list
    M_fte_adjusted = case_when(
        n_rows > 1 ~ D_empl_fte / n_rows,
      TRUE ~ D_empl_fte
    )
  ) %>%
  ungroup()



```

```{r}

print(sum(budgeted_positions_raw$D_empl_fte))

print(sum(budgeted_positions_raw$M_fte_adjusted))


budgeted_positions_raw %>%
  group_by(D_employee_id) %>% 
  filter(sum(M_fte_adjusted) >1)


```

# Now, performing traditional FTE adjustments for the remaining employees above

```{r}

budgeted_positions_raw <- budgeted_positions_raw %>%
  group_by(D_employee_id) %>%
  mutate(
    # total FTE across same job title & ID
    total_fte = sum(M_fte_adjusted, na.rm = TRUE),
    # count of rows in that group
    n_rows = n(),
    # adjust only if total_fte > 1, job titles are the same,
    # and IDs are not in the excluded list
    M_fte_adjusted = case_when(
      total_fte > 1 ~ D_empl_fte / total_fte,
      TRUE ~ M_fte_adjusted
    )
  ) %>%
  ungroup() %>%
  select(-H_fte_sum, -total_fte, -n_rows)


```

```{r}

print(sum(budgeted_positions_raw$D_empl_fte))

print(sum(budgeted_positions_raw$M_fte_adjusted))


budgeted_positions_raw %>%
  group_by(D_employee_id) %>% 
  filter(sum(M_fte_adjusted) >1)

```

# Part 2: Bring in Vaccancy Data

```{r}
#4.Bring in files 
#budgeted positions
vaccancies <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = "Budgeted Vacancies.xlsx", sheet_name = "Position Inquiry Export")



```

\`

## 2.3 Merge with crosswalks

```{r}

vaccancies <- vaccancies %>%
  separate(`Account Details`, into = c("Fund Code", "Function Code", "Special Cost Center", "Object", "Subject Area", "OPU", "Grade Level", "Job Assignment", "Budget Supervisor", "Extra"), sep = "-")

temp <- chart_of_accounts %>% 
  filter(`Seg Title` == "Fund")

vaccancies$Fund_Description <- temp$`Seg Description`[match(vaccancies$`Fund Code`, temp$`Seg Value`)]

temp <- chart_of_accounts %>% 
  filter(`Seg Title` == "Function")

vaccancies$Function_Description <- temp$`Seg Description`[match(vaccancies$`Function Code`, temp$`Seg Value`)]

temp <- chart_of_accounts %>% 
  filter(`Seg Title` == "Special Cost Center")

vaccancies$Cost_Center_Description <- temp$`Seg Description`[match(vaccancies$`Special Cost Center`, temp$`Seg Value`)]

temp <- chart_of_accounts %>% 
  filter(`Seg Title` == "Object")

vaccancies$Object_Description <- temp$`Seg Description`[match(vaccancies$`Object`, temp$`Seg Value`)]

temp <- chart_of_accounts %>% 
  filter(`Seg Title` == "Subject Area")

vaccancies$Subject_Description <- temp$`Seg Description`[match(vaccancies$`Subject Area`, temp$`Seg Value`)]

temp <- chart_of_accounts %>% 
  filter(`Seg Title` == "OPU")

vaccancies$OPU_Description <- temp$`Seg Description`[match(vaccancies$`OPU`, temp$`Seg Value`)]

temp <- chart_of_accounts %>% 
  filter(`Seg Title` == "Grade Level")

vaccancies$Grade_Description <- temp$`Seg Description`[match(vaccancies$`Grade Level`, temp$`Seg Value`)]

temp <- chart_of_accounts %>% 
  filter(`Seg Title` == "Job Assignment")


vaccancies$Job_Assignment_Description <- temp$`Seg Description`[match(vaccancies$`Job Assignment`, temp$`Seg Value`)]

temp <- chart_of_accounts %>% 
  filter(`Seg Title` == "Budget Supervisor")

vaccancies$Budget_Supervisor_Description <- temp$`Seg Description`[match(vaccancies$`Budget Supervisor`, temp$`Seg Value`)]

```

# Part 3: Clean Up Budgeted Positions

## 3.0 Column Mapping

```{r}

#    # Keep ERS names exactly as is - fill with NAs if missing
vaccancies <- vaccancies %>% 
  mutate(

    #Project INfo
    # Offset for the original positions
    H_record_ID = row_number() + 10018,
    H_district = district_folder,
    H_LEAID = LEAID,
    H_project_year = project_year,
    H_project = project,
    
    #FTE
    D_employee_id = 0, #critical
    D_empl_fte = `CY FTE Percent`, #critical
   
    
    #account codes
    D_activity_code     = `Grade Level`,
    D_allocation_code = NA_character_,
    D_class_code   = `Job Assignment`,
    D_cost_center_code  = `Special Cost Center`,
    D_fund_code    = `Fund Code`, #critical
    D_funct_code   = `Function Code`,
    D_dept_code         = OPU,
    D_location_code      = Location, #critical
    D_position_type_code    = `Job Class`,
    D_position_title_code    = Position,
    D_object_code  = Object,
    D_program_code = Subject_Description,
    D_project_code = NA_character_,
    
    #account descriptions
    D_activity_desc     = Grade_Description,
    D_allocation_desc  = `Account Desc`,
    D_class_desc   = Job_Assignment_Description,
    D_cost_center_desc  = Cost_Center_Description,
    D_fund_desc    = Fund_Description, #critical
    D_funct_desc   = Function_Description,
    D_dept_desc    = OPU_Description, 
    D_location_desc = `Location Desc`, #critical
    D_position_type_desc     = `Job Class Desc`, 
    D_position_title_desc    = `Position Desc`, #critical
    D_object_desc  = Object_Description,
    D_program_desc = Subject_Description,
    D_project_desc = NA_character_,
  
    
   #other disrict specific fields to add... update as needed
    D_other_1 =  Budget_Supervisor_Description,
    D_other_2 =  `Budget Supervisor`,
    D_other_3 =  NA_character_,
    D_other_4 =  NA_character_,
    
     .keep = "none"
   
  ) %>%
 select(where(~ !all(is.na(.x))))

```

## 3.1 Manual Checks

-   Look at Data Validation table in OneNote and complete (right year, detailed enough, etc.)

## 3.2 Review Table Structure

```{r}
#check number of columns/rows. 
#does this seem right? 
#ALso check for missing data and data types within the file
ers_column_info(vaccancies)
```

## 3.3. Remove Blank Rows/Columns

This code block deletes any rows or columns made up fully of blank values ("", NA," ", etc.). These rows and columns wouldn't give you any additional info so it's best practice to take them out early on.

```{r}

#pre-check
ncol(vaccancies)
nrow(vaccancies)

#remove empty rows and columns
budgeted_positions_raw <- remove_empty(budgeted_positions_raw,
                                 which = c("rows", "cols"))

#post-check
nrow(vaccancies)
ncol(vaccancies)

```

## 3.4 Check for Duplicates/Remove

If any rows are repetitive or not required for your analysis you can remove them. For example - you have multiple years of data but only need one.

```{r}

#Check for duplicates
vaccancies %>%
 get_dupes(-H_record_ID)

```

## 3.5 Review and modify column data types

Review the data type of each column and modify if needed. Refer to the Data Dictionary for the correct data type for each column. For more information on different data types in R and why it matters, please refer to this: <https://app.tettra.co/teams/ersknowledge/pages/32-r-data-types>

```{r}
#review data type of each column
#ALL Columns except FTE and record ID, should be charachter (this will make joining across tables easier)
col_info <- data.frame(data_type = sapply(vaccancies, class))
print(col_info)



#change data type of columns if needed (below is an example to convert to numeric values) - common functions are as.character, as.numeric, as.integer, as.Date


```

```{r}
# No FTE Adjustments needed since no employee ids and allocations are all 1.

vaccancies <- vaccancies %>% 
  mutate(M_fte_adjusted = D_empl_fte)

```

```{r}

budgeted_positions_raw <- rbind(budgeted_positions_raw, vaccancies)


```

# Part 6. Clean up Schools file

## 6.1 Bring in School Data

```{r}

#schools files
school_raw <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = school_file, sheet_name = 'Schools') %>% head(-1)

school_file_sped <- "SPED Student Counts by School Grade Classroom Type.xlsx"

school_sped_raw <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = school_file_sped, sheet_name = 'Sheet2')

```

## 6.2 Map District Col to ERS Columns

### a. construct school grade column

```{r}
# ---- 1. Define grade columns and friendly names ----
GRADE_COLS <- c(
   "KG", "1",
  "2", "3", "4",
  "5", "6", "7",
  "8", "9", "10", "11", "12", ">12"
)

GRADE_NAMES <- c("K","1","2","3","4","5","6","7","8","9","10","11","12", ">12")

# only keep columns that exist in your file
present_cols  <- intersect(GRADE_COLS, names(school_raw))
present_names <- GRADE_NAMES[match(present_cols, GRADE_COLS)]
print(present_names)

# ---- 2. Build grades_offered for each school ----
school_raw <- school_raw %>%
  rowwise() %>%
  mutate(
    grades_offered = list({
      vals <- c_across(all_of(present_cols))
      nums <- suppressWarnings(as.numeric(vals))
      nums[is.na(nums)] <- 0
      present_names[which(nums > 0)]
    })
  ) %>%
  ungroup()

# ---- 3. Create readable string and grade span ----
library(purrr)
school_raw <- school_raw %>%
  mutate(
    grades_offered_str = map_chr(grades_offered, ~ if (length(.x)) paste(.x, collapse = ", ") else NA_character_),
    grade_span = map_chr(grades_offered, ~ 
                           if (length(.x)) paste0(first(.x), "–", last(.x)) else NA_character_)
    )

# ---- 5. Sanity check ----
school_raw %>%
  select(grades_offered_str, grade_span) %>%
  head(102)
```

### c. clean up sped file 

```{r}
school_sped_clean <- school_sped_raw %>%
  group_by(`School Code`, `School Name`) %>%
 summarise(
    `LC` = sum(`LC`, na.rm = TRUE),
    `SELC` = sum(`SELC`, na.rm = TRUE),
    `SLC` = sum(`SLC`, na.rm = TRUE),
    .groups = "drop"
  ) %>%
    mutate(M_self_contained = `SELC` + `SLC`)

#clean nase and convert school code to numeric for merging
school_sped_clean <- school_sped_clean |>
  mutate(`School Code` = str_trim(`School Code`)) |>
  mutate(`School Code` = str_remove_all(`School Code`, "[^0-9]")) |>
  mutate(`School Code` = as.numeric(`School Code`))

#match check between school_sped_clean and school_raw
school_sped_clean %>%
  anti_join(school_raw, by = c("School Code" = "School Number")) %>%
  print(n = Inf)

#match check between school_sped_clean and school_raw
school_raw %>%
  anti_join(school_sped_clean, by = c("School Number" = "School Code")) %>%
  print(n = Inf)

#merge the files
school_raw <- school_raw |>
  left_join(
    school_sped_clean |>
      select(`School Code`, M_self_contained),
    by = c("School Number" = "School Code") # match the correct name
  )


```

### b. Create a school level flag

```{r}

school_raw <-school_raw %>%
  mutate(school_level= case_when(
    grade_span %in% c("PK–8","K–8") ~ "ES",
    grade_span %in% c("K–5", "PK–5", "PK–3","1–5", "PK–6", "K–3", "4–5", "3–8", "K–6") ~ "ES",
    grade_span %in% c("6–8") ~ "MS",
    grade_span %in% c("9–12","9–>12","9-10", "10–>12", "9–10") ~ "HS",
    grade_span %in% c("7–>12","7–12","6–12") ~ "HS", 
    grade_span %in% c("PK–PK") ~ "PK",
    TRUE ~ "Other")
  )


school_raw %>%
  group_by(grade_span,school_level) %>%
  arrange(grade_span) %>%
  count()

#changing column names and removing unneeded columns
#replace the column names after the "=" with column names from your data.You can press tab after the = to select column names available in the data

school_raw <- school_raw %>% 
  mutate(
    
    #Project INfo
    H_record_ID = row_number(),
    H_district = district_folder,
    H_LEAID = LEAID,
    H_project_year = project_year,
    H_project = project,
    
    #school info
    D_location_code = `School Number`,
    D_location_desc = `School Name`,
    D_school_level = `school_level`,
    D_school_type = NA_character_,
    D_grade_span = `grade_span`,
    D_enrollment_total = `Total Enrollment`,
    D_preK_enrollment = NA_character_,
    D_ell_total = `LEP - YES`,
    D_frl_total = `ED - YES`,
    D_swd_total = `IEP - YES`,
    D_self_contained = M_self_contained,
    .keep = "none"
  )
```

```{r}
#Waiting on this information

# Adding in self contained counts

#totals <- school_center_program %>%
  #group_by(`Oracle Location`) %>%
  #summarise(D_self_contained = sum(`Students Enrolled`, na.rm = TRUE), .groups = "drop")

#school_raw <- school_raw %>%
 # left_join(totals, by = c("D_location_code" = "Oracle Location")) %>% 
  #mutate(D_self_contained = replace_na(D_self_contained, 0))


```

## 6.3 Update Data Type

```{r}
#review data type of each column
#ALL Columns except FTE and record ID, should be charachter (this will make joining across tables easier)
col_info <- data.frame(data_type = sapply(school_raw, class))
print(col_info)



#change data type of columns if needed (below is an example to convert to numeric values) - common functions are as.character, as.numeric, as.integer, as.Date

school_raw$D_location_code <- sprintf("%03d", school_raw$D_location_code)

school_raw$D_location_code <- as.character(school_raw$D_location_code)
```

## 6.4 Review School Structure

```{r}
#check number of columns/rows. 
#does this seem right? 
#ALso check for missing data and data types within the file
ers_column_info(school_raw)
```

## 6.5 Check /Remove Duplicate School

### a. Investigate Dups

```{r}

#Check for duplicates
school_raw %>%
 get_dupes(-H_record_ID)


#deuplicate
school_raw<- school_raw %>%
  distinct()

#find shools with the same school ID
duplicates_by_id <- school_raw  %>%
  get_dupes(D_location_code)

```

## 6.6 Add Coding Columns

### a. School Level

```{r}
#school level coded above before assigning new columns

school_raw$C_school_level = school_raw$D_school_level
```

### b. Alternative School Flag

```{r}
#UPDATE BASED on DISTRICT'S ALT SCHOOLS
school_raw <- school_raw %>%
  mutate(C_school_flag = "Include: All") %>%   # start with default
  mutate(
    C_school_flag = case_when(
      D_location_desc ==  "Columbus Gifted Academy" ~ "Include: District-Wide Only",
      D_location_desc ==  "Columbus Scioto 6-12" ~ "Include: District-Wide Only",
      D_location_desc ==  "Beatty Park Elementary School" ~ "Include: District-Wide Only",
      D_location_desc ==  "Fort Hayes Career Center" ~ "Include: District-Wide Only",
      D_location_desc ==  "Columbus Downtown High School" ~ "Include: District-Wide Only",
      D_location_desc ==  "Columbus Global Academy" ~ "Include: District-Wide Only",
      D_location_desc ==  "Colerain Elementary School" ~ "Include: District-Wide Only",
      D_location_desc ==  "Columbus Online Academy" ~ "Include: District-Wide Only",
      TRUE ~ C_school_flag
    ),
    C_school_flag = case_when(
      D_school_level  %in% c("PK") ~ "Exclude",
      TRUE ~ C_school_flag
    )
  )

school_raw <- school_raw %>%
  mutate(C_school_flag_reason = "Neighborhood") %>%   # start with default
  mutate(
    C_school_flag_reason = case_when(
      D_location_desc ==  "Columbus Gifted Academy" ~ "Alternative",
      D_location_desc ==  "Columbus Scioto 6-12" ~ "Special Education",
      D_location_desc ==  "Beatty Park Elementary School" ~ "Special Education",
      D_location_desc ==  "Fort Hayes Career Center" ~ "Alternative",
      D_location_desc ==  "Columbus Downtown High School" ~ "Alternative",
      D_location_desc ==  "Columbus Global Academy" ~ "Alternative",
      D_location_desc ==  "Colerain Elementary School" ~ "Special Education",
      D_location_desc ==  "Columbus Online Academy" ~ "Virtual",
      TRUE ~ C_school_flag_reason
    ),
    C_school_flag_reason = case_when(
      D_school_level %in% "PK" ~ "PreK",
      TRUE ~ C_school_flag_reason
    )
  )

#check coding
school_raw %>%
  group_by(C_school_flag_reason) %>%
  count()

```

### c. Enrollment Characteristics

```{r}

#Update for your district's values 
school_raw <-school_raw %>%
  mutate(C_school_level_mapped = case_when(D_school_level == "ES"	~ "ES",
                                    D_school_level == "MS"	~ "SS",
                                    D_school_level == "HS"	~ "SS",
                                    D_school_level == "SS" ~ "SS",
                                    TRUE ~ "Other"))

school_raw %>%
  group_by(D_school_level,C_school_level_mapped) %>%
  count()

school_raw <- school_raw %>% 
  mutate(C_school_size = ifelse(D_school_level == "ES", ifelse(D_enrollment_total < 350, "Small", "Not Small"), ifelse(D_enrollment_total < 500, "Small", "Not Small")))

#NEED TO RETURN TO: Subtract self-contained students when that information is provided
school_raw <- school_raw %>%
  mutate(C_gen_ed = coalesce(D_enrollment_total, 0) - coalesce(D_self_contained, 0),
         M_ell_pct = D_ell_total / D_enrollment_total,
         M_swd_pct = as.numeric(D_swd_total) / D_enrollment_total,
         M_frl_pct = D_frl_total / D_enrollment_total)


# Adding in sanity cleanups - you should manually check any huge variations
school_raw <- school_raw %>%
  mutate(M_ell_pct = ifelse(is.na(M_ell_pct), 0, ifelse(M_ell_pct > 1, 1, M_ell_pct)),
         M_swd_pct = ifelse(is.na(M_swd_pct), 0, ifelse(M_swd_pct > 1, 1, M_swd_pct)),
         M_frl_pct = ifelse(is.na(M_frl_pct), 0, ifelse(M_frl_pct > 1, 1, M_frl_pct)))

#calculate %self-contained
school_raw %>%
  mutate(M_self_contained_pct = D_self_contained / D_enrollment_total,
         M_self_contained_pct = ifelse(is.na(M_self_contained_pct), 0, ifelse(M_self_contained_pct > 1, 1, M_self_contained_pct))) %>%
  select(D_location_code, D_location_desc, D_enrollment_total, D_self_contained, M_self_contained_pct) %>%

         print(n=Inf)
```

```{r}

# Create all levels quartiles flags
school_raw <- school_raw %>%
  group_by(C_school_flag) %>%
  mutate(
    M_swd_quartile_flag = case_when(
      M_swd_pct <= quantile(M_swd_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_swd_pct > quantile(M_swd_pct, 0.25, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.5, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag) %>%
  mutate(
    M_ell_quartile_flag = case_when(
      M_ell_pct <= quantile(M_ell_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_ell_pct > quantile(M_ell_pct, 0.25, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.5, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag) %>%
  mutate(
    M_frl_quartile_flag = case_when(
      M_frl_pct <= quantile(M_frl_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_frl_pct > quantile(M_frl_pct, 0.25, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.5, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


```

```{r}

# Create all levels quartiles flags
school_raw <- school_raw %>%
  group_by(C_school_flag, C_school_level_mapped) %>%
  mutate(
    M_swd_quartile_flag_sized = case_when(
      M_swd_pct <= quantile(M_swd_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_swd_pct > quantile(M_swd_pct, 0.25, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.5, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag, C_school_level_mapped) %>%
  mutate(
    M_ell_quartile_flag_sized = case_when(
      M_ell_pct <= quantile(M_ell_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_ell_pct > quantile(M_ell_pct, 0.25, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.5, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag, C_school_level_mapped) %>%
  mutate(
    M_frl_quartile_flag_sized = case_when(
      M_frl_pct <= quantile(M_frl_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_frl_pct > quantile(M_frl_pct, 0.25, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.5, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


```

```{r}

# Validating the quartiles
make_quartile_summary <- function(data, var, flag_var) {
  cutoffs <- quantile(data[[var]], probs = c(0.25, 0.5, 0.75), na.rm = TRUE)
  
  counts <- data %>%
    count(!!sym(flag_var)) %>%
    rename(quartile = !!sym(flag_var), count = n)
  
  tibble(
    metric = var,
    q25 = cutoffs[1],
    q50 = cutoffs[2],
    q75 = cutoffs[3]
  ) %>%
    crossing(counts)
}

# build tables for each metric (only looking at traditional campuses)

## Switch filters to small, not-small as needed
data_testing <- school_raw %>% 
  filter(C_school_flag == "Include: All")

library(tidyr)
frl_summary <- make_quartile_summary(data_testing, "M_frl_pct", "M_frl_quartile_flag")
swd_summary <- make_quartile_summary(data_testing, "M_swd_pct", "M_swd_quartile_flag")
ell_summary <- make_quartile_summary(data_testing, "M_ell_pct", "M_ell_quartile_flag")

# combine into one concise table
quartile_summary <- bind_rows(frl_summary, swd_summary, ell_summary)

print(quartile_summary)
```

## 6.7 Compare Enrollment to NCES

```{r}

public_data_comp_check <- public_data_comp_check %>%
  left_join(
    (school_raw %>%
       group_by(H_LEAID) %>%
       summarise(M_total_district_enrollment = sum(D_enrollment_total, na.rm = TRUE))),
    by = c("LEAID" = "H_LEAID")
  ) %>%
  mutate(
    M_enrollment_difference = M_total_district_enrollment - `Total Enrollment`
  )
```

# Part 7: Position + Schools Crosswalk

```{r}

# Fixing online HS location code to map to the budget

school_raw <- school_raw %>% 
  mutate(D_location_code = case_when(
    D_location_code == "718" ~ "981",
    TRUE ~ D_location_code
  ))

#fix budget file so it merges Wooward Park @ walden with Woodward Park
budgeted_positions_raw <- budgeted_positions_raw %>% 
  mutate(D_location_code = case_when(
    D_location_code == "650" ~ "274",
    TRUE ~ D_location_code
  ))

```

### b. Join Position & School Tables

```{r}

#Create Crosswalk 
standardize_school_names <- function(name) {
  name %>%
    str_to_title() %>%
    str_trim() %>%
    str_replace_all("\\bEs\\b", "Elementary School") %>%
    str_replace_all("\\bHs\\b", "High School") %>%
    str_replace_all("\\bMs\\b", "Middle School") %>%
    str_replace_all("\\bPrimary\\b", "Elementary School") %>%
    str_replace_all("(Elementary School).*", "\\1")  %>%
    str_replace_all("\\bAlt\\b", "Alternative") %>%
    str_replace_all("(Alternative).*", "\\1") %>%
    str_replace_all("^Devonshire Alternative$", "Devonshire Elementary School") %>%
    str_replace_all("^Duxberry Park Alternative$", "Duxberry Park Elementary School") %>%
    str_replace_all("^Eastmoor Academy$", "Eastmoor Academy High School") %>%
    str_replace_all("^Fairwood Alternative$", "Fairwood Elementary School")
}

#Create Crosswalk 
location_crosswalk <- bind_rows(
   # --- Part 1: Prepare the Budget Positions Table ---
 budgeted_positions_raw %>%
  filter(!is.na(D_location_code)) %>%
  count(D_location_code, D_location_desc, name = "row_count") %>% 
  mutate( table = "Position") ,

  # --- Part 2: Prepare the Schools Table ---
  school_raw %>%
    count(D_location_code, D_location_desc, name = "row_count") %>% 
    mutate(table = "School")

) %>% 
  
   #Keep only the first school name for each ID ---
  group_by(D_location_code) %>%
  arrange(desc(table)) %>%
  mutate(D_location_desc = first(D_location_desc)) %>%
  ungroup() %>%

  # --- Part 3: Pivot the Combined Table ---
  pivot_wider(
    names_from = table,
    values_from = row_count,
    values_fill = 0 
  ) %>%
  arrange(D_location_desc)


#bring in Total FTE
location_crosswalk <- location_crosswalk %>%
  left_join(
    budgeted_positions_raw %>%
      group_by(D_location_code, D_location_desc) %>%
      summarise(M_total_FTE = sum(M_fte_adjusted)) %>%
      filter(M_total_FTE > 0) %>%
      filter(!is.na(D_location_code)) %>%
      select(D_location_code, M_total_FTE),
    by = "D_location_code"
  )

#Bring in Total Enrollment 
location_crosswalk <- location_crosswalk %>%
  left_join(
    school_raw %>%
      select(D_location_code, D_enrollment_total,D_school_type),
    by = "D_location_code"
  ) %>%
  mutate(M_fte_per_500_students = M_total_FTE / D_enrollment_total * 500) %>%
  arrange(desc(School))
  

#flag issues
location_crosswalk <- location_crosswalk %>%
  mutate(
    C_location_issue_flag = case_when(
      Position < 1 & School == 1 ~ "Not In Position File",
      School == 0 &
        Position > 0 ~ "Not In School File",
      M_fte_per_500_students < 40 |
        M_fte_per_500_students > 90 ~ "Low/High FTE per 500",
      D_enrollment_total < 10 &
        D_enrollment_total > 0 ~ "Low Enrollment",
      TRUE ~ "Okay"
    )
  )

print(location_crosswalk)

```

### c. Add Flags

Note this will have to be updated for each individual districts

```{r}

location_crosswalk <- location_crosswalk %>%
  left_join(
    school_raw %>%
      select(D_location_code, C_school_flag,C_school_flag_reason),
    by = "D_location_code")
```

# Part 8: Machine Learning

## 8.0 Set-Up

```{r}
#Add Reporting flag for proxy..
budgeted_positions_raw <- budgeted_positions_raw %>%
  left_join(
    school_raw %>%
      select(D_location_code, D_location_desc_new = D_location_desc),
    by = "D_location_code"
  ) %>%
  mutate(C_reporting_0 = if_else(is.na(D_location_desc_new), 
                               "Non-School", "School")) %>%
  select(-D_location_desc_new)
```

```{r}

library(tm)
library(textstem) #for turning words into their stem form
library(randomForest) #for random forest model
library(ranger) #for random forest model
library(xgboost) #for random forest model

#set sharepoint site URL for access to our org-wide and external files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)

client_work_drive <- orgfiles_site$get_drive("Client Work")

# Modified read_sharepoint logic for RDS files
read_sharepoint_ers_obj <- function(drive_name = client_work_drive, 
                                folder_path, 
                                file_name_with_extension, 
                                sheet_name = NULL, 
                                skip_rows = 0,
                                na_values = c("","N/A","NA","Missing")) {
  # Construct the full data path
  data_path <- file.path(folder_path, file_name_with_extension)
  
  file_extension <- paste0(".", tools::file_ext(file_name_with_extension))
  
  # Create a temporary file with the appropriate file extension
  temp_file <- tempfile(fileext = file_extension)
  
  # Download the file from the specified drive to the temporary file
  drive_name$download_file(data_path, dest = temp_file)
  
  # Read the downloaded file based on its file extension
  data_loaded <- readRDS(temp_file)
  
  # Return the loaded data
  return(data_loaded)
}



district_personnel <- budgeted_positions_raw

training_bag_columns_position <- ers_read_sharepoint(folder_path = "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/1.%20BB%20Resources%20&%20Admin/3.%20BB%20Methodology%20&%20Guidance/Machine%20Learning%20Models",
  file_name_with_extension =  "position_type_training_bag_columns_9.24.24.csv")

model_position <- read_sharepoint_ers_obj(folder_path = "Other/Budget Benchmarking/1. BB Resources & Admin/3. BB Methodology & Guidance/Machine Learning Models", file_name_with_extension =  "position_type_model_9.24.24.rds")


training_bag_columns_function <- ers_read_sharepoint(folder_path = "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/1.%20BB%20Resources%20&%20Admin/3.%20BB%20Methodology%20&%20Guidance/Machine%20Learning%20Models",
  file_name_with_extension =  "function_training_bag_columns_9.24.24.csv")

model_function <- read_sharepoint_ers_obj(folder_path = "Other/Budget Benchmarking/1. BB Resources & Admin/3. BB Methodology & Guidance/Machine Learning Models", file_name_with_extension =  "function_model_9.24.24.rds")
```

## 8.1 Modify Data

```{r}

# Add a reporting proxy. To train the ML model, we need a reporting coding, in leiu of a complete set of coding, we can use any proxy to get the job done. In Indy they had a similar column structure that we could use, otherwise you can use whether or not there is a match in the school's file as a nice proxy

#district_personnel <- district_personnel %>% 
 # mutate(D_dept_desc = ifelse(D_dept_desc == "School", "School", "Non-School"))

district_personnel <- district_personnel %>%
  rename(C_reporting = C_reporting_0, `D_position_title` = `D_position_title_desc`)
  


```

## 8.2 Load Functions

```{r}

# All functions

#building a prefix function that allows me to add a tag into all words to mark which variable they come from
add_prefix_to_words <- function(string, prefix) {
  # Split the string into words
  words <- unlist(strsplit(string, " "))
  
  # Add prefix to each word
  prefixed_words <- paste0(prefix, words)
  
  # Rejoin the words into a single string
  prefixed_string <- paste(prefixed_words, collapse = " ")
  
  return(prefixed_string)
}

#subset to columns needed (district_data_concat is combination of all fields relevant for prediction and standardized to only regular characters)
prediction_data_prep <- function(data,
                                 predicted_value) {
  data %>%
  mutate(D_position_title = str_replace_all(D_position_title,
                                           "[^[:alnum:]]",
                                           " "),
         D_location_desc = str_replace_all(D_location_desc,
                                           "[^[:alnum:]]",
                                           " "),
         C_reporting = str_replace_all(C_reporting,
                                           "[^[:alnum:]]",
                                           " "),
         D_position_title = sapply(D_position_title,
                                 add_prefix_to_words,
                                 "pt"),
         D_location_desc = sapply(D_location_desc,
                                  add_prefix_to_words,
                                  "ld"),
         C_reporting = sapply(C_reporting,
                              add_prefix_to_words,
                              "rp"),
         district_data_concat = paste(D_position_title,
                                      D_location_desc,
                                      C_reporting,
                                      sep = " ")) %>%
  select(district_data_concat)
}


# Warnings here are fine
bag_prep <- function(data) {
  Corpus(VectorSource(data$district_data_concat)) %>%
  tm_map(content_transformer(tolower)) %>% 
  tm_map(removePunctuation) %>% 
  tm_map(removeNumbers) %>% 
  tm_map(removeWords, stopwords("english")) %>% 
  tm_map(stripWhitespace) %>% 
  DocumentTermMatrix() %>%
  as.matrix() %>% 
  as.data.frame()
}

#adding key terms if they're not already in data (so that select doesn't break)
add_each_col <- function(dataset, col_name) {
  if (!(col_name %in% names(dataset))) {
    dataset <- dataset %>% 
      mutate(!!col_name := 0)
  }
  return(dataset)
}

add_all_cols <- function(dataset) {
  dataset %>% 
  add_each_col("pttchr") %>% 
  add_each_col("pttch") %>% 
  add_each_col("ptasst") %>% 
  add_each_col("ptpara") %>% 
  add_each_col("ptsvcs") %>% 
  add_each_col("ptmgr") %>% 
  add_each_col("ptcoor") %>% 
  add_each_col("ptcoord") %>% 
  add_each_col("ptteacher") %>% 
  add_each_col("ptassistant") %>% 
  add_each_col("ptparaprofessional") %>% 
  add_each_col("ptservices") %>% 
  add_each_col("ptmanager") %>% 
  add_each_col("ptcoordinator")
}

fix_abbrevs <- function(data) {
  data %>% 
  mutate(ptteacher = ifelse(pttchr == 1 | pttch == 1 | ptteacher == 1, 1, 0),
         ptassistant = ifelse(ptasst == 1 | ptassistant == 1, 1, 0),
         ptparaprofessional = ifelse(ptpara ==1 | ptparaprofessional ==1, 1, 0),
         ptservices = ifelse(ptservices == 1 | ptsvcs == 1, 1, 0),
         ptmanager = ifelse(ptmanager == 1 | ptmgr == 1, 1, 0),
         ptcoordinator=ifelse(ptcoord==1|ptcoor==1|ptcoordinator==1, 1, 0)) %>% 
  select(-c(pttchr,
            pttch,
            ptasst,
            ptpara,
            ptsvcs,
            ptmgr,
            ptcoord,
            ptcoor))
}


#removing words that weren't present in training data

align_bags <- function(list_of_cols, test_data) {

  non_training_words <- setdiff(colnames(test_data),
                                list_of_cols$`.`)

  adjusted_test_data <- test_data %>%
    select(-any_of(non_training_words))

  return(adjusted_test_data)
}

add_missing_cols <- function(list_of_cols, test_df) {

  missing_columns <- setdiff(list_of_cols$`.`,
                             names(test_df))

  zero_values_df <- as_tibble(
    matrix(0, ncol = length(missing_columns),
           nrow = nrow(test_df),
           dimnames = list(NULL, missing_columns))
  )

  adjusted_test_df <- bind_cols(test_df,
                                zero_values_df)

  return(adjusted_test_df)
}


prediction_data_prep_exp <- function(data) {
  data %>%
  mutate(D_position_title_h = str_replace_all(D_position_title,
                                           "[^[:alnum:]]",
                                           " "),
         D_location_desc_h = str_replace_all(D_location_desc,
                                           "[^[:alnum:]]",
                                           " "),
         C_reporting_h = str_replace_all(C_reporting,
                                           "[^[:alnum:]]",
                                           " "),
         D_position_title_h = sapply(D_position_title_h,
                                 add_prefix_to_words,
                                 "pt"),
         D_location_desc_h = sapply(D_location_desc_h,
                                  add_prefix_to_words,
                                  "ld"),
         C_reporting_h = sapply(C_reporting_h,
                              add_prefix_to_words,
                              "rp"),
         district_data_concat = paste(D_position_title_h,
                                      D_location_desc_h,
                                      C_reporting_h,
                                      sep = " ")) %>% 
    select(-c(D_location_desc_h, D_position_title_h, C_reporting_h))
}




```

```{r}

# ERS content functions


turn_position_type_into_values <- function(dataset) {
  dataset %>% 
    mutate(C_position_type = case_when(
      C_position_type == 0 ~ "Substitute",
      C_position_type == 1 ~ "Teacher",
      C_position_type == 2 ~ "Other",
      C_position_type == 3 ~ "Nurse",
      C_position_type == 4 ~ "TA",
      C_position_type == 5 ~ "Sec/Clerk/Other Admin",
      C_position_type == 6 ~ "Coordinator/Manager",
      C_position_type == 7 ~ "Principal",
      C_position_type == 8 ~ "Speech Therapist",
      C_position_type == 9 ~ "Guidance Counselor",
      C_position_type == 10 ~ "Custodian",
      C_position_type == 11 ~ "Psychologist",
      C_position_type == 12 ~ "Instructional Coach",
      C_position_type == 13 ~ "Vice Principal",
      C_position_type == 14 ~ "Social Worker",
      C_position_type == 15 ~ "School Monitor/Security",
      C_position_type == 16 ~ "Club Advisor/Coach",
      C_position_type == 17 ~ "Librarian",
      C_position_type == 18 ~ "(Exec) Director",
      C_position_type == 19 ~ "Non-Position",
      C_position_type == 20 ~ "Nurse Aide",
      C_position_type == 21 ~ "Area Officers",
      C_position_type == 22 ~ "Occupational Therapist",
      C_position_type == 23 ~ "Physical Therapist",
      C_position_type == 24 ~ "Instructional Coach Supervisors",
      TRUE ~ NA
    ))
}


turn_function_into_values <- function(dataset) {
  dataset %>% 
    mutate(C_function = case_when(
      C_function == 0 ~ "Other Non-Compensation",
      C_function == 1 ~ "Teacher Compensation",
      C_function == 2 ~ "Aides Compensation",
      C_function == 3 ~ "Substitute Compensation",
      C_function == 4 ~ "Library & Media",
      C_function == 5 ~ "Instructional Materials & Supplies",
      C_function == 6 ~ "Other Compensation",
      C_function == 7 ~ "Extended Time & Tutoring",
      C_function == 8 ~ "Enrichment",
      C_function == 9 ~ "Social & Emotional",
      C_function == 10 ~ "Physical Health & Services",
      C_function == 11 ~ "Career & Academic Counseling",
      C_function == 12 ~ "Parent & Community Relations",
      C_function == 13 ~ "Professional Growth",
      C_function == 14 ~ "Curriculum Development",
      C_function == 15 ~ "Recruitment",
      C_function == 16 ~ "Special Population Program Management & Support",
      C_function == 17 ~ "Governance",
      C_function == 18 ~ "School Supervision",
      C_function == 19 ~ "School Administration",
      C_function == 20 ~ "Research & Accountability",
      C_function == 21 ~ "Communications",
      C_function == 22 ~ "Student Assignment",
      C_function == 23 ~ "Facilities & Maintenance",
      C_function == 24 ~ "Security & Safety",
      C_function == 25 ~ "Food Services",
      C_function == 26 ~ "Student Transportation",
      C_function == 27 ~ "Utilities",
      C_function == 28 ~ "Human Resources",
      C_function == 29 ~ "Finance, Budget, Purchasing & Distribution",
      C_function == 30 ~ "Data Processing & Information Services",
      C_function == 31 ~ "Facilities Planning",
      C_function == 32 ~ "Development & Fundraising",
      C_function == 33 ~ "Legal",
      C_function == 34 ~ "Insurance",
      TRUE ~ "Other Non-Compensation"))
}


```

## 8.3 Train Position Type Model

```{r}


district_personnel_position_type <- prediction_data_prep(district_personnel,
                                                         "C_position_type")

#making all district dataframes a single row per combination of values
district_personnel_position_type <- district_personnel_position_type %>% 
  group_by(district_data_concat) %>% 
  summarize()

# Prep the bags
district_personnel_position_type_test_bag <- bag_prep(district_personnel_position_type)

# Clean abreviations and columns
district_personnel_position_type_test_bag <- add_all_cols(district_personnel_position_type_test_bag)
district_personnel_position_type_test_bag <- fix_abbrevs(district_personnel_position_type_test_bag)


district_personnel_position_type_test_bag <- align_bags(training_bag_columns_position,
                                                        district_personnel_position_type_test_bag)

# Identify and add missing columns in the test data
district_personnel_position_type_test_bag <- add_missing_cols(training_bag_columns_position,                              district_personnel_position_type_test_bag)

set.seed(223)

#ordering columns the exact same order across training and test

  ## Get the column names except "C_position_type"
column_sorting <- sort(setdiff(training_bag_columns_position$'.', "C_position_type"))

new_column_order_test <- c(column_sorting)

position_type_district_test_bag <- district_personnel_position_type_test_bag[, new_column_order_test]



# Make predictions on the test data
rf_position_type_predictions_district <- predict(model_position,
                                                 as.matrix(position_type_district_test_bag))

#turning the predictions into a probability dataframe for each row in original data
probability_matrix_district_position_type <- data.frame(
  row_number = rep(c(1:nrow(district_personnel_position_type)),
                   each = 25),
  C_position_type = rep(c(0:24), nrow(district_personnel_position_type)),
  confidence = rf_position_type_predictions_district)

#shifting back to character values for position type
probability_matrix_district_position_type <-
  turn_position_type_into_values(probability_matrix_district_position_type)

#filtering to row with highest prediction value
probability_matrix_district_position_type <- probability_matrix_district_position_type %>% 
  group_by(row_number) %>% 
  filter(confidence == max(confidence))

#adding prediction and confidence level to original data
district_personnel_position_type <- district_personnel_position_type %>% 
  ungroup() %>% 
  mutate(predicted_value = probability_matrix_district_position_type$C_position_type,
         confidence = probability_matrix_district_position_type$confidence)

## Cleaning up the original data
district_data_export_position <- prediction_data_prep_exp(district_personnel)

district_data_export_position <- district_data_export_position %>% 
  left_join(district_personnel_position_type %>% select(district_data_concat, predicted_value, confidence),
            by = "district_data_concat")


corrected_predictions <- district_data_export_position$predicted_value

corrected_predictions[corrected_predictions == "Teacher"] <- "temp_teacher"
corrected_predictions[corrected_predictions == "Other"] <- "Teacher"
corrected_predictions[corrected_predictions == "Substitute"] <- "Other"
corrected_predictions[corrected_predictions == "temp_teacher"] <- "Substitute"

district_data_export_position$predicted_value <- corrected_predictions

```

## 8.4 Train Function Model

```{r}

district_personnel_function <- prediction_data_prep(district_personnel,
                                                         "C_function")

#making all district dataframes a single row per combination of values
district_personnel_function <- district_personnel_function %>% 
  group_by(district_data_concat) %>% 
  summarize()

# Prep the bags
district_personnel_function_test_bag <- bag_prep(district_personnel_function)

# Clean abreviations and columns
district_personnel_function_test_bag <- add_all_cols(district_personnel_function_test_bag)
district_personnel_function_test_bag <- fix_abbrevs(district_personnel_function_test_bag)

district_personnel_function_test_bag <- align_bags(training_bag_columns_function,
                                                        district_personnel_function_test_bag)
# Identify and add missing columns in the test data
district_personnel_function_test_bag <- add_missing_cols(training_bag_columns_function,                              district_personnel_function_test_bag)


#setting seed for replicability
set.seed(223)

#ordering columns the exact same order across training and test

  ## Get the column names except "C_position_type"
column_sorting <- sort(setdiff(training_bag_columns_function$'.', "C_function"))

new_column_order_test <- c(column_sorting)

function_district_test_bag <- district_personnel_function_test_bag[, new_column_order_test]


# Make predictions on the test data
rf_function_predictions_district <- predict(model_function,
                                                 as.matrix(function_district_test_bag))

#turning the predictions into a probability dataframe for each row in original data
probability_matrix_district_function <- data.frame(
  row_number = rep(c(1:nrow(district_personnel_function)),
                   each = 35),
  C_function = rep(c(0:34), nrow(district_personnel_function)),
  confidence = rf_function_predictions_district)

#shifting back to character values for position type
probability_matrix_district_function <-
  turn_function_into_values(probability_matrix_district_function)

#filtering to row with highest prediction value
probability_matrix_district_function <- probability_matrix_district_function %>% 
  group_by(row_number) %>% 
  filter(confidence == max(confidence))

#adding prediction and confidence level to original data
district_personnel_function <- district_personnel_function %>% 
  ungroup() %>% 
  mutate(predicted_value = probability_matrix_district_function$C_function,
         confidence = probability_matrix_district_function$confidence)


district_data_export_function <- prediction_data_prep_exp(district_personnel)

district_data_export_function <- district_data_export_function %>% 
  left_join(district_personnel_function %>% select(district_data_concat, predicted_value, confidence),
            by = "district_data_concat")

```

## 8.5 Merging predictions together

```{r}

## Change back any columns you changed in 8.1

#district_data_export_function <- district_data_export_function %>%
#  rename(`D_dept_desc` = `C_reporting`, `D_position_title_desc` = `D_position_title`)

#district_data_export_position <- district_data_export_position %>%
#  rename(`D_dept_desc` = `C_reporting`, `D_position_title_desc` = `D_position_title`)

shared_cols <- intersect(names(budgeted_positions_raw), names(district_data_export_position))
shared_cols <- intersect(shared_cols, names(district_data_export_function))


budgeted_positions_raw <- budgeted_positions_raw %>%
  # join position predictions
  left_join(
    district_data_export_position %>%
      select(all_of(shared_cols), ML_predicted_position_type = predicted_value),
    by = shared_cols
  ) %>%
  # join function predictions
  left_join(
    district_data_export_function %>%
      select(all_of(shared_cols), ML_predicted_function = predicted_value),
    by = shared_cols
  )


```

# Part 9: Upload File to Sharepoint

## Part 9.1 Any final clean up

Add any final clean up to your file below (remove unncessary columns, update column names to be standard etc.)

```{r}
processed_data_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2025%20Cohort/1.%20Data%20&%20Analysis%20-%20Secure/1.%20Data%20Processed"

#budgeted positions
ers_write_sharepoint(data = budgeted_positions_raw,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = paste0(district_folder,"_budgeted_positions_processed.csv"))


#Schools File
ers_write_sharepoint(data = school_raw,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = paste0(district_folder,"_schools_processed.csv"))

#location crosswalk
ers_write_sharepoint(data = location_crosswalk,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = paste0(district_folder,"_location_crosswalk.csv"))
```