---
title: "01_bb_validation"
author: "Jenny Katz"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3   # Customize the depth of the table of contents for HTML document
    toc_float:
      collapsed: true    # Start with the table of contents expanded
      smooth_scroll: true  # Enable smooth scrolling
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "knitted files", output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 0: Read Libraries and connect to Sharepoint

```{r}
erstools::ers_load_packages()
options(scipen = 99) #prevent any numbers from showing up as scientific notation
```

```{r}
#set sharepoint site URL for access to our org-wide files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"

#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)

#bring in pub available data
nces_data <- read_excel("NCES Data for BB Districts.xlsx")
```

# Part 1: Update Variables

```{r}

#1. folder path to data
raw_data_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2025%20Cohort/1.%20Data%20&%20Analysis%20-%20Secure/0.%20Data%20Received/"

#2. district folder
district_folder <-"Hamilton"


#3. file names 
budgeted_positions_file <- "HCSFY26PositionBudgetFile.xlsx"
school_file <- "District - Demographic Breakdown by School - Current Enrollment (3).xlsx"

#4.District & Project INfo
LEAID = "4701590" #Open up nces_data to find
project_year = "2025"
project = "Budget Benchmarking 2025"

```

# Part 2: Bring in Data

```{r}
#4.Bring in files 
#budgeted positions
budgeted_positions_raw <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = budgeted_positions_file)

#rules 
rules_db <-ers_read_sharepoint(folder_path = "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/1.%20BB%20Resources%20&%20Admin/3.%20BB%20Methodology%20&%20Guidance",
  file_name_with_extension =  "Budget Benchmarking Coding Rules Database.xlsx",
  sheet_name = district_folder,
  skip_rows = 4)
```

```{r}
# Loading in crosswalks


dept_crosswalk <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = "HCS-Department.xlsx")

object_crosswalk <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = "HCS-Object.xlsx")


```

```{r}
budgeted_positions_raw <- budgeted_positions_raw %>%
  separate(`Account Details`, into = c("Fund Code", "Agency Code", "Department Code", "Location Code", "Level Code", "Function Code", "Object Code", "Future Code", "Grant Code", "Project"), sep = "-")


budgeted_positions_raw <- budgeted_positions_raw %>% 
  select(-`Location Code`, -`Function Code`)


# Making manual crosswalks for the few codes we have

budgeted_positions_raw <- budgeted_positions_raw %>% 
  mutate(Fund_Description = case_when(
    `Fund Code` == 000 ~ "Trustee",
    `Fund Code` == 005 ~ "Payroll",
    `Fund Code` == 141 ~ "General Purpose",
    `Fund Code` == 143 ~ "School Nutrition",
    `Fund Code` == 146 ~ "Internal Service",
    `Fund Code` == 147 ~ "Capital Maintenance",
    `Fund Code` == 900 ~ "Capital Assets",
  ))


budgeted_positions_raw <- budgeted_positions_raw %>% 
  mutate(Agency_Description = case_when(
    `Agency Code` == 0 ~ "Any fund other than General Purpose (141)",
    `Agency Code` == 1 ~ "General Purpose/State Grants",
    `Agency Code` == 2 ~ "Federal Grants",
    `Agency Code` == 5 ~ "Self-Funded Grants",
  ))


budgeted_positions_raw <- budgeted_positions_raw %>% 
  mutate(Level_Description = case_when(
    `Level Code` == 0 ~ "None",
    `Level Code` == 1 ~ "Elementary",
    `Level Code` == 2 ~ "Middle",
    `Level Code` == 3 ~ "High",
    `Level Code` == 4 ~ "Combo Schools",
    `Level Code` == 5 ~ "Private Schools",
    `Level Code` == 6 ~ "Preschool",
    `Level Code` == 7 ~ "Programs",
    `Level Code` == 8 ~ "Charter School",
  ))


```

```{r}

budgeted_positions_raw$Object_Description <- object_crosswalk$Description[match(budgeted_positions_raw$`Object Code`, object_crosswalk$Object)]


budgeted_positions_raw$Department_Description <- dept_crosswalk$Description[match(budgeted_positions_raw$`Department Code`, dept_crosswalk$Code)]

```

# Part 3: Clean Up Budgeted Positions

## 3.0 Column Mapping

```{r}

#    # Keep ERS names exactly as is - fill with NAs if missing
budgeted_positions_raw <- budgeted_positions_raw %>% 
  mutate(

    #Project INfo
    H_record_ID = row_number(),
    H_district = district_folder,
    H_LEAID = LEAID,
    H_project_year = project_year,
    H_project = project,
    
    #FTE
    D_employee_id = `Employee Number`, #critical
    D_empl_fte = `Bud FTE`, #critical
   
    
    #account codes
    D_activity_code     = `Agency Code`,
    D_allocation_code = `Level Code`,
    D_class_code   = NA_character_,
    D_cost_center_code  = Location,
    D_fund_code    = `Fund Code`, #critical
    D_funct_code   = `Function`,
    D_dept_code         = `Department Code`,
    # To match what their account book calls location
    D_location_code      = `GL Loc`, #critical
    D_position_type_code    = NA_character_,
    D_position_title_code    = `Position`,
    D_object_code  = `Object Code`,
    D_program_code = `Grant Code`,
    D_project_code = `Project`,
    
    #account descriptions
    D_activity_desc     = Agency_Description,
    D_allocation_desc  = Level_Description,
    D_class_desc   = `Position Group`,
    D_cost_center_desc  = `Location Desc`,
    D_fund_desc    = `Fund_Description`, #critical
    D_funct_desc   = `Function Desc`,
    D_dept_desc    = `Department_Description`, 
    D_location_desc = `GL Loc Description`, #critical
    D_position_type_desc     = `Type`, 
    D_position_title_desc    = `Position Desc`, #critical
    D_object_desc  = Object_Description,
    D_program_desc = NA_character_,
    D_project_desc = NA_character_,
  
    
   #other disrict specific fields to add... update as needed
    D_other_1 =  `Anc Code`,
    D_other_2 =  `Loc Bud Group`,
    D_other_3 =  `Funding Source`,
    D_other_4 =  `Future Code`,
    
     .keep = "none"
   
  ) %>%
 select(where(~ !all(is.na(.x))))

```

## 3.1 Manual Checks

-   Look at Data Validation table in OneNote and complete (right year, detailed enough, etc.)

## 3.2 Review Table Structure

```{r}
#check number of columns/rows. 
#does this seem right? 
#ALso check for missing data and data types within the file
ers_column_info(budgeted_positions_raw)
```

## 3.3. Remove Blank Rows/Columns

This code block deletes any rows or columns made up fully of blank values ("", NA," ", etc.). These rows and columns wouldn't give you any additional info so it's best practice to take them out early on.

```{r}

#pre-check
ncol(budgeted_positions_raw)

#remove empty rows and columns
budgeted_positions_raw <- remove_empty(budgeted_positions_raw,
                                 which = c("rows", "cols"))

#post-check
nrow(budgeted_positions_raw)
ncol(budgeted_positions_raw)

```

## 3.4 Check for Duplicates/Remove

If any rows are repetitive or not required for your analysis you can remove them. For example - you have multiple years of data but only need one.

```{r}

#Check for duplicates
budgeted_positions_raw %>%
  filter(!(is.na(D_employee_id))) %>% 
 get_dupes(-H_record_ID)

```

## 3.5 Review and modify column data types

Review the data type of each column and modify if needed. Refer to the Data Dictionary for the correct data type for each column. For more information on different data types in R and why it matters, please refer to this: <https://app.tettra.co/teams/ersknowledge/pages/32-r-data-types>

```{r}
#review data type of each column
#ALL Columns except FTE and record ID, should be charachter (this will make joining across tables easier)
col_info <- data.frame(data_type = sapply(budgeted_positions_raw, class))
print(col_info)



#change data type of columns if needed (below is an example to convert to numeric values) - common functions are as.character, as.numeric, as.integer, as.Date

#budgeted_positions_raw$D_empl_FTE <- as.numeric(budgeted_positions_raw$D_empl_FTE)

```

# Part 4: FTE Adjustments

Employees may have multiple positions or may work in multiple locations and have a FTE of 1 for each position. To get accurate FTE for each employee, project teams should adjust for this and split the FTE across positions/locations.

## 4.0 Compare FTEs to NCES

```{r}
#compare total FTE in budgeted positions with NCES data
public_data_comp_check <- nces_data %>%
  left_join((budgeted_positions_raw %>%
                group_by(H_LEAID) %>%
              summarise(M_total_district_fte = sum(D_empl_fte, na.rm = TRUE))),
              by = c("LEAID" = "H_LEAID")) %>%
  mutate(M_fte_difference = M_total_district_fte  - `Total Staff FTE`)

```

## 4.1 Investigate FTES

First investigate employees whose FTE sums over 1, so we can investigate why and make sure it's not a larger issue with data that we need to go back to district for

```{r}

#count of employees who have an sum of FTE over 1
# 1) compute per-employee adjusted FTE 
budgeted_positions_raw <- budgeted_positions_raw %>%
  group_by(D_employee_id) %>%
  mutate(H_fte_sum = sum(D_empl_fte, na.rm = TRUE)) %>%
  ungroup() 


#2) Review employees whose fte sums > 1
budgeted_positions_raw %>%
  select(D_employee_id,D_position_title_desc,D_location_desc,H_fte_sum) %>%
  filter(!is.na(D_employee_id)) %>% 
  filter(H_fte_sum >1)

#
```

## 4.2 Adjust FTEs

If issue is not that big, that we can adjust the employee FTE Down

```{r}
#NOTE:  DO NOT ADJUST VACANT FTE DOWN IF THEY HAVE THE SAME EMPLOYEE ID (NA,0, 1, blanks)

budgeted_positions_raw <- budgeted_positions_raw %>%
  group_by(D_employee_id) %>%
  mutate(
    # total FTE across same job title & ID
    total_fte = sum(D_empl_fte, na.rm = TRUE),
    # count of rows in that group
    n_rows = n(),
    # adjust only if total_fte > 1, job titles are the same,
    # and IDs are not in the excluded list
    M_fte_adjusted = case_when(
      !(is.na(D_employee_id)) &
        total_fte > 1 ~ D_empl_fte / total_fte,
      TRUE ~ D_empl_fte
    )
  ) %>%
  ungroup() %>%
  select(-H_fte_sum, -total_fte, -n_rows)

# Some checks for Indy

print(sum(budgeted_positions_raw$D_empl_fte))

print(sum(budgeted_positions_raw$M_fte_adjusted))

#2) Double Check that no adjusted FTES are greater than 1
budgeted_positions_raw %>%
  select(D_employee_id,D_position_title_desc,D_location_desc,M_fte_adjusted) %>%
  filter(M_fte_adjusted >1)

```

# Part 5: Apply Rules (If applicable)

## 5.1 Create Apply Rules Function

```{r}
# ---- deps ----
library(dplyr)
library(purrr)
library(stringr)
library(readr)
library(rlang)

# ---------- helpers ----------
normalize_val <- function(x) {
  # Uppercase + squish; preserve NA
  if (is.null(x)) return(NA_character_)
  y <- suppressWarnings(as.character(x))
  y[is.na(y)] <- NA_character_
  y <- stringr::str_squish(y)
  y[!is.na(y)] <- stringr::str_to_upper(y[!is.na(y)])
  y
}

mk_outcol <- function(ers_type) {
paste0(
"C_",
ers_type |>
str_trim() |>
 str_to_lower() |>
 str_replace_all("[^a-z0-9]+", "_") |>
 str_replace_all("^_|_$", "")
 )
}

# Treat "Unclear" as not-coded (VECTORIZED)
is_unclear <- function(x) {
  y <- tolower(trimws(as.character(x)))
  !is.na(y) & y == "unclear"   # <-- single '&' for element-wise
}


# Try to resolve rule parent names to real columns in data:
# exact -> *_desc -> *_code fallback
resolve_parent <- function(df, p) {
 p <- str_trim(p)
if (p %in% names(df)) return(p)
cand <- paste0(p, "_desc"); if (cand %in% names(df)) return(cand)
cand2 <- paste0(p, "_code"); if (cand2 %in% names(df)) return(cand2)
 p
}

# Remove C_predicted_* parents AND the aligned pattern tokens from a rule
strip_predicted <- function(parents, pattern_str) {
  parents <- str_trim(parents)

  # split pattern into same-length tokens as parents; pad if needed
  parts <- str_split(pattern_str, "\\|", n = length(parents))[[1]]
  parts <- str_trim(parts)
  if (length(parts) < length(parents)) {
    parts <- c(parts, rep(NA_character_, length(parents) - length(parts)))
  }

  drop_idx <- str_starts(parents, "C_predicted_position") |
              str_starts(parents, "C_predicted_fun")
  keep_idx <- !drop_idx

  if (!any(keep_idx)) {
    return(list(parents = character(0), pattern = NA_character_))
  }

  parents_adj <- parents[keep_idx]
  pattern_adj <- parts[keep_idx]
  pattern_adj <- normalize_val(paste(pattern_adj[!is.na(pattern_adj)], collapse = " | "))

  list(parents = parents_adj, pattern = pattern_adj)
}

# ---------- core: apply rules for ONE ERS Coding Type ----------
apply_rules_one <- function(positions,
                            rules,
                            ers_type,
                            default_value = "Unclear") {

  out_col <- mk_outcol(ers_type)

  # base columns for pass 1 (only *_desc from the positions frame)
  base_desc_cols <- names(positions)[str_detect(names(positions), "_desc$")]

  # prep rules: split parents, strip predicted pieces, resolve aliases, flag C_ parents
  rules_sub <- rules |>
    filter(`ERS Coding Type` == ers_type) |>
    mutate(
      rule_num   = parse_number(`Rule Number`),
      parent_raw = `District Concat Account Code Parent`,
      parents0   = str_split(parent_raw, "\\|") |> map(~ str_trim(.x)),
      adj        = map2(parents0, `District Concat Account Desc`, strip_predicted),
      parents    = map(adj, "parents"),
      pattern    = map_chr(adj, "pattern")
    ) |>
    mutate(
      # resolve to actual column names in *positions*
      parents = map(parents, ~ map_chr(.x, ~ resolve_parent(positions, .x))),
      uses_c_parent = map_lgl(parents, ~ any(str_starts(.x, "C_")))
    ) |>
    filter(map_int(parents, length) > 0L, !is.na(pattern)) |>
    transmute(parents, uses_c_parent, pattern, coding = Coding, rule_num)

  # initialize output + audit if missing
  audit_rule <- paste0(".rule__", out_col)
  audit_on   <- paste0(".on__", out_col)
  audit_val  <- paste0(".val__", out_col)
  if (!out_col    %in% names(positions)) positions[[out_col]]    <- NA_character_
  if (!audit_rule %in% names(positions)) positions[[audit_rule]] <- NA_real_
  if (!audit_on   %in% names(positions)) positions[[audit_on]]   <- NA_character_
  if (!audit_val  %in% names(positions)) positions[[audit_val]]  <- NA_character_

  # ---- internal pass runner (NA-safe; uses up-to-date df INCLUDING new C_ cols) ----
  run_pass <- function(df, rules_pass, allowed_cols = names(df)) {
    if (nrow(rules_pass) == 0L) return(df)

    # keep only rules whose parents are available & allowed
    rules_ok <- rules_pass |>
      filter(map_lgl(parents, ~ all(.x %in% allowed_cols)))

    if (nrow(rules_ok) == 0L) return(df)

    # PRECOMPUTE concatenations from a LOCAL, normalized view of df,
    # restricted to only the columns we need for this pass
    parent_sets <- unique(rules_ok$parents)
    parent_key  <- map_chr(parent_sets, ~ paste(.x, collapse = " | "))

    needed_cols <- sort(unique(unlist(parent_sets, use.names = FALSE)))
    norm_local  <- df
    char_needed <- needed_cols[needed_cols %in% names(norm_local)]
    norm_local[char_needed] <- lapply(norm_local[char_needed], normalize_val)

    row_parent_value <- function(norm_df, parents) {
      parents <- str_trim(parents)
      stopifnot(all(parents %in% names(norm_df)))
      if (length(parents) == 1L) {
        norm_df[[parents[1]]]
      } else {
        do.call(paste, c(norm_df[parents], sep = " | "))
      }
    }

    parent_map  <- map(parent_sets, ~ row_parent_value(norm_local, .x)) |>
      set_names(parent_key)

    hits <- pmap_dfr(rules_ok, function(parents, uses_c_parent, pattern, coding, rule_num) {
      key <- paste(parents, collapse = " | ")
      pv_norm <- parent_map[[key]]                  # already normalized
      matched <- !is.na(pv_norm) & !is.na(pattern) & (pv_norm == pattern)
      if (!any(matched, na.rm = TRUE)) {
        return(tibble(.row = integer(), coding = character(),
                      rule_num = double(), on = character(), val = character()))
      }
      # Drop "Unclear" codings — they do NOT count as a fill
      if (is_unclear(coding)) {
        return(tibble(.row = integer(), coding = character(),
                      rule_num = double(), on = character(), val = character()))
      }
      tibble(
        .row = which(matched),
        coding = coding,
        rule_num = rule_num,
        on = key,
        val = pattern
      )
    })

    if (nrow(hits) == 0L) return(df)

    # choose lowest rule_num per row; only fill where currently NA or "Unclear"
    choose <- hits |>
      group_by(.row) |>
      slice_min(rule_num, n = 1, with_ties = FALSE) |>
      ungroup()

    to_fill <- which(is.na(df[[out_col]]) | is_unclear(df[[out_col]]))
    choose  <- choose |> filter(.row %in% to_fill)

    if (nrow(choose)) {
      df[[out_col]][choose$.row]    <- choose$coding
      df[[audit_rule]][choose$.row] <- choose$rule_num
      df[[audit_on]][choose$.row]   <- choose$on
      df[[audit_val]][choose$.row]  <- choose$val
    }
    df
  }
  
  # PASS 2: rules that depend on coded C_ parents (now any existing columns allowed)
  positions <- run_pass(
    positions,
    rules_sub |> filter(uses_c_parent),
    allowed_cols = names(positions)
  )

  # PASS 1: only non-C_ parents AND restricted to *_desc columns
  positions <- run_pass(
    positions,
    rules_sub |> filter(!uses_c_parent),
    allowed_cols = base_desc_cols
  )

  

  # default any remaining NAs
  positions[[out_col]] <- coalesce(positions[[out_col]], default_value)
  positions
}

# ---------- run ALL ERS Coding Types in your specified order ----------
apply_rules_all <- function(positions,
                            rules,
                            default_value = "Unclear",
                            preferred_order = c(
                              "Operating Status",
                              "Reporting",
                              "Position Type",
                              "Function",
                              "PreK Status",
                              "Student Type",
                              "Sharing"
                            )) {

  # all ERS Coding Types present in rules
  all_types <- unique(rules$`ERS Coding Type`)

  # case-insensitive exact match helper
  ci_match <- function(target, pool) {
    m <- match(tolower(target), tolower(pool))
    pool[!is.na(m)]
  }

  # # preferred types first (if present), then any remaining
  # ordered_pref <- unlist(lapply(preferred_order, ci_match, pool = all_types), use.names = FALSE)
  # print("ordered_pref:")
  # print(ordered_pref)
  # remaining    <- setdiff(all_types, ordered_pref)
  # print("remaining")
  # print(remaining)
  # run_order    <- c(ordered_pref, remaining)
  # print("run order")
  # print(run_order)
  # #print("helloooooo")
  
  run_order <- c("Operating Status", "Reporting", "Position Type", "Function","Prek_Expense", "Student_Type", "Sharing")

  # run in order
  purrr::reduce(
    run_order,
    .init = positions,
    .f = function(df, et) apply_rules_one(df, rules, ers_type = et, default_value = default_value)
  )
}


```

## 5.2 Run Apply Rules Function

```{r}
# Example names — replace with yours:
# budgeted_positions_raw <- read_excel("Detroit_positions_25_26.xlsx")
# detroit_rules          <- read_excel("Rules_24_25.xlsx")

budgeted_positions_raw <- apply_rules_all(
  positions = budgeted_positions_raw,
  rules     = rules_db,
  default_value = "Unclear"
)

budgeted_positions_raw <-  budgeted_positions_raw %>%
rename_with(~ paste0(., "_0"), .cols = starts_with("C_"))

# #Or start with a single field while testing:
# #detroit_function <- apply_rules_one(
# #positions = detroit_position,
# rules     = rules_db,
# ers_type  = "Function",
# default_value = "Unclear"
# )

```

## 5.3 Test Rules

```{r}
coverage_all <- budgeted_positions_raw %>%
  pivot_longer(cols = starts_with("C_"),
               names_to = "field", values_to = "val") %>%
  group_by(field) %>%
  summarise(
    total_rows   = n(), # includes Unclear + NA
    coded_rows   = sum(!is.na(val) & val != "Unclear"),
    unclear_rows = sum(val == "Unclear", na.rm = TRUE),
    na_rows      = sum(is.na(val)),
    pct_coded    = round(100 * coded_rows / total_rows, 1),
    .groups = "drop"
  ) %>%
  arrange(desc(pct_coded))

coverage_all

```

# Part 6: Machine Learning

## 6.0 Set-Up

```{r}

library(tm)
library(textstem) #for turning words into their stem form
library(randomForest) #for random forest model
library(ranger) #for random forest model
library(xgboost) #for random forest model

#set sharepoint site URL for access to our org-wide and external files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)

client_work_drive <- orgfiles_site$get_drive("Client Work")

# Modified read_sharepoint logic for RDS files
read_sharepoint_ers_obj <- function(drive_name = client_work_drive, 
                                folder_path, 
                                file_name_with_extension, 
                                sheet_name = NULL, 
                                skip_rows = 0,
                                na_values = c("","N/A","NA","Missing")) {
  # Construct the full data path
  data_path <- file.path(folder_path, file_name_with_extension)
  
  file_extension <- paste0(".", tools::file_ext(file_name_with_extension))
  
  # Create a temporary file with the appropriate file extension
  temp_file <- tempfile(fileext = file_extension)
  
  # Download the file from the specified drive to the temporary file
  drive_name$download_file(data_path, dest = temp_file)
  
  # Read the downloaded file based on its file extension
  data_loaded <- readRDS(temp_file)
  
  # Return the loaded data
  return(data_loaded)
}



district_personnel <- budgeted_positions_raw

training_bag_columns_position <- ers_read_sharepoint(folder_path = "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/Template/BB%20Methodology%20&%20Guidance/Machine Learning Models",
  file_name_with_extension =  "position_type_training_bag_columns_9.24.24.csv")

model_position <- read_sharepoint_ers_obj(folder_path = "Other/Budget Benchmarking/Template/BB Methodology & Guidance/Machine Learning Models", file_name_with_extension =  "position_type_model_9.24.24.rds")


training_bag_columns_function <- ers_read_sharepoint(folder_path = "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/Template/BB%20Methodology%20&%20Guidance/Machine Learning Models",
  file_name_with_extension =  "function_training_bag_columns_9.24.24.csv")

model_function <- read_sharepoint_ers_obj(folder_path = "Other/Budget Benchmarking/Template/BB Methodology & Guidance/Machine Learning Models", file_name_with_extension =  "function_model_9.24.24.rds")

```

## 6.1 Modify Data

```{r}

# Add a reporting proxy. To train the ML model, we need a reporting coding, in leiu of a complete set of coding, we can use any proxy to get the job done. In Indy they had a similar column structure that we could use, otherwise you can use whether or not there is a match in the school's file as a nice proxy

district_personnel <- district_personnel %>%
  mutate(C_reporting = ifelse(D_other_2 == "Centralized Support", "Non-School", "School")) %>% 
  rename(`D_position_title` = `D_position_title_desc`)
  


```

## 6.2 Load Functions

```{r}

# All functions

#building a prefix function that allows me to add a tag into all words to mark which variable they come from
add_prefix_to_words <- function(string, prefix) {
  # Split the string into words
  words <- unlist(strsplit(string, " "))
  
  # Add prefix to each word
  prefixed_words <- paste0(prefix, words)
  
  # Rejoin the words into a single string
  prefixed_string <- paste(prefixed_words, collapse = " ")
  
  return(prefixed_string)
}

#subset to columns needed (district_data_concat is combination of all fields relevant for prediction and standardized to only regular characters)
prediction_data_prep <- function(data,
                                 predicted_value) {
  data %>%
  mutate(D_position_title = str_replace_all(D_position_title,
                                           "[^[:alnum:]]",
                                           " "),
         D_location_desc = str_replace_all(D_location_desc,
                                           "[^[:alnum:]]",
                                           " "),
         C_reporting = str_replace_all(C_reporting,
                                           "[^[:alnum:]]",
                                           " "),
         D_position_title = sapply(D_position_title,
                                 add_prefix_to_words,
                                 "pt"),
         D_location_desc = sapply(D_location_desc,
                                  add_prefix_to_words,
                                  "ld"),
         C_reporting = sapply(C_reporting,
                              add_prefix_to_words,
                              "rp"),
         district_data_concat = paste(D_position_title,
                                      D_location_desc,
                                      C_reporting,
                                      sep = " ")) %>%
  select(district_data_concat)
}


# Warnings here are fine
bag_prep <- function(data) {
  Corpus(VectorSource(data$district_data_concat)) %>%
  tm_map(content_transformer(tolower)) %>% 
  tm_map(removePunctuation) %>% 
  tm_map(removeNumbers) %>% 
  tm_map(removeWords, stopwords("english")) %>% 
  tm_map(stripWhitespace) %>% 
  DocumentTermMatrix() %>%
  as.matrix() %>% 
  as.data.frame()
}

#adding key terms if they're not already in data (so that select doesn't break)
add_each_col <- function(dataset, col_name) {
  if (!(col_name %in% names(dataset))) {
    dataset <- dataset %>% 
      mutate(!!col_name := 0)
  }
  return(dataset)
}

add_all_cols <- function(dataset) {
  dataset %>% 
  add_each_col("pttchr") %>% 
  add_each_col("pttch") %>% 
  add_each_col("ptasst") %>% 
  add_each_col("ptpara") %>% 
  add_each_col("ptsvcs") %>% 
  add_each_col("ptmgr") %>% 
  add_each_col("ptcoor") %>% 
  add_each_col("ptcoord") %>% 
  add_each_col("ptteacher") %>% 
  add_each_col("ptassistant") %>% 
  add_each_col("ptparaprofessional") %>% 
  add_each_col("ptservices") %>% 
  add_each_col("ptmanager") %>% 
  add_each_col("ptcoordinator")
}

fix_abbrevs <- function(data) {
  data %>% 
  mutate(ptteacher = ifelse(pttchr == 1 | pttch == 1 | ptteacher == 1, 1, 0),
         ptassistant = ifelse(ptasst == 1 | ptassistant == 1, 1, 0),
         ptparaprofessional = ifelse(ptpara ==1 | ptparaprofessional ==1, 1, 0),
         ptservices = ifelse(ptservices == 1 | ptsvcs == 1, 1, 0),
         ptmanager = ifelse(ptmanager == 1 | ptmgr == 1, 1, 0),
         ptcoordinator=ifelse(ptcoord==1|ptcoor==1|ptcoordinator==1, 1, 0)) %>% 
  select(-c(pttchr,
            pttch,
            ptasst,
            ptpara,
            ptsvcs,
            ptmgr,
            ptcoord,
            ptcoor))
}


#removing words that weren't present in training data

align_bags <- function(list_of_cols, test_data) {

  non_training_words <- setdiff(colnames(test_data),
                                list_of_cols$`.`)

  adjusted_test_data <- test_data %>%
    select(-any_of(non_training_words))

  return(adjusted_test_data)
}

add_missing_cols <- function(list_of_cols, test_df) {

  missing_columns <- setdiff(list_of_cols$`.`,
                             names(test_df))

  zero_values_df <- as_tibble(
    matrix(0, ncol = length(missing_columns),
           nrow = nrow(test_df),
           dimnames = list(NULL, missing_columns))
  )

  adjusted_test_df <- bind_cols(test_df,
                                zero_values_df)

  return(adjusted_test_df)
}


prediction_data_prep_exp <- function(data) {
  data %>%
  mutate(D_position_title_h = str_replace_all(D_position_title,
                                           "[^[:alnum:]]",
                                           " "),
         D_location_desc_h = str_replace_all(D_location_desc,
                                           "[^[:alnum:]]",
                                           " "),
         C_reporting_h = str_replace_all(C_reporting,
                                           "[^[:alnum:]]",
                                           " "),
         D_position_title_h = sapply(D_position_title_h,
                                 add_prefix_to_words,
                                 "pt"),
         D_location_desc_h = sapply(D_location_desc_h,
                                  add_prefix_to_words,
                                  "ld"),
         C_reporting_h = sapply(C_reporting_h,
                              add_prefix_to_words,
                              "rp"),
         district_data_concat = paste(D_position_title_h,
                                      D_location_desc_h,
                                      C_reporting_h,
                                      sep = " ")) %>% 
    select(-c(D_location_desc_h, D_position_title_h, C_reporting_h))
}




```

```{r}

# ERS content functions


turn_position_type_into_values <- function(dataset) {
  dataset %>% 
    mutate(C_position_type = case_when(
      C_position_type == 0 ~ "Substitute",
      C_position_type == 1 ~ "Teacher",
      C_position_type == 2 ~ "Other",
      C_position_type == 3 ~ "Nurse",
      C_position_type == 4 ~ "TA",
      C_position_type == 5 ~ "Sec/Clerk/Other Admin",
      C_position_type == 6 ~ "Coordinator/Manager",
      C_position_type == 7 ~ "Principal",
      C_position_type == 8 ~ "Speech Therapist",
      C_position_type == 9 ~ "Guidance Counselor",
      C_position_type == 10 ~ "Custodian",
      C_position_type == 11 ~ "Psychologist",
      C_position_type == 12 ~ "Instructional Coach",
      C_position_type == 13 ~ "Vice Principal",
      C_position_type == 14 ~ "Social Worker",
      C_position_type == 15 ~ "School Monitor/Security",
      C_position_type == 16 ~ "Club Advisor/Coach",
      C_position_type == 17 ~ "Librarian",
      C_position_type == 18 ~ "(Exec) Director",
      C_position_type == 19 ~ "Non-Position",
      C_position_type == 20 ~ "Nurse Aide",
      C_position_type == 21 ~ "Area Officers",
      C_position_type == 22 ~ "Occupational Therapist",
      C_position_type == 23 ~ "Physical Therapist",
      C_position_type == 24 ~ "Instructional Coach Supervisors",
      TRUE ~ NA
    ))
}


turn_function_into_values <- function(dataset) {
  dataset %>% 
    mutate(C_function = case_when(
      C_function == 0 ~ "Other Non-Compensation",
      C_function == 1 ~ "Teacher Compensation",
      C_function == 2 ~ "Aides Compensation",
      C_function == 3 ~ "Substitute Compensation",
      C_function == 4 ~ "Library & Media",
      C_function == 5 ~ "Instructional Materials & Supplies",
      C_function == 6 ~ "Other Compensation",
      C_function == 7 ~ "Extended Time & Tutoring",
      C_function == 8 ~ "Enrichment",
      C_function == 9 ~ "Social & Emotional",
      C_function == 10 ~ "Physical Health & Services",
      C_function == 11 ~ "Career & Academic Counseling",
      C_function == 12 ~ "Parent & Community Relations",
      C_function == 13 ~ "Professional Growth",
      C_function == 14 ~ "Curriculum Development",
      C_function == 15 ~ "Recruitment",
      C_function == 16 ~ "Special Population Program Management & Support",
      C_function == 17 ~ "Governance",
      C_function == 18 ~ "School Supervision",
      C_function == 19 ~ "School Administration",
      C_function == 20 ~ "Research & Accountability",
      C_function == 21 ~ "Communications",
      C_function == 22 ~ "Student Assignment",
      C_function == 23 ~ "Facilities & Maintenance",
      C_function == 24 ~ "Security & Safety",
      C_function == 25 ~ "Food Services",
      C_function == 26 ~ "Student Transportation",
      C_function == 27 ~ "Utilities",
      C_function == 28 ~ "Human Resources",
      C_function == 29 ~ "Finance, Budget, Purchasing & Distribution",
      C_function == 30 ~ "Data Processing & Information Services",
      C_function == 31 ~ "Facilities Planning",
      C_function == 32 ~ "Development & Fundraising",
      C_function == 33 ~ "Legal",
      C_function == 34 ~ "Insurance",
      TRUE ~ "Other Non-Compensation"))
}


```

## 6.3 Train Position Type Model

```{r}


district_personnel_position_type <- prediction_data_prep(district_personnel,
                                                         "C_position_type")

#making all district dataframes a single row per combination of values
district_personnel_position_type <- district_personnel_position_type %>% 
  group_by(district_data_concat) %>% 
  summarize()

# Prep the bags
district_personnel_position_type_test_bag <- bag_prep(district_personnel_position_type)

# Clean abreviations and columns
district_personnel_position_type_test_bag <- add_all_cols(district_personnel_position_type_test_bag)
district_personnel_position_type_test_bag <- fix_abbrevs(district_personnel_position_type_test_bag)


district_personnel_position_type_test_bag <- align_bags(training_bag_columns_position,
                                                        district_personnel_position_type_test_bag)

# Identify and add missing columns in the test data
district_personnel_position_type_test_bag <- add_missing_cols(training_bag_columns_position,                              district_personnel_position_type_test_bag)

set.seed(223)

#ordering columns the exact same order across training and test

  ## Get the column names except "C_position_type"
column_sorting <- sort(setdiff(training_bag_columns_position$'.', "C_position_type"))

new_column_order_test <- c(column_sorting)

position_type_district_test_bag <- district_personnel_position_type_test_bag[, new_column_order_test]



# Make predictions on the test data
rf_position_type_predictions_district <- predict(model_position,
                                                 as.matrix(position_type_district_test_bag))

#turning the predictions into a probability dataframe for each row in original data
probability_matrix_district_position_type <- data.frame(
  row_number = rep(c(1:nrow(district_personnel_position_type)),
                   each = 25),
  C_position_type = rep(c(0:24), nrow(district_personnel_position_type)),
  confidence = rf_position_type_predictions_district)

#shifting back to character values for position type
probability_matrix_district_position_type <-
  turn_position_type_into_values(probability_matrix_district_position_type)

#filtering to row with highest prediction value
probability_matrix_district_position_type <- probability_matrix_district_position_type %>% 
  group_by(row_number) %>% 
  filter(confidence == max(confidence))

#adding prediction and confidence level to original data
district_personnel_position_type <- district_personnel_position_type %>% 
  ungroup() %>% 
  mutate(predicted_value = probability_matrix_district_position_type$C_position_type,
         confidence = probability_matrix_district_position_type$confidence)

## Cleaning up the original data
district_data_export_position <- prediction_data_prep_exp(district_personnel)

district_data_export_position <- district_data_export_position %>% 
  left_join(district_personnel_position_type %>% select(district_data_concat, predicted_value, confidence),
            by = "district_data_concat")


corrected_predictions <- district_data_export_position$predicted_value

corrected_predictions[corrected_predictions == "Teacher"] <- "temp_teacher"
corrected_predictions[corrected_predictions == "Other"] <- "Teacher"
corrected_predictions[corrected_predictions == "Substitute"] <- "Other"
corrected_predictions[corrected_predictions == "temp_teacher"] <- "Substitute"

district_data_export_position$predicted_value <- corrected_predictions

```

## 6.4 Train Function Model

```{r}

district_personnel_function <- prediction_data_prep(district_personnel,
                                                         "C_function")

#making all district dataframes a single row per combination of values
district_personnel_function <- district_personnel_function %>% 
  group_by(district_data_concat) %>% 
  summarize()

# Prep the bags
district_personnel_function_test_bag <- bag_prep(district_personnel_function)

# Clean abreviations and columns
district_personnel_function_test_bag <- add_all_cols(district_personnel_function_test_bag)
district_personnel_function_test_bag <- fix_abbrevs(district_personnel_function_test_bag)

district_personnel_function_test_bag <- align_bags(training_bag_columns_function,
                                                        district_personnel_function_test_bag)
# Identify and add missing columns in the test data
district_personnel_function_test_bag <- add_missing_cols(training_bag_columns_function,                              district_personnel_function_test_bag)


#setting seed for replicability
set.seed(223)

#ordering columns the exact same order across training and test

  ## Get the column names except "C_position_type"
column_sorting <- sort(setdiff(training_bag_columns_function$'.', "C_function"))

new_column_order_test <- c(column_sorting)

function_district_test_bag <- district_personnel_function_test_bag[, new_column_order_test]


# Make predictions on the test data
rf_function_predictions_district <- predict(model_function,
                                                 as.matrix(function_district_test_bag))

#turning the predictions into a probability dataframe for each row in original data
probability_matrix_district_function <- data.frame(
  row_number = rep(c(1:nrow(district_personnel_function)),
                   each = 35),
  C_function = rep(c(0:34), nrow(district_personnel_function)),
  confidence = rf_function_predictions_district)

#shifting back to character values for position type
probability_matrix_district_function <-
  turn_function_into_values(probability_matrix_district_function)

#filtering to row with highest prediction value
probability_matrix_district_function <- probability_matrix_district_function %>% 
  group_by(row_number) %>% 
  filter(confidence == max(confidence))

#adding prediction and confidence level to original data
district_personnel_function <- district_personnel_function %>% 
  ungroup() %>% 
  mutate(predicted_value = probability_matrix_district_function$C_function,
         confidence = probability_matrix_district_function$confidence)


district_data_export_function <- prediction_data_prep_exp(district_personnel)

district_data_export_function <- district_data_export_function %>% 
  left_join(district_personnel_function %>% select(district_data_concat, predicted_value, confidence),
            by = "district_data_concat")

```

## 6.5 Merging predictions together

```{r}

## Change back any columns you changed in 8.1

# # INDy Specific
# district_data_export_function <- district_data_export_function %>%
#   rename(`C_reporting_0` = `C_reporting`, `D_position_title_desc` = `D_position_title`)
# 
# district_data_export_position <- district_data_export_position %>%
#   rename(`C_reporting_0` = `C_reporting`, `D_position_title_desc` = `D_position_title`)

shared_cols <- intersect(names(budgeted_positions_raw), names(district_data_export_position))
shared_cols <- intersect(shared_cols, names(district_data_export_function))


budgeted_positions_raw <- budgeted_positions_raw %>%
  # join position predictions
  left_join(
    district_data_export_position %>%
      select(all_of(shared_cols), ML_predicted_position_type = predicted_value),
    by = shared_cols
  ) %>%
  # join function predictions
  left_join(
    district_data_export_function %>%
      select(all_of(shared_cols), ML_predicted_function = predicted_value),
    by = shared_cols
  )


```

# Part 7: Upload Budget File to Sharepoint

Add any final clean up to your file below (remove unnecessary columns, update column names to be standard etc.)

```{r}
processed_data_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2025%20Cohort/1.%20Data%20&%20Analysis%20-%20Secure/1.%20Data%20Processed"

#budgeted positions
ers_write_sharepoint(data = budgeted_positions_raw,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = paste0(district_folder,"_budgeted_positions_processed.csv"))

```

# Part 8. Clean up Schools file

## 8.1 Bring in School Data

```{r}

#schools files
school_raw <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = school_file, sheet_name = "Consolidated")



school_raw <- school_raw %>% 
  filter(`School Name` != "Overall - Total")


```

## 8.2 Map District Col to ERS Columns

```{r}
#changing column names and removing unneeded columns
#replace the column names after the "=" with column names from your data.You can press tab after the = to select column names available in the data

school_raw <- school_raw %>% 
  mutate(
    
    #Project INfo
    H_record_ID = row_number(),
    H_district = district_folder,
    H_LEAID = LEAID,
    H_project_year = project_year,
    H_project = project,
    
    #school info
    D_location_code = `School ID`,
    D_location_desc = `School Name`,
    D_school_level = `Model Type`,
    D_school_type = `Program Status`,
    D_grade_span = `Grade Levels`,
    D_enrollment_total = `Student Count`,
    D_preK_enrollment = 0,
    D_ell_total = `ELL Count`,
    D_frl_total = `ED Count`,
    D_swd_total = `SPED Count`,
    # Need to update self contained data
    D_self_contained = 0,
    
    
     .keep = "none"
  )
  

```

## 8.3 Update Data Type

```{r}
#review data type of each column
#ALL Columns except FTE and record ID, should be charachter (this will make joining across tables easier)
col_info <- data.frame(data_type = sapply(school_raw, class))
print(col_info)



#change data type of columns if needed (below is an example to convert to numeric values) - common functions are as.character, as.numeric, as.integer, as.Date

#budgeted_positions_raw$D_empl_FTE <- as.numeric(budgeted_positions_raw$D_empl_FTE)
```

## 8.4 Review School Structure

```{r}
#check number of columns/rows. 
#does this seem right? 
#ALso check for missing data and data types within the file

# 2 that are missing FRL data are excluded anyways
ers_column_info(school_raw)
```

## 8.5 Check /Remove Duplicate School

### a. Investigate Dups

```{r}

#Check for duplicates
school_raw %>%
 get_dupes(-H_record_ID)


#deuplicate
school_raw<- school_raw %>%
  distinct()

#find shools with the same school ID
duplicates_by_id <- school_raw  %>%
  get_dupes(D_location_code)

```

### b. Remove Duplicates

```{r}
#NOTE THIS WAS A DETROIT SPECIFIC DUP REMOVAL. MAY BE DIFF FOR YOUR DISTRICT

#DETROIT ONLY: update the school IDs so they combine into one row.. 

# No duplicates

```

## 8.6 Add Coding Columns

### a. School Level

```{r}
#view school level
school_raw %>%
  group_by(D_grade_span) %>%
  count()

#Update for your district's values 
school_raw <-school_raw %>%
  mutate(C_school_level = case_when(D_grade_span == "6th - 12th"	~ "HS",
                                    D_grade_span == "6th - 8th"	~ "MS",
                                    D_grade_span == "9th - 12th"	~ "HS",
                                    D_grade_span == "K - 12th" ~ "Other",
                                    D_grade_span == "K-5" ~ "ES",
                                    D_grade_span == "K-8th" ~ "ES",
                                    TRUE ~ "Exclude"))

#check coding
school_raw %>%
  group_by(D_grade_span,C_school_level) %>%
  count()

```

### b. School Include Flag

```{r}
#UPDATE BASED on DISTRICT'S ALT SCHOOLS
school_raw <- school_raw %>%
  mutate(
    C_school_flag = case_when(
      D_school_level == "Exclude" ~ "Exclude",
      D_location_desc == "Hamilton County Virtual School" ~ "Include: District-Wide Only",
      TRUE ~ "Include: All"
    ),
    C_school_flag_reason = case_when(
      D_school_level == "Exclude" ~ "Charter/Out of District",
      D_location_desc == "Hamilton County Virtual School" ~ "Virtual School",
      TRUE ~ "Traditional/Neighborhood"))




#check coding
school_raw %>%
  group_by(C_school_flag, C_school_flag_reason) %>%
  count()

```

### c. Enrollment Characteristics

### c.1 Calc Self-contained based on 2024 data (NOTE EVENTUALLY UPDATE)

```{r}
#NEED TO UPDATE WITH REAL NUMBERS EVENTUALLY
#last year % of students that was self-contained was 1.9% so going to calculate SC number at each school 
school_raw <- school_raw %>%
  mutate(D_self_contained = round(D_enrollment_total * 0.019,0))
```

```{r}

#Update for your district's values 
school_raw <-school_raw %>%
  mutate(C_school_level_mapped = case_when(C_school_level == "ES"	~ "ES",
                                    C_school_level == "MS"	~ "SS",
                                    C_school_level == "HS"	~ "SS",
                                    TRUE ~ "Other"))

school_raw %>%
  group_by(C_school_level,C_school_level_mapped) %>%
  count()

school_raw <- school_raw %>% 
  mutate(C_school_size = ifelse(C_school_level == "ES", ifelse(D_enrollment_total < 350, "Small", "Not Small"), ifelse(D_enrollment_total < 500, "Small", "Not Small")))


school_raw <- school_raw %>%
  mutate(C_gen_ed = D_enrollment_total - D_self_contained,
         M_ell_pct = D_ell_total / D_enrollment_total,
         M_swd_pct = as.numeric(D_swd_total) / D_enrollment_total,
         M_frl_pct = D_frl_total / D_enrollment_total)

# Adding in sanity cleanups - you should manually check any huge variations
school_raw <- school_raw %>%
  mutate(M_ell_pct = ifelse(is.na(M_ell_pct), 0, ifelse(M_ell_pct > 1, 1, M_ell_pct)),
         M_swd_pct = ifelse(is.na(M_swd_pct), 0, ifelse(M_swd_pct > 1, 1, M_swd_pct)),
         M_frl_pct = ifelse(is.na(M_frl_pct), 0, ifelse(M_frl_pct > 1, 1, M_frl_pct)))


```

```{r}

# Create all levels quartiles flags
school_raw <- school_raw %>%
  group_by(C_school_flag) %>%
  mutate(
    M_swd_quartile_flag = case_when(
      M_swd_pct <= quantile(M_swd_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_swd_pct > quantile(M_swd_pct, 0.25, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.5, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag) %>%
  mutate(
    M_ell_quartile_flag = case_when(
      M_ell_pct <= quantile(M_ell_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_ell_pct > quantile(M_ell_pct, 0.25, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.5, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag) %>%
  mutate(
    M_frl_quartile_flag = case_when(
      M_frl_pct <= quantile(M_frl_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_frl_pct > quantile(M_frl_pct, 0.25, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.5, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


```

```{r}

# Create all levels quartiles flags
school_raw <- school_raw %>%
  group_by(C_school_flag, C_school_level_mapped) %>%
  mutate(
    M_swd_quartile_flag_sized = case_when(
      M_swd_pct <= quantile(M_swd_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_swd_pct > quantile(M_swd_pct, 0.25, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.5, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag, C_school_level_mapped) %>%
  mutate(
    M_ell_quartile_flag_sized = case_when(
      M_ell_pct <= quantile(M_ell_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_ell_pct > quantile(M_ell_pct, 0.25, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.5, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag, C_school_level_mapped) %>%
  mutate(
    M_frl_quartile_flag_sized = case_when(
      M_frl_pct <= quantile(M_frl_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_frl_pct > quantile(M_frl_pct, 0.25, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.5, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


```

```{r}

# Validating the quartiles
make_quartile_summary <- function(data, var, flag_var) {
  cutoffs <- quantile(data[[var]], probs = c(0.25, 0.5, 0.75), na.rm = TRUE)
  
  counts <- data %>%
    count(!!sym(flag_var)) %>%
    rename(quartile = !!sym(flag_var), count = n)
  
  tibble(
    metric = var,
    q25 = cutoffs[1],
    q50 = cutoffs[2],
    q75 = cutoffs[3]
  ) %>%
    crossing(counts)
}

# build tables for each metric (only looking at traditional campuses)

## Switch filters to small, not-small as needed
data_testing <- school_raw %>% 
  filter(C_school_flag == "Include: All")

frl_summary <- make_quartile_summary(data_testing, "M_frl_pct", "M_frl_quartile_flag")
swd_summary <- make_quartile_summary(data_testing, "M_swd_pct", "M_swd_quartile_flag")
ell_summary <- make_quartile_summary(data_testing, "M_ell_pct", "M_ell_quartile_flag")

# combine into one concise table
quartile_summary <- bind_rows(frl_summary, swd_summary, ell_summary)

print(quartile_summary)

```

## 8.7 Compare Enrollment to NCES

```{r}

public_data_comp_check <- public_data_comp_check %>%
  left_join(
    (school_raw %>%
       group_by(H_LEAID) %>%
       summarise(M_total_district_enrollment = sum(D_enrollment_total, na.rm = TRUE))),
    by = c("LEAID" = "H_LEAID")
  ) %>%
  mutate(
    M_enrollment_difference = M_total_district_enrollment - `Total Enrollment`
  )
```

# Part 8: Position + Schools Crosswalk

### a. Fix mappings of schools code

```{r}

school_raw$D_location_code <- as.numeric(school_raw$D_location_code)

school_raw$D_location_code <- sprintf("%04d", school_raw$D_location_code)

```

### b. Join Position & School Tables

```{r}
#Create Crosswalk 
location_crosswalk <- bind_rows(
   # --- Part 1: Prepare the Budget Positions Table ---
 budgeted_positions_raw %>%
  filter(!is.na(D_location_code)) %>%
  count(D_location_code, D_location_desc, name = "row_count") %>% 
  mutate( table = "Position") ,

  # --- Part 2: Prepare the Schools Table ---
  school_raw %>%
    count(D_location_code, D_location_desc, name = "row_count") %>% 
    mutate(table = "School")

) %>% 
  
   #Keep only the first school name for each ID ---
  group_by(D_location_code) %>%
  arrange(desc(table)) %>%
  mutate(D_location_desc = first(D_location_desc)) %>%
  ungroup() %>%

  # --- Part 3: Pivot the Combined Table ---
  pivot_wider(
    names_from = table,
    values_from = row_count,
    values_fill = 0 
  ) %>%
  arrange(D_location_desc)


#bring in Total FTE
location_crosswalk <- location_crosswalk %>%
  left_join(
    budgeted_positions_raw %>%
      group_by(D_location_code, D_location_desc) %>%
      summarise(M_total_FTE = sum(M_fte_adjusted)) %>%
      filter(M_total_FTE > 0) %>%
      filter(!is.na(D_location_code)) %>%
      select(D_location_code, M_total_FTE),
    by = "D_location_code"
  )

#Bring in Total Enrollment 
location_crosswalk <- location_crosswalk %>%
  left_join(
    school_raw %>%
      select(D_location_code, D_enrollment_total,D_school_type),
    by = "D_location_code"
  ) %>%
  mutate(M_fte_per_500_students = M_total_FTE / D_enrollment_total * 500) %>%
  arrange(desc(School))
  

#flag issues
location_crosswalk <- location_crosswalk %>%
  mutate(
    C_location_issue_flag = case_when(
      Position < 1 & School == 1 ~ "Not In Position File",
      School == 0 &
        Position > 0 ~ "Not In School File",
      M_fte_per_500_students < 40 |
        M_fte_per_500_students > 90 ~ "Low/High FTE per 500",
      D_enrollment_total < 10 &
        D_enrollment_total > 0 ~ "Low Enrollment",
      TRUE ~ "Okay"
    )
  )
print(location_crosswalk)   
```

### c. Post Crosswalk Fixes - ID

Note this will have to be updated for each individual districts

```{r}

# to return to

```

### d. Add flags to location crosswalk

```{r}
location_crosswalk <- location_crosswalk %>%
  left_join(
    school_raw %>%
      select(D_location_code, C_school_flag,C_school_flag_reason),
    by = "D_location_code")
```

# Part 9: Upload School & Location Crosswalk Files to Sharepoint

```{r}
processed_data_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2025%20Cohort/1.%20Data%20&%20Analysis%20-%20Secure/1.%20Data%20Processed"

# Schools File
ers_write_sharepoint(data = school_raw,
                    folder_path = processed_data_folder_path,
                    file_name_with_extension = paste0(district_folder,"_schools_processed.csv"))

#location crosswalk
ers_write_sharepoint(data = location_crosswalk,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = paste0(district_folder,"_location_crosswalk.csv"))
```