---
title: "01_bb_validation"
author: "Jenny Katz"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3   # Customize the depth of the table of contents for HTML document
    toc_float:
      collapsed: true    # Start with the table of contents expanded
      smooth_scroll: true  # Enable smooth scrolling
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "knitted files", output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 0: Read Libraries and connect to Sharepoint

```{r}
erstools::ers_load_packages()
options(scipen = 99) #prevent any numbers from showing up as scientific notation
```

```{r}
#set sharepoint site URL for access to our org-wide files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"

#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)

#bring in pub available data
nces_data <- read_excel("NCES Data for BB Districts.xlsx")
```

# Part 1: Update Variables

```{r}

#1. folder path to data
raw_data_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2025%20Cohort/1.%20Data%20&%20Analysis%20-%20Secure/0.%20Data%20Received/"

#2. district folder
district_folder <-"Indianapolis"


#3. file names 
budgeted_positions_file <- "FY26 Budgeted Position File - 09172025.xlsx"
school_file <- "FY26 Enrollment Data - 09162025.xlsx"

#4.District & Project INfo
LEAID = "1804770" #Open up nces_data to find
project_year = "2025"
project = "Budget Benchmarking 2025"

```

# Part 2: Bring in Data

```{r}
#4.Bring in files 
#budgeted positions
cold_spring <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder, "/Innovation Data"),
                   file_name_with_extension = "Cold Spring School #315.xlsx", skip_rows = 1)

edison <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder, "/Innovation Data"),
                   file_name_with_extension = "Comparability Report for Innovation Schools - Edison School of the Arts (4).xlsx", sheet_name = "Sheet1", skip_rows = 1)

sankofa <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder, "/Innovation Data"),
                   file_name_with_extension = "Comparability Report for Innovation Schools (SANKOFA) (1) (1).xlsx", skip_rows = 1)

tgns <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder, "/Innovation Data"),
                   file_name_with_extension = "TGNS 2025 - Comparability Report for Innovation Schools.xlsx", skip_rows = 1)

wins <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder, "/Innovation Data"),
                   file_name_with_extension = "WINS 2025 - Comparability Report for Innovation Schools.xlsx", skip_rows = 1)

#rules 

```

## Cleaning up the files

```{r}

cold_spring <- cold_spring %>% 
  filter(!is.na(`Last Name`)) %>% 
  select(-`...9`) %>% 
  mutate(`School Name` = "Cold Spring School",
         `School Number` = 315) %>% 
  select(-c(`School Name/No.#`)) %>% 
  rename(`Total Salary Only                       (No Hourly Rate)` = `Total Salary Only`)


edison <- edison %>% 
  filter(!is.na(`Last Name`)) %>% 
  filter(`Last Name` != "Federally Funded Positions not included in Count") %>% 
  mutate(`School Name` = "Edison School of the Arts",
         `School Number` = 47) %>% 
  select(-c(`School Name/No.#`))
  

sankofa <- sankofa %>% 
  mutate(`School Name` = "Sankofa at Arlington Woods",
         `School Number` = 99) %>% 
  select(-c(`School Name/No.#`))


tgns <- tgns %>% 
  mutate(`School Name` = "Thomas Gregg Neighborhood School",
         `School Number` = 15) %>% 
  select(-c(`School Name/No.#`))


wins <- wins %>% 
  mutate(`School Name` = "Washington Irving Neighborhood School",
         `School Number` = 14) %>% 
  select(-c(`School Name/No.#`))



```

## 2.2 Stacking Together

```{r}

budgeted_positions_raw <- rbind(cold_spring,
                                edison,
                                sankofa,
                                tgns,
                                wins)




```


# Part 3: Clean Up Budgeted Positions

## 3.0 Column Mapping

```{r}

#    # Keep ERS names exactly as is - fill with NAs if missing
budgeted_positions_raw <- budgeted_positions_raw %>% 
  mutate(

    #Project INfo
    H_record_ID = row_number(),
    H_district = district_folder,
    H_LEAID = LEAID,
    H_project_year = project_year,
    H_project = project,
    
    #FTE
    D_employee_id = NA_character_, #critical
    D_empl_fte = FTE, #critical
   
    
    #account codes
    D_activity_code     = NA_character_,
    D_allocation_code = NA_character_,
    D_class_code   = NA_character_,
    D_cost_center_code  = NA_character_,
    D_fund_code    = NA_character_, #critical
    D_funct_code   = NA_character_,
    D_dept_code         = NA_character_,
    D_location_code      = `School Number` + 5500, #critical
    D_position_type_code    = NA_character_,
    D_position_title_code    = NA_character_,
    D_object_code  = NA_character_,
    D_program_code = NA_character_,
    D_project_code = NA_character_,
    
    #account descriptions
    D_activity_desc     = NA_character_,
    D_allocation_desc  = NA_character_,
    D_class_desc   = NA_character_,
    D_cost_center_desc  = "Innovation Data",
    D_fund_desc    = `Funding Source`, #critical
    D_funct_desc   = NA_character_,
    D_dept_desc    = NA_character_, 
    D_location_desc = `School Name`, #critical
    D_position_type_desc     = NA_character_, 
    D_position_title_desc    = `Job Title`, #critical
    D_object_desc  = NA_character_,
    D_program_desc = NA_character_,
    D_project_desc = NA_character_,
  
    
   #other disrict specific fields to add... update as needed
    D_other_1 =  `Total Salary Only                       (No Hourly Rate)`,
    D_other_2 =  NA_character_,
    D_other_3 =  NA_character_,
    D_other_4 =  NA_character_,
    
     .keep = "none"
   
  ) %>%
 select(where(~ !all(is.na(.x))))

```

## 3.1 Manual Checks

-   Look at Data Validation table in OneNote and complete (right year, detailed enough, etc.)

## 3.2 Review Table Structure

```{r}
#check number of columns/rows. 
#does this seem right? 
#ALso check for missing data and data types within the file
ers_column_info(budgeted_positions_raw)
```

## 3.3. Remove Blank Rows/Columns

This code block deletes any rows or columns made up fully of blank values ("", NA," ", etc.). These rows and columns wouldn't give you any additional info so it's best practice to take them out early on.

```{r}

#pre-check
ncol(budgeted_positions_raw)
nrow(budgeted_positions_raw)

#remove empty rows and columns
budgeted_positions_raw <- remove_empty(budgeted_positions_raw,
                                 which = c("rows", "cols"))

#post-check
nrow(budgeted_positions_raw)
ncol(budgeted_positions_raw)

```

## 3.4 Check for Duplicates/Remove

If any rows are repetitive or not required for your analysis you can remove them. For example - you have multiple years of data but only need one.

```{r}

#Check for duplicates
budgeted_positions_raw %>%
 get_dupes(-H_record_ID)

# Duplicates are caused by the lack of employee id (no action needed)

```

## 3.5 Review and modify column data types

Review the data type of each column and modify if needed. Refer to the Data Dictionary for the correct data type for each column. For more information on different data types in R and why it matters, please refer to this: <https://app.tettra.co/teams/ersknowledge/pages/32-r-data-types>

```{r}
#review data type of each column
#ALL Columns except FTE and record ID, should be charachter (this will make joining across tables easier)
col_info <- data.frame(data_type = sapply(budgeted_positions_raw, class))
print(col_info)



#change data type of columns if needed (below is an example to convert to numeric values) - common functions are as.character, as.numeric, as.integer, as.Date

#budgeted_positions_raw$D_empl_FTE <- as.numeric(budgeted_positions_raw$D_empl_FTE)

```

# NO need to adjust fte or apply rules. Will merge with school data once the full data sources are appended together.

```{r}
budgeted_positions_raw <- budgeted_positions_raw %>% 
  mutate(M_fte_adjusted = D_empl_fte)

```

# Part 6. Clean up Schools file

## 6.1 Bring in School Data

```{r}

#schools files
school_raw <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = school_file,
                   skip_rows = 1)


# Removing rows that are used for totals and section headers
school_raw <- school_raw %>%
  dplyr::slice(-c(45:47))


```

## 6.2 Map District Col to ERS Columns

```{r}
#changing column names and removing unneeded columns
#replace the column names after the "=" with column names from your data.You can press tab after the = to select column names available in the data

school_raw <- school_raw %>% 
  mutate(
    
    #Project INfo
    H_record_ID = row_number(),
    H_district = district_folder,
    H_LEAID = LEAID,
    H_project_year = project_year,
    H_project = project,
    
    #school info
    D_location_code = `School ID`,
    D_location_desc = School,
    D_school_level = `Grade Configuration`,
    D_school_type = Program,
    D_grade_span = NA_character_,
    D_enrollment_total = as.numeric(`K-12 Gen Ed`) + coalesce(`Self Contained`, 0),
    D_preK_enrollment = `Pre-K`,
    D_ell_total = `ELL 1` + `ELL 2` + `ELL 3` + `ELL 4`,
    D_frl_total = `Direct Cert %` * D_enrollment_total,
    D_swd_total = NA_character_,
    D_self_contained = `Self Contained`,
    
    
     .keep = "none"
  )
  

```

6.3 Update Data Type

```{r}
#review data type of each column
#ALL Columns except FTE and record ID, should be charachter (this will make joining across tables easier)
col_info <- data.frame(data_type = sapply(school_raw, class))
print(col_info)



#change data type of columns if needed (below is an example to convert to numeric values) - common functions are as.character, as.numeric, as.integer, as.Date

#budgeted_positions_raw$D_empl_FTE <- as.numeric(budgeted_positions_raw$D_empl_FTE)
```

## 6.4 Review School Structure

```{r}
#check number of columns/rows. 
#does this seem right? 
#ALso check for missing data and data types within the file
ers_column_info(school_raw)
```

## 6.5 Check /Remove Duplicate School

### a. Investigate Dups

```{r}

#Check for duplicates
school_raw %>%
 get_dupes(-H_record_ID)


#deuplicate
school_raw<- school_raw %>%
  distinct()

#find shools with the same school ID
duplicates_by_id <- school_raw  %>%
  get_dupes(D_location_code)

```

### b. Remove Duplicates

```{r}
# No duplicates
```

## 6.6 Add Coding Columns

### a. School Level

```{r}
#view school level
school_raw %>%
  group_by(D_school_level) %>%
  count()

#Update for your district's values 
school_raw <-school_raw %>%
  mutate(C_school_level = case_when(D_school_level == "ES"	~ "ES",
                                    D_school_level == "K-12"	~ "Other",
                                    D_school_level == "K-7"	~ "ES",
                                    D_school_level == "HS" ~ "HS",
                                    D_school_level == "MS" ~ "MS",
                                    D_school_level == "K-8" ~ "ES",
                                    D_school_level == "K-5" ~ "ES",
                                    TRUE ~ "Other"))




#check coding
school_raw %>%
  group_by(D_school_level,C_school_level) %>%
  count()

```

### b. Alternative School Flag

```{r}
#UPDATE BASED on DISTRICT'S ALT SCHOOLS

# KF to Update
school_raw <- school_raw %>%
  mutate(
    C_school_flag = case_when(
      D_location_desc %in% c(
    "Liberty Grove at Elder Diggs",
    "Phalen Leadership Academy at Louis B. Russell Jr. School 48",
    "Matchbook Learning @ Wendell Phillips",
    "Emma Donnan Elementary and Middle School 72",
    "Phalen Leadership Academy@George H Fisher",
    "PLA @ 103",
    "KIPP Unite",
    "Impact Academy 1",
    "Impact Academy 2",
    "Impact Academy 3",
    "Roots Off Campus Instruction",
    "REACH Program",
    "Project Search/Site",
    "Global Preparatory Academy @ Riverside (K-6)",
    "The PATH School at Stephen Foster School 67",
    "Positive Support Academy",
    "Step Ahead Academy",
    "Monarca",
    "Enlace",
    "KIPP College Prep",
    "Avondale Meadows Middle School",
    "Phalen Virtual",
    "Newcomer",
    "Herron High School",
    "Herron Riverside High School",
    "Purdue Polytech",
    "Purdue Polytech North",
    "KIPP Legacy High School",
    "Hope Academy",
    "PSOE Virtual",
    "Christel House Academy South",
    "Simon Youth Academy"
    ) ~ "Include: District-Wide Only",  
      TRUE ~ "Include: All"
    )
  )


school_raw <- school_raw %>%
  mutate(
    C_school_flag_reason = case_when(
      D_location_desc %in% c(
    "Liberty Grove at Elder Diggs",
    "Phalen Leadership Academy at Louis B. Russell Jr. School 48",
    "Matchbook Learning @ Wendell Phillips",
    "Emma Donnan Elementary and Middle School 72",
    "Phalen Leadership Academy@George H Fisher",
    "PLA @ 103",
    "KIPP Unite",
    "Impact Academy 1",
    "Impact Academy 2",
    "Impact Academy 3",
    "Roots Off Campus Instruction",
    "REACH Program",
    "Project Search/Site",
    "Global Preparatory Academy @ Riverside (K-6)",
    "The PATH School at Stephen Foster School 67",
    "Positive Support Academy",
    "Step Ahead Academy",
    "Monarca",
    "Enlace",
    "KIPP College Prep",
    "Avondale Meadows Middle School",
    "Phalen Virtual",
    "Newcomer",
    "Herron High School",
    "Herron Riverside High School",
    "Purdue Polytech",
    "Purdue Polytech North",
    "KIPP Legacy High School",
    "Hope Academy",
    "PSOE Virtual",
    "Christel House Academy South",
    "Simon Youth Academy"
    ) ~ "2024 Analysis",  
      TRUE ~ "Neighborhood"
    )
  )

#check coding
school_raw %>%
  group_by(C_school_flag) %>%
  count()

```

### c. Enrollment Characteristics

```{r}

#Update for your district's values 
school_raw <-school_raw %>%
  mutate(C_school_level_mapped = case_when(C_school_level == "ES"	~ "ES",
                                    C_school_level == "MS"	~ "SS",
                                    C_school_level == "HS"	~ "SS",
                                    TRUE ~ "Other"))

school_raw %>%
  group_by(C_school_level,C_school_level_mapped) %>%
  count()

school_raw <- school_raw %>% 
  mutate(C_school_size = ifelse(C_school_level == "ES", ifelse(D_enrollment_total < 350, "Small", "Not Small"), ifelse(D_enrollment_total < 500, "Small", "Not Small")))


school_raw <- school_raw %>%
  mutate(C_gen_ed = D_enrollment_total - D_self_contained,
         M_ell_pct = D_ell_total / D_enrollment_total,
         M_swd_pct = as.numeric(D_swd_total) / D_enrollment_total,
         M_frl_pct = D_frl_total / D_enrollment_total)


# Adding in sanity cleanups - you should manually check any huge variations
school_raw <- school_raw %>%
  mutate(M_ell_pct = ifelse(is.na(M_ell_pct), 0, ifelse(M_ell_pct > 1, 1, M_ell_pct)),
         M_swd_pct = ifelse(is.na(M_swd_pct), 0, ifelse(M_swd_pct > 1, 1, M_swd_pct)),
         M_frl_pct = ifelse(is.na(M_frl_pct), 0, ifelse(M_frl_pct > 1, 1, M_frl_pct)))


```

```{r}

# Create all levels quartiles flags
school_raw <- school_raw %>%
  group_by(C_school_flag) %>%
  mutate(
    M_swd_quartile_flag = case_when(
      M_swd_pct <= quantile(M_swd_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_swd_pct > quantile(M_swd_pct, 0.25, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.5, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag) %>%
  mutate(
    M_ell_quartile_flag = case_when(
      M_ell_pct <= quantile(M_ell_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_ell_pct > quantile(M_ell_pct, 0.25, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.5, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag) %>%
  mutate(
    M_frl_quartile_flag = case_when(
      M_frl_pct <= quantile(M_frl_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_frl_pct > quantile(M_frl_pct, 0.25, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.5, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


```

```{r}

# Create all levels quartiles flags
school_raw <- school_raw %>%
  group_by(C_school_flag, C_school_level_mapped) %>%
  mutate(
    M_swd_quartile_flag_sized = case_when(
      M_swd_pct <= quantile(M_swd_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_swd_pct > quantile(M_swd_pct, 0.25, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.5, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag, C_school_level_mapped) %>%
  mutate(
    M_ell_quartile_flag_sized = case_when(
      M_ell_pct <= quantile(M_ell_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_ell_pct > quantile(M_ell_pct, 0.25, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.5, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag, C_school_level_mapped) %>%
  mutate(
    M_frl_quartile_flag_sized = case_when(
      M_frl_pct <= quantile(M_frl_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_frl_pct > quantile(M_frl_pct, 0.25, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.5, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


```

```{r}

# Validating the quartiles
make_quartile_summary <- function(data, var, flag_var) {
  cutoffs <- quantile(data[[var]], probs = c(0.25, 0.5, 0.75), na.rm = TRUE)
  
  counts <- data %>%
    count(!!sym(flag_var)) %>%
    rename(quartile = !!sym(flag_var), count = n)
  
  tibble(
    metric = var,
    q25 = cutoffs[1],
    q50 = cutoffs[2],
    q75 = cutoffs[3]
  ) %>%
    crossing(counts)
}

# build tables for each metric (only looking at traditional campuses)

## Switch filters to small, not-small as needed
data_testing <- school_raw %>% 
  filter(C_school_flag == "Include: All") %>% 
  filter(C_school_size == "Not Small")

frl_summary <- make_quartile_summary(data_testing, "M_frl_pct", "M_frl_quartile_flag")
swd_summary <- make_quartile_summary(data_testing, "M_swd_pct", "M_swd_quartile_flag")
ell_summary <- make_quartile_summary(data_testing, "M_ell_pct", "M_ell_quartile_flag")

# combine into one concise table
quartile_summary <- bind_rows(frl_summary, swd_summary, ell_summary)

print(quartile_summary)

```

## 6.7 Compare Enrollment to NCES

```{r}

public_data_comp_check <- public_data_comp_check %>%
  left_join(
    (school_raw %>%
       group_by(H_LEAID) %>%
       summarise(M_total_district_enrollment = sum(D_enrollment_total, na.rm = TRUE))),
    by = c("LEAID" = "H_LEAID")
  ) %>%
  mutate(
    M_enrollment_difference = M_total_district_enrollment - `Total Enrollment`
  )
```

# Part 7: Position + Schools Crosswalk

### b. Join Position & School Tables

```{r}

#Create Crosswalk 
location_crosswalk <- bind_rows(
   # --- Part 1: Prepare the Budget Positions Table ---
 budgeted_positions_raw %>%
  filter(!is.na(D_location_code)) %>%
  count(D_location_code, D_location_desc, name = "row_count") %>% 
  mutate( table = "Position") %>% 
  rename("School ID" = D_location_code, 
         "School Name" = D_location_desc),

  # --- Part 2: Prepare the Schools Table ---
  school_raw %>%
    count(D_location_code, D_location_desc, name = "row_count") %>% 
    mutate(table = "School") %>%                           
    rename("School ID" = D_location_code, 
           "School Name" = D_location_desc) 

) %>% 
  
   #Keep only the first school name for each ID ---
  group_by(`School ID`) %>%
  mutate(`School Name` = first(`School Name`)) %>%
  ungroup() %>%

  # --- Part 3: Pivot the Combined Table ---
  pivot_wider(
    names_from = table,
    values_from = row_count,
    values_fill = 0 
  ) %>%
  arrange(`School Name`)



#bring in Total FTE
location_crosswalk <- location_crosswalk %>%
  left_join(
    budgeted_positions_raw %>%
      group_by(D_location_code, D_location_desc) %>%
      summarise(M_total_FTE = sum(M_fte_adjusted)) %>%
      filter(M_total_FTE > 0) %>%
      filter(!is.na(D_location_code)) %>%
      select(D_location_code, M_total_FTE),
    by = c("School ID" = "D_location_code")
  )

#Bring in Total Enrollment 
location_crosswalk <- location_crosswalk %>%
  left_join(
    school_raw %>%
      select(D_location_code, D_enrollment_total,C_school_flag),
    by = c("School ID" = "D_location_code")
  ) %>%
  mutate(M_fte_per_500_students = M_total_FTE / D_enrollment_total * 500) %>%
  arrange(desc(School),C_school_flag)
  

print(location_crosswalk)
```

### c. Fix Location Codes

Note this will have to be updated for each individual districts

```{r}


## Will return to, but Indy location codes in the schools file should add 5500 to the codes that are less than 300

```


# Part 9: Upload File to Sharepoint

## Part 9.1 Any final clean up

Add any final clean up to your file below (remove unncessary columns, update column names to be standard etc.)

```{r}
processed_data_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2025%20Cohort/1.%20Data%20&%20Analysis%20-%20Secure/1.%20Data%20Processed"

#budgeted positions
ers_write_sharepoint(data = budgeted_positions_raw,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = paste0(district_folder,"_innovation_budgeted_positions_processed.csv"))


# #Schools File
# ers_write_sharepoint(data = school_raw,
#                      folder_path = processed_data_folder_path,
#                      file_name_with_extension = paste0(district_folder,"_schools_processed.csv"))
# 
# #location crosswalk
# ers_write_sharepoint(data = location_crosswalk,
#                      folder_path = processed_data_folder_path,
#                      file_name_with_extension = paste0(district_folder,"_location_crosswalk.csv"))
```
