---
title: "01_bb_validation"
author: "Jenny Katz"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3   # Customize the depth of the table of contents for HTML document
    toc_float:
      collapsed: true    # Start with the table of contents expanded
      smooth_scroll: true  # Enable smooth scrolling
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "knitted files", output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 0: Read Libraries and connect to Sharepoint

```{r}
erstools::ers_load_packages()
options(scipen = 99) #prevent any numbers from showing up as scientific notation
```

```{r}
#set sharepoint site URL for access to our org-wide files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"

#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)

#bring in pub available data
nces_data <- read_excel("NCES Data for BB Districts.xlsx")
```

# Part 1: Update Variables

```{r}

#1. folder path to data
raw_data_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2025%20Cohort/1.%20Data%20&%20Analysis%20-%20Secure/0.%20Data%20Received/"

#2. district folder
district_folder <-"Prince William"


#3. file names 
budgeted_positions_file <- "Positions Budget vs Filled 093025.xlsx"
school_file <- "Enrollment September 30 2024.xlsx"

#4.District & Project INfo
LEAID = "5103130" #Open up nces_data to find
project_year = "2025"
project = "Budget Benchmarking 2025"

```

# Part 2: Bring in Data

```{r}
#4.Bring in files 
#budgeted positions
budgeted_positions_raw <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = budgeted_positions_file,
                   skip_rows=0)

## Bring in additional PWCS Specific Files
crosswalks <- "PWCS Chart of Accounts.xlsx"

#budgeted positions
crosswalk_schools <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = crosswalks,
                   sheet_name = "Departments",
                   skip_rows=0)

#budgeted positions
crosswalk_object_codes <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = crosswalks,
                   sheet_name = "Object",
                   skip_rows=0)

#budgeted positions
crosswalk_fund_codes <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = crosswalks,
                   sheet_name = "Fund Codes",
                   skip_rows=0)

#budgeted positions
crosswalk_unit_codes <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = crosswalks,
                   sheet_name = "Unit",
                   skip_rows=0)

#budgeted positions
crosswalk_grant_codes <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = crosswalks,
                   sheet_name = "Grant",
                   skip_rows=0)

#budgeted positions
crosswalk_activity_codes <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = crosswalks,
                   sheet_name = "Activity",
                   skip_rows=0)
```

# Part 3: Clean Up Budgeted Positions

# PWCS-specific merges

### a. schools crosswalk

```{r}
#crosswalk schools
crosswalk_schools <- crosswalk_schools %>%
  mutate(DEPT = str_pad(`Department Codes`, width = 3, pad = "0"))

# remove NAs
crosswalk_schools <- crosswalk_schools %>%
  filter(!if_all(everything(), is.na))

# remove Innovation ES
crosswalk_schools <- crosswalk_schools %>%
  filter(!`Department Description`=="Rosemount Lewis ES") # Removing because there were two schools with the same ID

# Merge crosswalk_schools by department codes
budgeted_positions_raw <- left_join(budgeted_positions_raw, crosswalk_schools, by = "DEPT")

```

### b. object code descriptions

```{r}
# Rename in order to match
crosswalk_object_codes$OBJCD <- as.character(crosswalk_object_codes$`Object Code`)

# Merge in object codes crosswalk
budgeted_positions_raw <- left_join(budgeted_positions_raw, crosswalk_object_codes, by = "OBJCD")
```

### c. Bring in other descriptions

```{r}

fund_xw <- crosswalk_fund_codes %>%
  transmute(
    `Fund Code` = str_pad(as.character(`Fund Code`), 
                          width = 3, side = "left", pad = "0"),
    `Fund Code Description`
  ) %>%
  distinct(`Fund Code`, .keep_all = TRUE)

activity_xw <- crosswalk_activity_codes %>%   # (fix from "acticity" if needed)
  transmute(
    `Activity Code` = str_pad(as.character(`Activity Code`),
                              width = 4, side = "left",pad = "0"),
    `Activity Definition`,
    `Activity Category`
  ) %>%
  distinct(`Activity Code`, .keep_all = TRUE)

unit_xw <- crosswalk_unit_codes %>%
  transmute(
    `Unit Code` = str_pad(as.character(`Unit Code`),
                          width =3,side = "left",pad = "0"),
    `Unit Definition`,
    `Unit Category`
  ) %>%
  distinct(`Unit Code`, .keep_all = TRUE)

# --- 3) Bring in only the fields you need -----------------------------------

budgeted_positions_raw <- budgeted_positions_raw %>%
  # Fund Code Description
  left_join(fund_xw,     by = c("FUND CD" = "Fund Code")) %>%
  # Activity Definition (you already have ACTV CD as the activity code)
  left_join(activity_xw, by = c("ACTV CD" = "Activity Code")) %>%
  # Unit Definition & Unit Category
  left_join(unit_xw,     by = c("UNIT"    = "Unit Code"))

```

## 3.0 Column Mapping

```{r}

#    # Keep ERS names exactly as is - fill with NAs if missing
budgeted_positions_raw <- budgeted_positions_raw %>% 
  mutate(

    #Project INfo
    H_record_ID = row_number(),
    H_district = district_folder,
    H_LEAID = LEAID,
    H_project_year = project_year,
    H_project = project,
    
    #FTE
    D_employee_id = NA_character_,
    #NO EMPLOYEE IDS PROVIDED
    D_empl_fte = `Auth FTE`, #critical
   
    
    #account codes
    D_activity_code     = `ACTV CD`,
    D_allocation_code = NA_character_,
    D_class_code   = UNIT,
    D_cost_center_code  = NA_character_,
    D_fund_code    = `FUND CD`, #critical
    D_funct_code   = NA_character_,
    D_dept_code         = NA_character_,
    D_location_code      = `DEPT`, #critical
    D_position_type_code    = `POSITION ID`,
    D_position_title_code    = NA_character_,
    D_object_code  = `OBJCD`,
    D_program_code = NA_character_,
    D_project_code = NA_character_,
    
    #account descriptions <- NOTE THAT THERE ARE MORE, BUT I'M NOT SURE WHAT ACTUALLY MAPS WHERE
    D_activity_desc = `Activity Definition`,
    D_allocation_desc = `Activity Category`,
    D_class_desc   = `Unit Definition`,
    D_cost_center_desc  = NA_character_,
    D_fund_desc    = `Fund Code Description`, #critical
    D_funct_desc   = NA_character_,
    D_dept_desc    = NA_character_,
    D_location_desc = `Department Description`, #critical
    D_position_type_desc = NA_character_, 
    D_position_title_desc = `TITLE DESCRIPTION`, #critical
    D_object_desc  = `Object Definition`,
    D_program_desc = `Object Category`,
    D_project_desc = NA_character_,
  
    
   #other disrict specific fields to add... update as needed
    D_other_1 =  NA_character_,
    D_other_2 =  `Filled FTE`,
    D_other_3 =  `Vacant FTE`,
    D_other_4 =  NA_character_,
    
     .keep = "none"
   
  ) %>%
 select(where(~ !all(is.na(.x))))

```

## 3.1 Manual Checks

-   Look at Data Validation table in OneNote and complete (right year, detailed enough, etc.)

## 3.2 Review Table Structure

```{r}
#check number of columns/rows. 
#does this seem right? 
#ALso check for missing data and data types within the file
ers_column_info(budgeted_positions_raw)
```

## 3.3. Remove Blank Rows/Columns

This code block deletes any rows or columns made up fully of blank values ("", NA," ", etc.). These rows and columns wouldn't give you any additional info so it's best practice to take them out early on.

```{r}

#pre-check
ncol(budgeted_positions_raw)

#remove empty rows and columns
budgeted_positions_raw <- remove_empty(budgeted_positions_raw,
                                 which = c("rows", "cols"))

#post-check
nrow(budgeted_positions_raw)
ncol(budgeted_positions_raw)

# Remove row with total
# Remove rows where D_location_code is blank
budgeted_positions_raw <- budgeted_positions_raw %>%
  filter(!is.na(D_location_code) & D_location_code != "")
```

## 3.4 Check for Duplicates/Remove

If any rows are repetitive or not required for your analysis you can remove them. For example - you have multiple years of data but only need one.

```{r}

#Check for duplicates
budgeted_positions_raw %>%
 get_dupes(-H_record_ID)

```

## 3.5 Review and modify column data types

Review the data type of each column and modify if needed. Refer to the Data Dictionary for the correct data type for each column. For more information on different data types in R and why it matters, please refer to this: <https://app.tettra.co/teams/ersknowledge/pages/32-r-data-types>

```{r}
#review data type of each column
#ALL Columns except FTE and record ID, should be charachter (this will make joining across tables easier)
col_info <- data.frame(data_type = sapply(budgeted_positions_raw, class))
print(col_info)

#change data type of columns if needed (below is an example to convert to numeric values) - common functions are as.character, as.numeric, as.integer, as.Date

budgeted_positions_raw <- budgeted_positions_raw %>%
  mutate(across(.cols = !c(H_record_ID, D_empl_fte), 
                .fns = as.character))

# check after update
col_info <- data.frame(data_type = sapply(budgeted_positions_raw, class))
print(col_info)
```

# Part 4: FTE Adjustments

Employees may have multiple positions or may work in multiple locations and have a FTE of 1 for each position. To get accurate FTE for each employee, project teams should adjust for this and split the FTE across positions/locations.

## 4.0 Compare FTEs to NCES

```{r}
#compare total FTE in budgeted positions with NCES data
public_data_comp_check <- nces_data %>%
  left_join((budgeted_positions_raw %>%
                group_by(H_LEAID) %>%
              summarise(M_total_district_fte = sum(D_empl_fte, na.rm = TRUE))),
              by = c("LEAID" = "H_LEAID")) %>%
  mutate(M_fte_difference = M_total_district_fte  - `Total Staff FTE`)

```

## 4.1 Investigate FTES

First investigate employees whose FTE sums over 1, so we can investigate why and make sure it's not a larger issue with data that we need to go back to district for

PWCS DID NOT PROVIDE FTE BY INDIVIDUAL EMPLOYEE, SO NOT ADJUSTING

```{r}

#count of employees who have an sum of FTE over 1
# 1) compute per-employee adjusted FTE 
#budgeted_positions_raw <- budgeted_positions_raw %>%
#  group_by(D_employee_id) %>%
#  mutate(H_fte_sum = sum(D_empl_fte, na.rm = TRUE)) %>%
#  ungroup() 


#2) Review employees whose fte sums > 1
#budgeted_positions_raw %>%
#  select(D_employee_id,D_position_title_desc,D_location_desc,H_fte_sum) %>%
 # filter(H_fte_sum >1)

#
```

## 4.2 Adjust FTEs

If issue is not that big, that we can adjust the employee FTE Down

```{r}
#NOTE:  DO NOT ADJUST VACANT FTE DOWN IF THEY HAVE THE SAME EMPLOYEE ID (NA,0, 1, blanks)

#1 Adjust the FTE Down
#  budgeted_positions_raw <- budgeted_positions_raw %>%
#  mutate(M_fte_adjusted = D_empl_fte / H_fte_sum) %>%
#  select(-H_fte_sum)

#1.PWCS
budgeted_positions_raw <- budgeted_positions_raw %>%
  mutate(M_fte_adjusted = D_empl_fte) 

#2) Double Check that no adjusted FTES are greater than 1
#budgeted_positions_raw %>%
 # select(D_employee_id,D_position_title_desc,D_location_desc,M_fte_adjusted) %>%
  #filter(M_fte_adjusted >1)
```

# Part 5: Apply Rules (If applicable)

## 5.1 DELETED BC NOT REPEAT DISTRICT

```         
```

## 5.2 Run Apply Rules Function

```         
```

## 5.3 Test Rules

```         
```

# Part 6. Clean up Schools file

## 6.1 Bring in School Data

```{r}
#schools files
school_raw <- ers_read_sharepoint(folder_path = paste0(raw_data_folder_path,district_folder),
                   file_name_with_extension = school_file,
                   skip_rows = 0)
```

## 6.2 Map District Col to ERS Columns

### a. PWCS ONLY: Create a grade span flag

```{r}
# ---- 1. Define grade columns and friendly names ----
GRADE_COLS <- c(
   "KG (C09)", "Grade 1 (C10)", "Grade 2 (C11)",
  "Grade 3 (C12)", "Grade 4 (C13)", "Grade 5 (C14)",
  "Grade 6 (C15)", "Grade 7 (C16)", "Grade 8 (C17)",
  "Grade 9 (C18)", "Grade 10 (C19)", "Grade 11 (C20)", "Grade 12 (C21)"
)

GRADE_NAMES <- c("K","1","2","3","4","5","6","7","8","9","10","11","12")

# only keep columns that exist in your file
present_cols  <- intersect(GRADE_COLS, names(school_raw))
present_names <- GRADE_NAMES[match(present_cols, GRADE_COLS)]

# ---- 2. Build grades_offered for each school ----
school_raw <- school_raw %>%
  rowwise() %>%
  mutate(
    grades_offered = list({
      vals <- c_across(all_of(present_cols))
      nums <- suppressWarnings(as.numeric(vals))
      nums[is.na(nums)] <- 0
      present_names[which(nums > 0)]
    })
  ) %>%
  ungroup()

# ---- 3. Create readable string and grade span ----
school_raw <- school_raw %>%
  mutate(
    grades_offered_str = map_chr(grades_offered, ~ if (length(.x)) paste(.x, collapse = ", ") else NA_character_),
    D_grade_span = map_chr(grades_offered, ~ 
      if (length(.x)) paste0(first(.x), "–", last(.x)) else NA_character_)
  )

# ---- 4. Label school level from span ----
school_raw <- school_raw %>%
  mutate(
    D_grade_span = if_else(`School Number` == 244, "PK", D_grade_span),
  )

# ---- 5. Sanity check ----
school_raw %>%
  select(grades_offered_str, D_grade_span) %>%
  head(102)

```

### b. PWCS ONLY: Create a school level flag

```{r}

school_raw <-school_raw %>%
  mutate(D_school_level= case_when(
    D_grade_span %in% c("1–8","K–8") ~ "K8",
    D_grade_span %in% c("K–5") ~ "ES",
    D_grade_span %in% c("6–8") ~ "MS",
    D_grade_span %in% c("9–12") ~ "HS",
    D_grade_span %in% c("7–12") ~ "SS",
    TRUE ~ "Other")
    
  )


school_raw %>%
  group_by(D_grade_span,D_school_level) %>%
  arrange(D_grade_span) %>%
  count()
```

```{r}
# Print names in school data frame
cat("\nData Frame Column Names:\n")
print(names(school_raw))

#changing column names and removing unneeded columns
#replace the column names after the "=" with column names from your data.You can press tab after the = to select column names available in the data

school_raw <- school_raw %>% 
  mutate(
    
    #Project INfo
    H_record_ID = row_number(),
    H_district = district_folder,
    H_LEAID = LEAID,
    H_project_year = project_year,
    H_project = project,
    
    #school info
    D_location_code = as.character(`School Number`),
    D_location_desc = School,
    D_school_level = D_school_level,
    D_school_type = NA_character_,
    D_grade_span = D_grade_span,
    D_enrollment_total = `Total (w/all PK)` - replace_na(`Pre-School (Head Start)`, 0) - replace_na(`Pre-School (VPI)`, 0),
    D_ell_total = `ESOL Total (Lvls. 1 - 6 [yrs.1-4] and Opt Outs)`,
    D_frl_total = `Economically Disadvantaged`,
    D_swd_total = `Special Education`,
    D_self_contained = `Total Self-Contained`,
    
     .keep = "none"
  )




```

```{r}
#PWCS ONLY
school_raw <- school_raw %>%
  select(
    H_record_ID, H_district, H_LEAID, H_project_year, H_project,
    D_location_code, D_location_desc, D_school_level, D_school_type,
    D_grade_span, D_enrollment_total, D_ell_total, D_frl_total,
    D_swd_total, D_self_contained
  )
```

## 6.3 Update Data Type

```{r}
#review data type of each column
col_info <- data.frame(data_type = sapply(school_raw, class))
print(col_info)

#change data type of columns if needed (below is an example to convert to numeric values) - common functions are as.character, as.numeric, as.integer, as.Date

school_raw$D_ell_total <- as.numeric(school_raw$D_ell_total)
school_raw$D_frl_total <- as.numeric(school_raw$D_frl_total)
school_raw$D_swd_total <- as.numeric(school_raw$D_swd_total)


#budgeted_positions_raw$D_empl_FTE <- as.numeric(budgeted_positions_raw$D_empl_FTE)

# Remove rows where D_location_code is blank
school_raw <- school_raw %>%
  filter(!is.na(D_location_code) & D_location_code != "")

# Remove PK
school_raw <- school_raw %>%
  filter(D_location_code != "244")
```

## 6.4 Review School Structure

```{r}
#check number of columns/rows. 
#does this seem right? 
#ALso check for missing data and data types within the file
ers_column_info(school_raw)
```

## 6.5 Check /Remove Duplicate School

### a. Investigate Dups

```{r}

#Check for duplicates
school_raw %>%
 get_dupes(-H_record_ID)


#deuplicate
school_raw<- school_raw %>%
  distinct()

#find shools with the same school ID
duplicates_by_id <- school_raw  %>%
  get_dupes(D_location_code)

```

### b. Remove Duplicates

There are no duplicates, so we don't need to remove any.

```{r}
# #NOTE THIS WAS A DETROIT SPECIFIC DUP REMOVAL. MAY BE DIFF FOR YOUR DISTRICT
# 
# # Add a count column 'n' based on school_id
# school_raw <- school_raw %>%
#   add_count(D_location_code)
# 
# # Separate the duplicates (groups where n > 1)
# duplicates_to_summarize <- school_raw %>%
#   filter(n > 1)
# 
# # Separate the unique rows (where n == 1)
# school_raw <- school_raw %>%
#   filter(n == 1)
# 
# #summarize duplicates
# #ThiS Take the two different grade spans/names and concats to one cell
# duplicates_to_summarize <- duplicates_to_summarize %>%
#   group_by(D_location_code) %>%
#   mutate(
#     # These calculations are applied to the whole group...
#     D_location_desc = paste(unique(D_location_desc), collapse = " / "),
#     D_enrollment_total = sum(D_enrollment_total, na.rm = TRUE),
#     D_grade_span = paste(unique(D_grade_span), collapse = " / ")
#   ) %>%
#   # ...then we just keep the first row of the group.
#   slice_head(n = 1) %>%
#   ungroup() # Always ungroup after you're done.
# 
# #stack it back onto school data
# school_raw <-school_raw %>%
#   bind_rows(duplicates_to_summarize) %>% 
#               select(-n)

```

## 6.6 Add Coding Columns

### a. School Level

```{r}
#view school level
school_raw %>%
  group_by(D_school_level) %>%
  count()

#Update for your district's values 
school_raw <-school_raw %>%
  mutate(C_school_level = case_when(D_school_level %in% c("ES","K8")	~ "ES",
                                    D_school_level == "HS" ~ "HS",
                                    D_school_level == "MS" ~ "MS",
                                    TRUE ~ "Other"))

#check coding
school_raw %>%
  group_by(D_school_level,C_school_level) %>%
  count()
```

### b. School Include Flag

```{r}
#UPDATE BASED on DISTRICT'S ALT SCHOOLS
school_raw <- school_raw %>%
  mutate(
    C_school_flag = case_when(
      D_location_desc %in% c(
        "Independence Nontraditional School - ED Center",
        "Independence Nontraditional School - High",
        "Independence Nontraditional School - Middle",
        "PACE West"
        ) ~ "Include: District-Wide Only",  
      TRUE ~ "Include: All"
    )
  )

#UPDATE BASED on DISTRICT'S ALT SCHOOLS
school_raw <- school_raw %>%
  mutate(
    C_school_flag_reason = case_when(
      D_location_desc %in% c(
        "Independence Nontraditional School - ED Center",
        "Independence Nontraditional School - High",
        "Independence Nontraditional School - Middle") ~ "Alternative",  
      D_location_desc == "PACE West" ~ "Special Education",
      TRUE ~ "Traditional/Neighborhood"
    )
  )



#check coding
school_raw %>%
  group_by(C_school_flag) %>%
  count()

```

### c. PWCS ONLY: Fix Location Codes for Independence Schools

```{r}
#Fix location codes for   "Independence Nontraditional Schools" so they all have the same location code and can join onto the position control file.

school_raw <- school_raw %>%
  mutate(
    D_location_code = case_when(
      D_location_desc %in% c(
        "Independence Nontraditional School - ED Center",
        "Independence Nontraditional School - High",
        "Independence Nontraditional School - Middle") ~ "240",
      TRUE ~ D_location_code
    ))


#Now we combine these locations into one field
 # Add a count column 'n' based on school_id
 school_raw <- school_raw %>%
  add_count(D_location_code)
 
 #Separate the duplicates (groups where n > 1)
 duplicates_to_summarize <- school_raw %>%
  filter(n > 1)
 
# # Separate the unique rows (where n == 1)
 school_raw <- school_raw %>%
  filter(n == 1)
# 
# #summarize duplicates
# #ThiS Take the two different grade spans/names and concats to one cell
 duplicates_to_summarize <- duplicates_to_summarize %>%
   group_by(D_location_code) %>%
  mutate(
    # These calculations are applied to the whole group...
    D_location_desc = paste(unique(D_location_desc), collapse = " / "),
    D_enrollment_total = sum(D_enrollment_total, na.rm = TRUE),
    D_ell_total = sum(D_ell_total , na.rm = TRUE),
    D_frl_total = sum(D_frl_total , na.rm = TRUE),
    D_swd_total = sum(D_swd_total , na.rm = TRUE),
    D_self_contained = sum(D_self_contained, na.rm = TRUE),
    D_school_level = paste(unique(D_school_level ), collapse = " / "),
    C_school_level = "Other",
    C_school_flag = "Include: District-Wide Only",
    C_school_flag_reason = "Alternative"
    
   ) %>%
   # ...then we just keep the first row of the group.
   slice_head(n = 1) %>%
  ungroup() # Always ungroup after you're done.
 
# #stack it back onto school data
school_raw <-school_raw %>%
   bind_rows(duplicates_to_summarize) %>% 
              select(-n)

school_raw <-school_raw %>%
  select(-C_school_type)
```

### c. Enrollment Characteristics

```{r}

#Update for your district's values 
school_raw <-school_raw %>%
  mutate(C_school_level_mapped = case_when(C_school_level == "ES"	~ "ES",
                                    C_school_level == "MS"	~ "SS",
                                    C_school_level == "HS"	~ "SS",
                                    TRUE ~ "Other"))

school_raw %>%
  group_by(C_school_level,C_school_level_mapped) %>%
  count()

school_raw <- school_raw %>% 
  mutate(C_school_size = ifelse(C_school_level == "ES", ifelse(D_enrollment_total < 350, "Small", "Not Small"), ifelse(D_enrollment_total < 500, "Small", "Not Small")))


school_raw <- school_raw %>%
  mutate(C_gen_ed = coalesce(D_enrollment_total - D_self_contained,
                             D_enrollment_total),                     
         M_ell_pct = D_ell_total / D_enrollment_total,
         M_swd_pct = as.numeric(D_swd_total) / D_enrollment_total,
         M_frl_pct = D_frl_total / D_enrollment_total)


# Adding in sanity cleanups - you should manually check any huge variations
school_raw <- school_raw %>%
  mutate(M_ell_pct = ifelse(is.na(M_ell_pct), 0, ifelse(M_ell_pct > 1, 1, M_ell_pct)),
         M_swd_pct = ifelse(is.na(M_swd_pct), 0, ifelse(M_swd_pct > 1, 1, M_swd_pct)),
         M_frl_pct = ifelse(is.na(M_frl_pct), 0, ifelse(M_frl_pct > 1, 1, M_frl_pct)))



```

```{r}

# Create all levels quartiles flags
school_raw <- school_raw %>%
  group_by(C_school_flag) %>%
  mutate(
    M_swd_quartile_flag = case_when(
      M_swd_pct <= quantile(M_swd_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_swd_pct > quantile(M_swd_pct, 0.25, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.5, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag) %>%
  mutate(
    M_ell_quartile_flag = case_when(
      M_ell_pct <= quantile(M_ell_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_ell_pct > quantile(M_ell_pct, 0.25, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.5, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag) %>%
  mutate(
    M_frl_quartile_flag = case_when(
      M_frl_pct <= quantile(M_frl_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_frl_pct > quantile(M_frl_pct, 0.25, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.5, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


```

```{r}

# Create all levels quartiles flags
school_raw <- school_raw %>%
  group_by(C_school_flag, C_school_level_mapped) %>%
  mutate(
    M_swd_quartile_flag_sized = case_when(
      M_swd_pct <= quantile(M_swd_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_swd_pct > quantile(M_swd_pct, 0.25, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.5, na.rm = TRUE) & M_swd_pct <= quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_swd_pct > quantile(M_swd_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag, C_school_level_mapped) %>%
  mutate(
    M_ell_quartile_flag_sized = case_when(
      M_ell_pct <= quantile(M_ell_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_ell_pct > quantile(M_ell_pct, 0.25, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.5, na.rm = TRUE) & M_ell_pct <= quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_ell_pct > quantile(M_ell_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


school_raw <- school_raw %>%
  group_by(C_school_flag, C_school_level_mapped) %>%
  mutate(
    M_frl_quartile_flag_sized = case_when(
      M_frl_pct <= quantile(M_frl_pct, 0.25, na.rm = TRUE) ~ "Low",
      M_frl_pct > quantile(M_frl_pct, 0.25, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.5, na.rm = TRUE) ~ "Low-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.5, na.rm = TRUE) & M_frl_pct <= quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High-Middle",
      M_frl_pct > quantile(M_frl_pct, 0.75, na.rm = TRUE) ~ "High",
      TRUE ~ "0" # Handle any missing or unexpected cases
    )
  ) %>%
  ungroup()


```

```{r}

# Validating the quartiles
make_quartile_summary <- function(data, var, flag_var) {
  cutoffs <- quantile(data[[var]], probs = c(0.25, 0.5, 0.75), na.rm = TRUE)
  
  counts <- data %>%
    count(!!sym(flag_var)) %>%
    rename(quartile = !!sym(flag_var), count = n)
  
  tibble(
    metric = var,
    q25 = cutoffs[1],
    q50 = cutoffs[2],
    q75 = cutoffs[3]
  ) %>%
    crossing(counts)
}

# build tables for each metric (only looking at traditional campuses)

## Switch filters to small, not-small as needed
data_testing <- school_raw %>% 
  filter(C_school_flag == "Include: All") %>% 
  filter(C_school_size == "Not Small")

frl_summary <- make_quartile_summary(data_testing, "M_frl_pct", "M_frl_quartile_flag")
swd_summary <- make_quartile_summary(data_testing, "M_swd_pct", "M_swd_quartile_flag")
ell_summary <- make_quartile_summary(data_testing, "M_ell_pct", "M_ell_quartile_flag")

# combine into one concise table
quartile_summary <- bind_rows(frl_summary, swd_summary, ell_summary)

print(quartile_summary)

```

## 6.7 Compare Enrollment to NCES

```{r}

public_data_comp_check <- public_data_comp_check %>%
  left_join(
    (school_raw %>%
       group_by(H_LEAID) %>%
       summarise(M_total_district_enrollment = sum(D_enrollment_total, na.rm = TRUE))),
    by = c("LEAID" = "H_LEAID")
  ) %>%
  mutate(
    M_enrollment_difference = M_total_district_enrollment - `Total Enrollment`
  )
```

# b. Calculate subgroups

```{r}
# Assuming your data frame is named 'schools_raw'

# 1. Calculate the overall sum for each individual subgroup and total enrollment
sum_ell <- sum(school_raw$D_ell_total, na.rm = TRUE)
 sum_frl <- sum(school_raw$D_frl_total, na.rm = TRUE)
sum_swd <- sum(school_raw$D_swd_total, na.rm = TRUE)
sum_enrollment <- sum(school_raw$D_enrollment_total, na.rm = TRUE)

# 2. Calculate the percentage of the sum of D_enrollment_total for each subgroup

# Percent for English Language Learners (ELL)
percent_ell <- (sum_ell / sum_enrollment) * 100

# Percent for Free or Reduced-Price Lunch (FRL)
percent_frl <- (sum_frl / sum_enrollment) * 100

# Percent for Students with Disabilities (SWD)
percent_swd <- (sum_swd / sum_enrollment) * 100

# 3. Display the results
cat("Overall Sums:\n")
cat("Sum of ELL (D_ell_total):", sum_ell, "\n")
 cat("Sum of FRL (D_frl_total):", sum_frl, "\n")
cat("Sum of SWD (D_swd_total):", sum_swd, "\n")
cat("Sum of Total Enrollment (D_enrollment_total):", sum_enrollment, "\n\n")

cat("Individual Subgroup Percentages of Total Enrollment:\n")
cat("Percent of ELL (D_ell_total):", round(percent_ell, 2), "%\n")
cat("Percent of FRL (D_frl_total):", round(percent_frl, 2), "%\n")
cat("Percent of SWD (D_swd_total):", round(percent_swd, 2), "%\n")
```

# Part 7: Position + Schools Crosswalk

### b. Join Position & School Tables

```{r}
#Create Crosswalk 
location_crosswalk <- bind_rows(
   # --- Part 1: Prepare the Budget Positions Table ---
 budgeted_positions_raw %>%
  filter(!is.na(D_location_code)) %>%
  count(D_location_code, D_location_desc, name = "row_count") %>% 
  mutate( table = "Position") ,

  # --- Part 2: Prepare the Schools Table ---
  school_raw %>%
    count(D_location_code, D_location_desc, name = "row_count") %>% 
    mutate(table = "School")

) %>% 
  
   #Keep only the first school name for each ID ---
  group_by(D_location_code) %>%
  arrange(desc(table)) %>%
  mutate(D_location_desc = first(D_location_desc)) %>%
  ungroup() %>%

  # --- Part 3: Pivot the Combined Table ---
  pivot_wider(
    names_from = table,
    values_from = row_count,
    values_fill = 0 
  ) %>%
  arrange(D_location_desc)


#bring in Total FTE
location_crosswalk <- location_crosswalk %>%
  left_join(
    budgeted_positions_raw %>%
      group_by(D_location_code, D_location_desc) %>%
      summarise(M_total_FTE = sum(M_fte_adjusted)) %>%
      filter(M_total_FTE > 0) %>%
      filter(!is.na(D_location_code)) %>%
      select(D_location_code, M_total_FTE),
    by = "D_location_code"
  )

#Bring in Total Enrollment 
location_crosswalk <- location_crosswalk %>%
  left_join(
    school_raw %>%
      select(D_location_code, D_enrollment_total,D_school_type),
    by = "D_location_code"
  ) %>%
  mutate(M_fte_per_500_students = M_total_FTE / D_enrollment_total * 500) %>%
  arrange(desc(School))
  

#flag issues
location_crosswalk <- location_crosswalk %>%
  mutate(
    C_location_issue_flag = case_when(
      Position < 1 & School == 1 ~ "Not In Position File",
      School == 0 &
        Position > 0 ~ "Not In School File",
      M_fte_per_500_students < 40 |
        M_fte_per_500_students > 90 ~ "Low/High FTE per 500",
      D_enrollment_total < 10 &
        D_enrollment_total > 0 ~ "Low Enrollment",
      TRUE ~ "Okay"
    )
  )
print(location_crosswalk)  
```

### c. Fix Location Codes

I Don't think we have any to fix here???

```{r}

location_crosswalk <- location_crosswalk %>%
  left_join(
    school_raw %>%
      select(D_location_code, C_school_flag,C_school_flag_reason),
    by = "D_location_code")

```

# Part 8: Machine Learning

## 8.0 Set-Up

```{r}

library(tm)
library(textstem) #for turning words into their stem form
library(randomForest) #for random forest model
library(ranger) #for random forest model
library(xgboost) #for random forest model

#set sharepoint site URL for access to our org-wide and external files
orgfiles_site_url <- "https://erstrategies1.sharepoint.com/sites/orgfiles"
#access sharepoint site URL you may be prompted to sign in here
orgfiles_site <- get_sharepoint_site(site_url = orgfiles_site_url)

client_work_drive <- orgfiles_site$get_drive("Client Work")

# Modified read_sharepoint logic for RDS files
read_sharepoint_ers_obj <- function(drive_name = client_work_drive, 
                                folder_path, 
                                file_name_with_extension, 
                                sheet_name = NULL, 
                                skip_rows = 0,
                                na_values = c("","N/A","NA","Missing")) {
  # Construct the full data path
  data_path <- file.path(folder_path, file_name_with_extension)
  
  file_extension <- paste0(".", tools::file_ext(file_name_with_extension))
  
  # Create a temporary file with the appropriate file extension
  temp_file <- tempfile(fileext = file_extension)
  
  # Download the file from the specified drive to the temporary file
  drive_name$download_file(data_path, dest = temp_file)
  
  # Read the downloaded file based on its file extension
  data_loaded <- readRDS(temp_file)
  
  # Return the loaded data
  return(data_loaded)
}



district_personnel <- budgeted_positions_raw

training_bag_columns_position <- ers_read_sharepoint(folder_path = "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/Template/BB%20Methodology%20&%20Guidance/Machine Learning Models",
  file_name_with_extension =  "position_type_training_bag_columns_9.24.24.csv")

model_position <- read_sharepoint_ers_obj(folder_path = "Other/Budget Benchmarking/Template/BB Methodology & Guidance/Machine Learning Models", file_name_with_extension =  "position_type_model_9.24.24.rds")


training_bag_columns_function <- ers_read_sharepoint(folder_path = "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/Template/BB%20Methodology%20&%20Guidance/Machine Learning Models",
  file_name_with_extension =  "function_training_bag_columns_9.24.24.csv")

model_function <- read_sharepoint_ers_obj(folder_path = "Other/Budget Benchmarking/Template/BB Methodology & Guidance/Machine Learning Models", file_name_with_extension =  "function_model_9.24.24.rds")

```

## 8.1 Modify Data

```{r}

# Add a reporting proxy. To train the ML model, we need a reporting coding, in leiu of a complete set of coding, we can use any proxy to get the job done. In Indy they had a similar column structure that we could use, otherwise you can use whether or not there is a match in the school's file as a nice proxy

district_personnel <- district_personnel %>%
  rename(`C_reporting` = 1, `D_position_title` = `D_position_title_desc`)
  
```

## 8.2 Load Functions

```{r}

# All functions

#building a prefix function that allows me to add a tag into all words to mark which variable they come from
add_prefix_to_words <- function(string, prefix) {
  # Split the string into words
  words <- unlist(strsplit(string, " "))
  
  # Add prefix to each word
  prefixed_words <- paste0(prefix, words)
  
  # Rejoin the words into a single string
  prefixed_string <- paste(prefixed_words, collapse = " ")
  
  return(prefixed_string)
}

#subset to columns needed (district_data_concat is combination of all fields relevant for prediction and standardized to only regular characters)
prediction_data_prep <- function(data,
                                 predicted_value) {
  data %>%
  mutate(D_position_title = str_replace_all(D_position_title,
                                           "[^[:alnum:]]",
                                           " "),
         D_location_desc = str_replace_all(D_location_desc,
                                           "[^[:alnum:]]",
                                           " "),
         C_reporting = str_replace_all(C_reporting,
                                           "[^[:alnum:]]",
                                           " "),
         D_position_title = sapply(D_position_title,
                                 add_prefix_to_words,
                                 "pt"),
         D_location_desc = sapply(D_location_desc,
                                  add_prefix_to_words,
                                  "ld"),
         C_reporting = sapply(C_reporting,
                              add_prefix_to_words,
                              "rp"),
         district_data_concat = paste(D_position_title,
                                      D_location_desc,
                                      C_reporting,
                                      sep = " ")) %>%
  select(district_data_concat)
}


# Warnings here are fine
bag_prep <- function(data) {
  Corpus(VectorSource(data$district_data_concat)) %>%
  tm_map(content_transformer(tolower)) %>% 
  tm_map(removePunctuation) %>% 
  tm_map(removeNumbers) %>% 
  tm_map(removeWords, stopwords("english")) %>% 
  tm_map(stripWhitespace) %>% 
  DocumentTermMatrix() %>%
  as.matrix() %>% 
  as.data.frame()
}

#adding key terms if they're not already in data (so that select doesn't break)
add_each_col <- function(dataset, col_name) {
  if (!(col_name %in% names(dataset))) {
    dataset <- dataset %>% 
      mutate(!!col_name := 0)
  }
  return(dataset)
}

add_all_cols <- function(dataset) {
  dataset %>% 
  add_each_col("pttchr") %>% 
  add_each_col("pttch") %>% 
  add_each_col("ptasst") %>% 
  add_each_col("ptpara") %>% 
  add_each_col("ptsvcs") %>% 
  add_each_col("ptmgr") %>% 
  add_each_col("ptcoor") %>% 
  add_each_col("ptcoord") %>% 
  add_each_col("ptteacher") %>% 
  add_each_col("ptassistant") %>% 
  add_each_col("ptparaprofessional") %>% 
  add_each_col("ptservices") %>% 
  add_each_col("ptmanager") %>% 
  add_each_col("ptcoordinator")
}

fix_abbrevs <- function(data) {
  data %>% 
  mutate(ptteacher = ifelse(pttchr == 1 | pttch == 1 | ptteacher == 1, 1, 0),
         ptassistant = ifelse(ptasst == 1 | ptassistant == 1, 1, 0),
         ptparaprofessional = ifelse(ptpara ==1 | ptparaprofessional ==1, 1, 0),
         ptservices = ifelse(ptservices == 1 | ptsvcs == 1, 1, 0),
         ptmanager = ifelse(ptmanager == 1 | ptmgr == 1, 1, 0),
         ptcoordinator=ifelse(ptcoord==1|ptcoor==1|ptcoordinator==1, 1, 0)) %>% 
  select(-c(pttchr,
            pttch,
            ptasst,
            ptpara,
            ptsvcs,
            ptmgr,
            ptcoord,
            ptcoor))
}


#removing words that weren't present in training data

align_bags <- function(list_of_cols, test_data) {

  non_training_words <- setdiff(colnames(test_data),
                                list_of_cols$`.`)

  adjusted_test_data <- test_data %>%
    select(-any_of(non_training_words))

  return(adjusted_test_data)
}

add_missing_cols <- function(list_of_cols, test_df) {

  missing_columns <- setdiff(list_of_cols$`.`,
                             names(test_df))

  zero_values_df <- as_tibble(
    matrix(0, ncol = length(missing_columns),
           nrow = nrow(test_df),
           dimnames = list(NULL, missing_columns))
  )

  adjusted_test_df <- bind_cols(test_df,
                                zero_values_df)

  return(adjusted_test_df)
}


prediction_data_prep_exp <- function(data) {
  data %>%
  mutate(D_position_title_h = str_replace_all(D_position_title,
                                           "[^[:alnum:]]",
                                           " "),
         D_location_desc_h = str_replace_all(D_location_desc,
                                           "[^[:alnum:]]",
                                           " "),
         C_reporting_h = str_replace_all(C_reporting,
                                           "[^[:alnum:]]",
                                           " "),
         D_position_title_h = sapply(D_position_title_h,
                                 add_prefix_to_words,
                                 "pt"),
         D_location_desc_h = sapply(D_location_desc_h,
                                  add_prefix_to_words,
                                  "ld"),
         C_reporting_h = sapply(C_reporting_h,
                              add_prefix_to_words,
                              "rp"),
         district_data_concat = paste(D_position_title_h,
                                      D_location_desc_h,
                                      C_reporting_h,
                                      sep = " ")) %>% 
    select(-c(D_location_desc_h, D_position_title_h, C_reporting_h))
}
```

```{r}
# ERS content functions
turn_position_type_into_values <- function(dataset) {
  dataset %>% 
    mutate(C_position_type = case_when(
      C_position_type == 0 ~ "Substitute",
      C_position_type == 1 ~ "Teacher",
      C_position_type == 2 ~ "Other",
      C_position_type == 3 ~ "Nurse",
      C_position_type == 4 ~ "TA",
      C_position_type == 5 ~ "Sec/Clerk/Other Admin",
      C_position_type == 6 ~ "Coordinator/Manager",
      C_position_type == 7 ~ "Principal",
      C_position_type == 8 ~ "Speech Therapist",
      C_position_type == 9 ~ "Guidance Counselor",
      C_position_type == 10 ~ "Custodian",
      C_position_type == 11 ~ "Psychologist",
      C_position_type == 12 ~ "Instructional Coach",
      C_position_type == 13 ~ "Vice Principal",
      C_position_type == 14 ~ "Social Worker",
      C_position_type == 15 ~ "School Monitor/Security",
      C_position_type == 16 ~ "Club Advisor/Coach",
      C_position_type == 17 ~ "Librarian",
      C_position_type == 18 ~ "(Exec) Director",
      C_position_type == 19 ~ "Non-Position",
      C_position_type == 20 ~ "Nurse Aide",
      C_position_type == 21 ~ "Area Officers",
      C_position_type == 22 ~ "Occupational Therapist",
      C_position_type == 23 ~ "Physical Therapist",
      C_position_type == 24 ~ "Instructional Coach Supervisors",
      TRUE ~ NA
    ))
}


turn_function_into_values <- function(dataset) {
  dataset %>% 
    mutate(C_function = case_when(
      C_function == 0 ~ "Other Non-Compensation",
      C_function == 1 ~ "Teacher Compensation",
      C_function == 2 ~ "Aides Compensation",
      C_function == 3 ~ "Substitute Compensation",
      C_function == 4 ~ "Library & Media",
      C_function == 5 ~ "Instructional Materials & Supplies",
      C_function == 6 ~ "Other Compensation",
      C_function == 7 ~ "Extended Time & Tutoring",
      C_function == 8 ~ "Enrichment",
      C_function == 9 ~ "Social & Emotional",
      C_function == 10 ~ "Physical Health & Services",
      C_function == 11 ~ "Career & Academic Counseling",
      C_function == 12 ~ "Parent & Community Relations",
      C_function == 13 ~ "Professional Growth",
      C_function == 14 ~ "Curriculum Development",
      C_function == 15 ~ "Recruitment",
      C_function == 16 ~ "Special Population Program Management & Support",
      C_function == 17 ~ "Governance",
      C_function == 18 ~ "School Supervision",
      C_function == 19 ~ "School Administration",
      C_function == 20 ~ "Research & Accountability",
      C_function == 21 ~ "Communications",
      C_function == 22 ~ "Student Assignment",
      C_function == 23 ~ "Facilities & Maintenance",
      C_function == 24 ~ "Security & Safety",
      C_function == 25 ~ "Food Services",
      C_function == 26 ~ "Student Transportation",
      C_function == 27 ~ "Utilities",
      C_function == 28 ~ "Human Resources",
      C_function == 29 ~ "Finance, Budget, Purchasing & Distribution",
      C_function == 30 ~ "Data Processing & Information Services",
      C_function == 31 ~ "Facilities Planning",
      C_function == 32 ~ "Development & Fundraising",
      C_function == 33 ~ "Legal",
      C_function == 34 ~ "Insurance",
      TRUE ~ "Other Non-Compensation"))
}


```

## 8.3 Train Position Type Model

```{r}


district_personnel_position_type <- prediction_data_prep(district_personnel,
                                                         "C_position_type")

#making all district dataframes a single row per combination of values
district_personnel_position_type <- district_personnel_position_type %>% 
  group_by(district_data_concat) %>% 
  summarize()

# Prep the bags
district_personnel_position_type_test_bag <- bag_prep(district_personnel_position_type)

# Clean abreviations and columns
district_personnel_position_type_test_bag <- add_all_cols(district_personnel_position_type_test_bag)
district_personnel_position_type_test_bag <- fix_abbrevs(district_personnel_position_type_test_bag)


district_personnel_position_type_test_bag <- align_bags(training_bag_columns_position,
                                                        district_personnel_position_type_test_bag)

# Identify and add missing columns in the test data
district_personnel_position_type_test_bag <- add_missing_cols(training_bag_columns_position,                              district_personnel_position_type_test_bag)

set.seed(223)

#ordering columns the exact same order across training and test

  ## Get the column names except "C_position_type"
column_sorting <- sort(setdiff(training_bag_columns_position$'.', "C_position_type"))

new_column_order_test <- c(column_sorting)

position_type_district_test_bag <- district_personnel_position_type_test_bag[, new_column_order_test]



# Make predictions on the test data
rf_position_type_predictions_district <- predict(model_position,
                                                 as.matrix(position_type_district_test_bag))

#turning the predictions into a probability dataframe for each row in original data
probability_matrix_district_position_type <- data.frame(
  row_number = rep(c(1:nrow(district_personnel_position_type)),
                   each = 25),
  C_position_type = rep(c(0:24), nrow(district_personnel_position_type)),
  confidence = rf_position_type_predictions_district)

#shifting back to character values for position type
probability_matrix_district_position_type <-
  turn_position_type_into_values(probability_matrix_district_position_type)

#filtering to row with highest prediction value
probability_matrix_district_position_type <- probability_matrix_district_position_type %>% 
  group_by(row_number) %>% 
  filter(confidence == max(confidence))

#adding prediction and confidence level to original data
district_personnel_position_type <- district_personnel_position_type %>% 
  ungroup() %>% 
  mutate(predicted_value = probability_matrix_district_position_type$C_position_type,
         confidence = probability_matrix_district_position_type$confidence)

## Cleaning up the original data
district_data_export_position <- prediction_data_prep_exp(district_personnel)

district_data_export_position <- district_data_export_position %>% 
  left_join(district_personnel_position_type %>% select(district_data_concat, predicted_value, confidence),
            by = "district_data_concat")


corrected_predictions <- district_data_export_position$predicted_value

corrected_predictions[corrected_predictions == "Teacher"] <- "temp_teacher"
corrected_predictions[corrected_predictions == "Other"] <- "Teacher"
corrected_predictions[corrected_predictions == "Substitute"] <- "Other"
corrected_predictions[corrected_predictions == "temp_teacher"] <- "Substitute"

district_data_export_position$predicted_value <- corrected_predictions

```

## 8.4 Train Function Model

```{r}

district_personnel_function <- prediction_data_prep(district_personnel,
                                                         "C_function")

#making all district dataframes a single row per combination of values
district_personnel_function <- district_personnel_function %>% 
  group_by(district_data_concat) %>% 
  summarize()

# Prep the bags
district_personnel_function_test_bag <- bag_prep(district_personnel_function)

# Clean abreviations and columns
district_personnel_function_test_bag <- add_all_cols(district_personnel_function_test_bag)
district_personnel_function_test_bag <- fix_abbrevs(district_personnel_function_test_bag)

district_personnel_function_test_bag <- align_bags(training_bag_columns_function,
                                                        district_personnel_function_test_bag)
# Identify and add missing columns in the test data
district_personnel_function_test_bag <- add_missing_cols(training_bag_columns_function,                              district_personnel_function_test_bag)


#setting seed for replicability
set.seed(223)

#ordering columns the exact same order across training and test

  ## Get the column names except "C_position_type"
column_sorting <- sort(setdiff(training_bag_columns_function$'.', "C_function"))

new_column_order_test <- c(column_sorting)

function_district_test_bag <- district_personnel_function_test_bag[, new_column_order_test]


# Make predictions on the test data
rf_function_predictions_district <- predict(model_function,
                                                 as.matrix(function_district_test_bag))

#turning the predictions into a probability dataframe for each row in original data
probability_matrix_district_function <- data.frame(
  row_number = rep(c(1:nrow(district_personnel_function)),
                   each = 35),
  C_function = rep(c(0:34), nrow(district_personnel_function)),
  confidence = rf_function_predictions_district)

#shifting back to character values for position type
probability_matrix_district_function <-
  turn_function_into_values(probability_matrix_district_function)

#filtering to row with highest prediction value
probability_matrix_district_function <- probability_matrix_district_function %>% 
  group_by(row_number) %>% 
  filter(confidence == max(confidence))

#adding prediction and confidence level to original data
district_personnel_function <- district_personnel_function %>% 
  ungroup() %>% 
  mutate(predicted_value = probability_matrix_district_function$C_function,
         confidence = probability_matrix_district_function$confidence)


district_data_export_function <- prediction_data_prep_exp(district_personnel)

district_data_export_function <- district_data_export_function %>% 
  left_join(district_personnel_function %>% select(district_data_concat, predicted_value, confidence),
            by = "district_data_concat")

```

## 8.5 Merging predictions together

```{r}
## Change back any columns you changed in 8.1

district_data_export_function <- district_data_export_function %>%
  rename(`D_dept_desc` = `C_reporting`, `D_position_title_desc` = `D_position_title`)

district_data_export_position <- district_data_export_position %>%
  rename(`D_dept_desc` = `C_reporting`, `D_position_title_desc` = `D_position_title`)

shared_cols <- intersect(names(budgeted_positions_raw), names(district_data_export_position))
shared_cols <- intersect(shared_cols, names(district_data_export_function))


budgeted_positions_raw <- budgeted_positions_raw %>%
  # join position predictions
  left_join(
    district_data_export_position %>%
      select(all_of(shared_cols), ML_predicted_position_type = predicted_value),
    by = shared_cols
  ) %>%
  # join function predictions
  left_join(
    district_data_export_function %>%
      select(all_of(shared_cols), ML_predicted_function = predicted_value),
    by = shared_cols
  )


```

# Part 9: Upload File to Sharepoint

## Part 9.1 Any final clean up

Add any final clean up to your file below (remove unncessary columns, update column names to be standard etc.)

```{r}
processed_data_folder_path <- "https://erstrategies1.sharepoint.com/sites/orgfiles/Client%20Work/Other/Budget%20Benchmarking/2025%20Cohort/1.%20Data%20&%20Analysis%20-%20Secure/1.%20Data%20Processed"

#budgeted positions
ers_write_sharepoint(data = budgeted_positions_raw,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = paste0(district_folder,"_budgeted_positions_processed.csv"))


#Schools File
ers_write_sharepoint(data = school_raw,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = paste0(district_folder,"_schools_processed.csv"))

#location crosswalk
ers_write_sharepoint(data = location_crosswalk,
                     folder_path = processed_data_folder_path,
                     file_name_with_extension = paste0(district_folder,"_location_crosswalk.csv"))
```
